---
title: "Figure3_ME20250408_CM20240815_reporter_only_hopping"
author: "Mathias & Christine"
date: "2024-08-15"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---



#Goal
Plot the hopping-induced deletion clones for figure 3, combining data from the three separate experiments.
Also plot the FACS data from -116kb hopping replicate 1, where we see the mCh- mT high cells. 

## Notes for re-running
To re-run this code, change the paths in 'File paths' to the correct location of datasets.

# Setup
Load the libraries and set the parameters.

```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(cowplot)
library(rtracklayer)
library(tidyverse)
# library(dtplyr)
library(dplyr)
# library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(caTools)
library(ggrastr)
library(RColorBrewer)
library(RcppRoll)
library(plotly)
library(readxl)
library(ggpubr)
library(ggpmisc)
library(ggridges)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')


split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```

# Paths & parameters
```{r paths_parameters}
#datatag
datetag = paste0("ME",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/Sox2/Figure_hopping_deletions/analysis_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}

library(knitr)
opts_chunk$set(fig.width = 20, 
               fig.height = 8, 
               cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "/figures_deletion_clones/")) 
pdf.options(useDingbats = FALSE)
```

## File paths
```{r}
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"

#mapped and filtered integrations clones
path_ints_clones_E2221 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2344_mapping_clones/CM20231218_E2221_locations_all_mapped_clones.rds"
path_ints_clones_E2344 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2344_mapping_clones/CM20231214_E2344_locations_all_mapped_clones.rds"
path_int_clones_E2709 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2709_mapping_clones/CM20241213_E2709_tib_clones_annot_filt1.rds"


#flowcytometry data clones
path_fcs_files_E2221 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2217_FACS_clones/gated_fcs_files/plate_reader'
path_fcs_files_E2259 = '/DATA/usr/m.eder/projects/FACS_data/E2259_MK03_23_34A_del_clones/fcs_for_R/ME20230518_E2259_gated'
path_fcs_files_E2270 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2270_hopping_mTurq_6gates_delDE/fcs_for_R/clones'
path_fcs_files_E2709 = '/DATA/projects/Sox2/Figure_hopping_deletions/fcs_for_R/E2709'

path_link_well_clonename_E2221 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2217_FACS_clones/CM20230309_E2204_link_wellindex_clonename.txt"

#WT controls
path_WT_ctrl_E2250 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2250_CRISPR_Sox2_delBclones_pools/fcs_for_R/export_ME20230418_E2250_pools_WT_001_Single Cells.fcs"
path_WT_ctrl_E2254 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2254_CRISPR_delA_clones/fcs_for_R/export_ME20230426_E2254_WT_001_Single Cells.fcs"   
path_WT_ctrl_E2709 = "/DATA/projects/Sox2/Figure_hopping_deletions/fcs_for_R/E2709/WT_ctrl/export_WT_B11_B11_Single Cells.fcs"

```


#Annotations
```{r}

# Location of enhancer / gene
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")
Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")


landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")

#CTCF sites based on our own mapping
CTCF_mm10 <- import.bed(path_CTCF_sites)
CTCF_mm10.chr3 <- CTCF_mm10[seqnames(CTCF_mm10)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)

# Plotting ranges
prange_zoom_paper = c(34.59E6, 34.83E6)

# Color annotations
## gene
red = 'red'
## SCR
brown = "#ffb000"
```


# Import Data
import mapping data
```{r}
all_mapped_clones_E2344 = readRDS(path_ints_clones_E2344)
all_mapped_clones_E2221 = readRDS(path_ints_clones_E2221)
filt_mapped_clones_E2709 = readRDS(path_int_clones_E2709)

```

Load all FCS files
```{r}
##load the file names
fcs_files_E2221 = dir(path_fcs_files_E2221, full.names = T)
fcs_files_E2259 = dir(path_fcs_files_E2259, full.names = T)
fcs_files_E2270 = dir(path_fcs_files_E2270, full.names = T)
fcs_files_E2709 = dir(path_fcs_files_E2709, full.names = T)
```

## E2221
```{r load_set_of_fcs_files}
#wrangle sample names E2221
names(fcs_files_E2221) = split_string(str_remove(fcs_files_E2221, '.*/'), "_", 2, 3)

link_well_clonename_E2221 = 
  read_tsv(path_link_well_clonename_E2221)


pdata_tib_E2221 = tibble(files = fcs_files_E2221, Plate_well_new = names(fcs_files_E2221)) %>%
  #link clone names to wells
  left_join(link_well_clonename_E2221) %>%
  mutate(sample = paste0("clone_", `clone name`, "_34")) %>%
  #add mapping
  left_join(all_mapped_clones_E2221) %>%
  mutate(sample = str_replace(sample,"clone_", "clone_E2221_")) %>%
  select(-c("clone name", "Well index culture plate")) 
```

## E2259
Manually add the location of the 23_34A control, the fcs file from this control is already in the same folder as the deletion clones so will be automatically loaded.
```{r}

#wrangle sample names E2259
names(fcs_files_E2259) = split_string(str_remove(fcs_files_E2259, '.*/'), "_", 2, 9)

#Adding the control sample to the E2259
info_ctrl_sample_E2259 = tibble(sample = "ctrl_E2259_23_34A_untreated", chr = "chr3",start = 34643961, #add info of the control sample
                          strand = "+", cell_line = "23_34A", experiment = "E2259",
                          expected_clonetype = "insertion", observed_clonetype_mapping = "insertion")

pdata_tib_E2259 = tibble(files = fcs_files_E2259,  sample_name = names(fcs_files_E2259)) %>%
  mutate(cell_line = case_when(grepl("23_34A", sample_name) ~ "23_34A",
                               grepl("23_MK3_C1", sample_name) ~ "23_MK03",
                               .default = "unclear")) %>%
           filter(cell_line == "23_34A") %>%
           mutate(clone_well = str_remove(sample_name, ".*_"),
                  sample1 = paste0("clone_E2259_", cell_line, "_", clone_well),
                  sample = case_when(sample_name == "ME20230518_E2259_23_34A_clones_23_34A_ctrl" ~ "ctrl_E2259_23_34A_untreated",
                            .default = sample1)) %>%
  left_join(all_mapped_clones_E2344) %>%
  rows_update(info_ctrl_sample_E2259, by = "sample") 
```

## E2270
```{r}
#wrangle sample names E2270
names(fcs_files_E2270) = split_string(str_remove(fcs_files_E2270, '.*/'), "_", 2, 6)

info_ctrl_sample = tibble(sample = "ctrl_E2270_23_34A_untreated", chr = "chr3",start = 34643961, #add info of the control sample
                          strand = "+", cell_line = "23_34A", experiment = "E2270",
                          expected_clonetype = "insertion", observed_clonetype_mapping = "insertion")

pdata_tib_E2270 = tibble(files = fcs_files_E2270,  sample_name = names(fcs_files_E2270)) %>%
  mutate(sample1 = paste0("clone_E2270_23_34A_", split_string(sample_name, "_",5))) %>%
  mutate(sample = case_when(sample_name == "23_34A_ctrl_C7_C07" ~ "ctrl_E2270_23_34A_untreated",
                            .default = sample1)) %>%
   left_join(all_mapped_clones_E2344) %>%
   rows_update(info_ctrl_sample, by = "sample") 

```

## E2709
#Clones E2709 mCh-/mT- deletions
In this experiment we mapped: mCh-/mT- clones, two 'control' mCh-/mT+ clones (hopping deletions as in the previous 3 experiments), as well as some mCh+/mT- clones (not used for the manuscript). Here we process them and select the ones with confident mapping.

Mark the confident clones:

A good deletion clone has: 
 - one loc with ratio 1 on fw_only 
 - one loc with ratio 1 on rv_only
 - no other hits
 
A good insertion clone has:
 - one loc with ratio 1 on both arms
 - no other hits

```{r manually_check_hits}
tib_clones_annot_filt2 = filt_mapped_clones_E2709 %>% 
  group_by(sample, expected_clonetype) %>%
    #try to deal with all deletion clones that have a couple of fw reads in the home location
  mutate(mapped_arms_corrected = case_when(mapped_arms == "both_arms" & ratio_reads_1 < 0.1 ~ "rv_mostly",
                                           mapped_arms == "both_arms" & ratio_reads_2 < 0.1 ~ "fw_mostly",
                                           .default = mapped_arms
                                           ))

classify_for_checking = tib_clones_annot_filt2 %>%
  summarize(n_hit_locations = n(),
            n_home_locations = sum(!hopped),
            n_both_arms = sum(mapped_arms_corrected == "both_arms"),
            n_fw_only = sum(mapped_arms_corrected == "fw_only") + sum(mapped_arms_corrected == "fw_mostly"),
            n_rv_only = sum(mapped_arms_corrected == "rv_only") + sum(mapped_arms_corrected == "rv_mostly")
            ) %>%
  mutate(confident_clone = case_when(n_hit_locations == 1 & n_both_arms == 1 ~ "confident_insertion",
                                     n_hit_locations == 2 & n_fw_only == 1 & n_rv_only == 1 ~ "confident_deletion",
                                     .default = "manual_check"),
         unhopped = case_when(confident_clone == "confident_insertion" ~ n_home_locations == 1,
                              .default = NA))

manual_check_clones = classify_for_checking %>% filter(confident_clone == "manual_check") %>% pull(sample)

tib_clones_annot_filt2 %>% filter(sample %in% manual_check_clones)
```

mC-mT-cl1: couple reads in home location, 34757122 locations is bleedthrough from cl12. Most likely: no integration found.
mC-mT-cl11: 10 rv only reads in home location, no fw found. Most likely: no integration found.
mC-mT-cl2: 14 rv only reads in home location, max 2 fw found. 34757122 locations is bleedthrough from cl12, 35112784 is unique to this reporter. 35112784 could be real
mC-mT-cl3: no hit found (only 'hit' has no location, is probably home location)
mC-mT-cl4: rv 34836695 location (4 reads) is unique, could be real, but no fw location found (home location is also rv only). Unexplained clone
mC-mT-cl5: fw 34790410 location (2 reads) is unique, could be real deletion
mC-mT-cl6: rv in home location, no fw found. Most likely: no integration found.
mC-mT-cl9: fw 34917554 (6 reads) is unique, 19 rv reads in home location. Most likely: real deletion

ME-cl1: seems real integration on chr10 (43601999), but only rv reads. Fw area is not clearly unmappable. Fw library did have less reads after parsing (104 fw vs 60313). Total sequencing of both should be enough (4 Mb and 11 Mb). Conclusion:  weird clone.
ME-cl17: 34832184 is bleedthrough from cl24, 35819378  is the real integration (almost unique to this clone).
ME-cl19: 34832184, 35693227, 154546977 are bleedthrough. Clone is either unhopped, or no reporter.
ME-cl23: 34832184 is bleedthrough, real location: 29665779
ME-cl26: 34612885 seems bleedthrough, home location probably also. Clone is either unhopped, or no reporter.
ME-cl31: all three integrations seem real and (almost) unique to this clone. Three integrations, not useable
ME-cl4: chr2 32326111 seems a real location, but also many (281) rv reads in home location. Weird clone, not useable

mC-mT-cl8: is sorted as a deletion clone, has 1 fw and 14 rv reads in the home location, so not really a confident unhopped insertion. Could be unhopped, or no reporter, not useable. 
mC-mT-cl10: has only 2 reads, in 34757122. This is most likely bleedthrough from cl12 (many clones have this bleedthrough). Most likely no reporter? (and if it does, its the same as cl12, so useless anyways)

```{r}
probably_no_reporter = c('mC-mT-cl1', 'mC-mT-cl11', 'mC-mT-cl3', 'mC-mT-cl6', 'ME-cl19', 'ME-cl26', 'mC-mT-cl8', 'mC-mT-cl10')
multiple_integrations = c('ME-cl31')

weird_clones = c('mC-mT-cl4', 'ME-cl1', 'ME-cl4')

confident_real = c("ME-cl17", "ME-cl23",'mC-mT-cl9')
potential_hits = c('mC-mT-cl2', 'mC-mT-cl5')

```

Remove the weird clones, clones with probably no reporter, clones with multiple integrations
```{r}

tib_clones_annot_filt3 = tib_clones_annot_filt2 %>%
  filter(! sample %in% c(probably_no_reporter, multiple_integrations, weird_clones)) %>%
  #remove locations that are bleedthrough from different clones
  filter(!(sample %in% c('mC-mT-cl2','mC-mT-cl5', 'mC-mT-cl9') & start == 34757122)) %>%
  filter(!(sample %in% c('ME-cl17','ME-cl23') & start == 34832184)) %>%
  mutate(annotation_confidence = case_when(sample %in% potential_hits ~ 'low',
                                           sample %in% confident_real  ~ 'medium',
                                           .default = 'good'))
```


NOTE: only load the FACS files of the samples we need plust the controls
```{r Wrangle E2709}

tib_clones_annot_filt3_try = tib_clones_annot_filt3 %>% 
  filter(expected_clonetype == "deletion" & hopped) %>%
  filter(!sample == "mC-mT-cl12") %>% #this clone is partially mT+ after sorting, so it does not match the phenotype we want to assess -> might be a double clone somehow
  mutate(experiment = "E2709",
         cell_line = "23_34A") %>% 
  rename("n_reads_fw" = read_count_1,
         "n_reads_rv" = read_count_2) %>%
  mutate(start_del = start(landingPad_23),
         end_del = end)

tib_ctrl_samples = data.frame(sample = c("WT", "F121", "23_41"), expected_clonetype = "none", experiment = "E2709", cell_line = c("WT", "F121", "23_41"))

tib_clones_annot_filt3_try = tib_clones_annot_filt3_try %>% left_join(tib_ctrl_samples)

#wrangle sample names E2709
names(fcs_files_E2709) = split_string(str_remove(fcs_files_E2709, '.*/'), "_", 2, 5) %>%
  str_remove("_Single Cells\\.fcs")


info_ctrl_sample_E2709 = tibble(sample = c("ctrl_E2709_23_34A_untreated", "WT", "F121", "23_41"), chr = "chr3", start = 34643961, #add info of the control sample
                          strand = "+", cell_line = c("23_34A", "WT", "F121", "23_41"), experiment = "E2709",
                          expected_clonetype = c("insertion", "ctrl", "ctrl", "insertion"))





pdata_tib_E2709 = tibble(files = fcs_files_E2709, sample_name =  names(fcs_files_E2709)) %>% 
  mutate(sample = case_when(sample_name == "23_34A_C11_C11" ~ "ctrl_E2709_23_34A_untreated",
                            sample_name == "mCh-_mT-_GFP+_B2" ~ "mC-mT-cl2",
                            sample_name == "mCh-_mT-_GFP+_E2" ~ "mC-mT-cl5",
                            sample_name == "mCh-_mT-_GFP+_H2" ~ "mC-mT-cl7",
                            sample_name == "mCh-_mT-_GFP+_E3" ~ "mC-mT-cl9",
                            sample_name == "mCh-_mT+_GFP+_B1" ~ "mC-mTctrl1",
                            sample_name == "mCh-_mT+_GFP+_C1" ~ "mC-mTctrl2",
                            sample_name == "WT_B11_B11" ~ "WT",
                            sample_name == "F121_9_B12_B12" ~ "F121",
                            sample_name == "23_41_C12_C12" ~ "23_41",
                            .default = sample_name)) %>% 
  left_join(tib_clones_annot_filt3_try) %>%
  rows_update(info_ctrl_sample_E2709, by = "sample") %>% 
  #only keep the relevant clones
  mutate(observed_clonetype_mapping = case_when(expected_clonetype == "deletion" ~ "deletion",
                                                .default = NA))

```


# Load fcs
```{r}
  #combine tibbles
pdata_tib_merged = bind_rows(pdata_tib_E2221, pdata_tib_E2259, pdata_tib_E2270, pdata_tib_E2709) %>%
  #select only clones that are mapped 
  filter(!is.na(chr)) %>%
  select(c(files, sample, chr, start, strand, read_count, n_reads_fw, n_reads_rv, cell_line, experiment, expected_clonetype,
           observed_clonetype_mapping, start_del, end_del))

pdata_merged_df = data.frame(pdata_tib_merged)
rownames(pdata_merged_df) = pdata_merged_df$sample

#load data
flowset = flowCore::read.flowSet(files = pdata_merged_df$files, 
                                       alter.names = T, truncate_max_range = F,
                                 column.pattern = 'BL.B..530_30.A|YG.D..610_20.A|V.F..450_50.A|FSC.A|SSC.A')
flowCore::sampleNames(flowset) = pdata_merged_df$sample
pData(flowset)= pdata_merged_df

total_cell_count = as.numeric(flowCore::keyword(flowset, "$TOT")[,'$TOT'])
pData(flowset)$cell_count = total_cell_count

flowset_fluo = flowset[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A')]
colnames(flowset_fluo) = c("GFP", "mCherry", "mTurq")
rm(flowset)
```

# Preprocessing
## Get summary data
calculate median
```{r}
#calculate central tendency
median_exprs = t(sapply(1:length(flowset_fluo), function(i){
  apply(exprs(flowset_fluo[[i]]),2,median)
}))
rownames(median_exprs) <- sampleNames(flowset_fluo)
colnames(median_exprs) <- paste0("median_", colnames(exprs(flowset_fluo[[1]])))

median_exprs = as_tibble( median_exprs, rownames = 'sample') %>% left_join(pdata_tib_merged)

``` 

## get close WT control
These clonal data don't have a WT control. Based on the 23_34A fluorescence I picked a close WT control from a different experiment.
I pick the closest controls:
For E2204 (E2221) use E2250 (me20230418)
For E2259 use E2254 (ME20230426)
For E2270 use E2250

```{r}
flowset_WT_ctrls = flowCore::read.flowSet(files = c(path_WT_ctrl_E2250, path_WT_ctrl_E2254, path_WT_ctrl_E2709), alter.names = T, truncate_max_range = F)[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A', 'FSC.A', 'SSC.A')]
colnames(flowset_WT_ctrls) = c("GFP", "mCherry", "mTurq",'FSC', 'SSC')

autofluo_est = bind_rows(lapply(1:3, function(i){
  tibble(GFP_WT = median(exprs(flowset_WT_ctrls[[i]])[,'GFP']),
         mCherry_WT = median(exprs(flowset_WT_ctrls[[i]])[,'mCherry']),
         mTurq_WT = median(exprs(flowset_WT_ctrls[[i]])[,'mTurq']))
})) %>%
  mutate(autofluo_experiment =  c("E2250", "E2254", "E2709"),
         date = c("me20230418", "ME20230426", "ME20241114"))

```

## Normalize

```{r}

fcs_tib_clones = median_exprs %>%
  mutate(hopped = start != 34643961) %>%
  #link the estimated autofluorescence
  mutate(autofluo_experiment = case_when(experiment %in% c("E2204", "E2270") ~  "E2250",
                                         experiment == "E2259" ~ "E2254",
                                         experiment == "E2709" ~ "E2709")) %>%
  left_join(autofluo_est) %>%
  #subtract autofluorescence  
  mutate(median_mTurq_cor = median_mTurq - mTurq_WT,
         median_mCherry_cor = median_mCherry - mCherry_WT,
         median_GFP_cor = median_GFP - GFP_WT) %>%
  #normalize to GFP
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)


#normalize to 23_34A (to compare between measurement days)
standard_tib = fcs_tib_clones %>% filter(!(sample %in% c("23_41", "F121", "WT"))) %>%
  filter(! hopped) %>%
  group_by(experiment, start) %>%
  summarize(
    #GFP normalized
    mTurq_norm_standard = mean(mTurq_norm),
    mCherry_norm_standard = mean(mCherry_norm),
    #not GFP normalized
    GFP_standard = mean(median_GFP_cor),
    mCherry_standard = mean(median_mCherry_cor)) %>%
  select(-start)


expression_tib_cor = fcs_tib_clones %>%
  left_join(standard_tib) %>%
  mutate(
    #GFP normalized
    rel_norm_mTurq = mTurq_norm / mTurq_norm_standard,
    rel_norm_mCherry = mCherry_norm / mCherry_norm_standard,
    #not GFP normalized
    rel_GFP_cor = median_GFP_cor / GFP_standard,
    rel_mCherry_cor = median_mCherry_cor / mCherry_standard
    )

``` 

## Summarize per location
```{r}
clones_by_pos = expression_tib_cor %>%
  group_by(experiment, chr, start, strand, cell_line, observed_clonetype_mapping, start_del, end_del) %>%
  summarize(
    #GFP normalized
    rel_mTurq_norm_mean = mean(rel_norm_mTurq),
    rel_mCherry_norm_mean = mean(rel_norm_mCherry),
    #not GFP normalized
    rel_GFP_cor_mean = mean(rel_GFP_cor),
    rel_mCherry_cor_mean = mean(rel_mCherry_cor)
    )

```


# Plot range of deletions

```{r number_of_clones}
expression_tib_cor %>% filter(observed_clonetype_mapping == "deletion") %>%nrow()

``` 

```{r unique_locations}
del_locations = # just ordered by size, unique locations
  filter(clones_by_pos, start >= prange_zoom_paper[1] & start <= prange_zoom_paper[2] & 
           observed_clonetype_mapping == "deletion") %>%
  #remove E2709, is only for supplemental figure as it was a follow-up experiment
  filter(experiment != "E2709") %>%
  distinct(experiment, end_del) %>%  
  ungroup()%>%
  mutate(end_del_exp =  paste0(end_del, "_", experiment)) %>%
  #order the factor from high to low so it looks nice in the plot (y axis is built from bottom to top)
  mutate(end_del_exp_fct = as_factor(end_del_exp)) %>%
  mutate(end_del_exp_fct = reorder(end_del_exp_fct, -end_del)) 
nrow(del_locations)
```
There are 48 unique deletion events (unique deletions or from separate experiments)

```{r shorted_largest_del}
del_locations %>%
  mutate(del_length = end_del - start_del) %>%
  pull(del_length) %>% summary()

```
Shortest deletion: 8395 bp, longest 113294 

```{r plot_range_deletions_c, fig.width=6, fig.height=4}

## plot
plot_range_dels = ggplot(del_locations, aes(x = end_del, y = end_del_exp_fct)) + 
  # annotate enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = red) +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = red) +
  #annotate landing pad
  # geom_vline(xintercept = start(landingPad_23), col = 'darkgrey') +
  # #annotate CTCF sites
  # geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty = '11') +
  
  #plot data
  geom_segment(aes(x = 34.63E6, xend = start_del,
                   yend = end_del_exp_fct),
               col = 'grey30') +
  geom_segment(aes(x = end_del, xend =  34.775E6,
                   yend = end_del_exp_fct),
               col = 'grey30') +
    geom_segment(aes(x = start_del, xend = end_del,
                   yend = end_del_exp_fct),
               col = 'grey80', linetype = 'dotted') +
  geom_point(size = 1, col = '#0077BB') +
 
  
  #layout
  labs(x='Genomic coordinate (Mb)') +
  theme_classic(base_size = 16) +
  theme(legend.position = "bottom",
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y  = element_blank(),
        axis.line.y = element_blank()) +
  
  scale_x_continuous(breaks = scales::pretty_breaks(n = 4),
                     labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=c(34.63E6, 34.775E6),
                     expand = c(0,0)) 
plot_range_dels
```


# Plot expression deletions

```{r rel_mTurq_del_clones, fig.width = 6, fig.height = 4}

colScale_ins_del =  scale_color_manual(name = "observed_clonetype_mapping",
                                       values  = c(insertion = "grey60", deletion = '#0077BB'),
                                       aesthetics = c("col","fill")) 

plot_expr_dels = ggplot(filter(clones_by_pos, experiment != "E2709"), aes(x = start, y = rel_mTurq_norm_mean, col = observed_clonetype_mapping)) +
  # annotate enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = red) +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = red) +
  #annotate landing pad
  geom_vline(xintercept = start(landingPad_23), col = 'darkgrey') +
  # #annotate CTCF sites
  # geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty = '11') +
 
  #indicate scaled expression of 1 (23_34A)
  geom_hline(yintercept = 1, linetype = 'dashed')+ 
  
  #layout
  labs(x='Genomic coordinate (Mb)', y = 'relative reporter \n(mTurq) expression') +
  theme_classic(base_size = 16) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 4),
                     labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=c(34.63E6, 34.775E6),
                     expand = c(0,0)
                     ) +
  scale_y_continuous(limits=c(0,NA),
                     expand = expansion(mult = c(0,0.05))) +
  
  #points
  geom_point( )+
  colScale_ins_del
plot_expr_dels

```

```{r BC_del_clones_combined, fig.width = 10, fig.height = 4}

cowplot::plot_grid(plotlist = list(plot_range_dels, plot_expr_dels), 
                   align = 'h', axis = c('tb'), nrow = 1, rel_widths =  c(1, 1.5))
```
## Numbers expression value
Maximum expression deletions:

```{r}
clones_by_pos %>% filter(observed_clonetype_mapping == "deletion") %>% pull(rel_mTurq_norm_mean) %>% max()
```

Difference clone ins dels
Within 20kb of SCR (not including inside the SCR!)
```{r}
clones_by_pos %>% filter( chr == "chr3" & start >= start(enhancer)-20E3 & start <= start(enhancer)) %>%
  group_by(observed_clonetype_mapping) %>%
  summarize(mean_mT = mean(rel_mTurq_norm_mean)) %>%
  pivot_wider(names_from= 'observed_clonetype_mapping', values_from = 'mean_mT') %>%
  mutate(deletion / insertion)

```

All clones overlapping the insertion range (wihtin 23.4kb)

```{r}
clones_by_pos %>% filter( chr == "chr3" & start >= 34.73E6 & start <= start(enhancer)) %>%
  group_by(observed_clonetype_mapping) %>%
  summarize(mean_mT = mean(rel_mTurq_norm_mean)) %>%
  pivot_wider(names_from= 'observed_clonetype_mapping', values_from = 'mean_mT') %>%
  mutate(deletion / insertion)

```
Correlation expression position:

```{r}
clones_by_pos_dels = clones_by_pos %>% filter(observed_clonetype_mapping == "deletion") %>%
  mutate(dist_SCR = start(enhancer) - start)

cor.test( clones_by_pos_dels$dist_SCR, clones_by_pos_dels$rel_mTurq_norm_mean, method = "spearman")
```

# To pick clones for PCR/Sanger confirmation
I picked a couple of clones with different deletions, from E2204 (so I can use lysates from those plates), for Sanger confirmation.

Here, plot the locations of the insertion/deletions and the SNPs.
```{r clones_Sanger_confirm, fig.width = 4, fig.height = 2}
ggplot(filter(expression_tib_cor, sample %in% c('clone_E2221_1-54_34', 'clone_E2221_1-64_34', 'clone_E2221_3-55_34') ),
               aes(x = start, y = rel_norm_mTurq, col = observed_clonetype_mapping,
                   lab = sample)) +
  # annotate enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = red) +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = red) +
  #annotate landing pad
  geom_vline(xintercept = start(landingPad_23), col = 'darkgrey') +
  
  #annotate SNPs
  geom_vline(xintercept = c(34691039, 34733400, 34752890), col = 'darkgreen', linewidth = 0.8, lty = 'dashed') +
 
  #indicate scaled expression of 1 (23_34A)
  geom_hline(yintercept = 1, linetype = 'dashed')+ 
  
  #layout
  labs(x='Genomic coordinate (Mb)', y = 'relative mTurq') +
  theme_classic(base_size = 16) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 4),
                     labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=c(34.63E6, 34.775E6),
                     expand = c(0,0)
                     ) +
  scale_y_continuous(limits=c(0,NA),
                     expand = expansion(mult = c(0,0.05))) +
  
  #points
  geom_segment(data = filter(expression_tib_cor, sample %in% c('clone_E2221_1-54_34', 'clone_E2221_1-64_34')),
               aes(x = start_del, xend = end_del, yend = rel_norm_mTurq), col  = 'darkgrey')+
  geom_point(size = 2) +
  
  colScale_ins_del
```

# Plot ridgeline plots of clones
Many deletion clones show the exact same deletion

```{r ridgeline plots of mCh-/mT+ del clones}

samples_clones_deletions = pData(flowset_fluo) %>% 
  filter(observed_clonetype_mapping == "deletion" &
           cell_count >= 1000) %>%
  pull(sample)

samples_clones_deletions_idx = pData(flowset_fluo)$sample %in% samples_clones_deletions

# Arrange them by deletion size
#order them by deletion sizes
level_order_clones = pData(flowset_fluo[samples_clones_deletions_idx]) %>%
  arrange(end_del) %>% pull(name)

#

# Plot as dotplot
#plot
ggplot(flowset_fluo[samples_clones_deletions_idx], aes(x = mTurq, fill = end_del)) +
  geom_density_ridges(aes(y = factor(name, levels = level_order_clones)),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 2,
                      alpha = 0.75,
                      quantile_lines = TRUE, vline_width = 0.5) +
  #facet_wrap(~ experiment) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42, limits = c(0, 5*10^4)) +
  scale_y_discrete(expand = expansion( add = c(0, 1.5))) +
  facet_wrap(~ end_del) +
  # geom_vline(data = filter(peaks_linear_tib, exp_name %in% pData(flowset_fluo_good[samples_clones_23_delA_idx])$exp_name),aes(xintercept = mTurq_peak)) +
  theme_bw() +
  theme(legend.position = 'bottom',
        axis.title.y = element_blank(), 
        aspect.ratio = 1.5) 


```

## Plot bimodality of deletion clones as ridgline plots
on the y-axis: experiment number, clone name, coordinate
in illustrator change this to rep1-3, coordinate (without clone name)
color by identical coordinate
```{r S3C_ridgeline_hopping_del_clones}
ggplot(
  flowset_fluo[samples_clones_deletions_idx], 
  aes(
    x = mTurq, 
    y = interaction(sample, end_del),  # separate ridges per sample per deletion position
    fill = factor(end_del)
  )
) +
#facet_wrap(~ end_del) +
stat_density_ridges(
    scale = 2, 
    alpha = 0.8, 
    rel_min_height = 0.01,
    quantile_lines = TRUE,
    quantiles = c(0.5),
    color = "black"
  ) +
  ggcyto::scale_x_flowjo_biexp(
    widthBasis = -100, 
    breaks = c(0, 1000, 10000, 100000), 
    limits = c(0, 5*10^4)
  ) +
  theme_bw() +
  labs(
    y = "mTurq Fluorescence", 
    x = "Sample per Deletion (sample_id:end_del)"
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )

```


# E2709 mCh-/mT- deletion clones 
Data was loaded earlier, here i plot the locations of the deletions. 





### Plot mapped deletion clones ordered by del size
```{r S3E_range_del_clones_E2709}
clones_by_pos_E2709 = expression_tib_cor %>% 
  filter(experiment == "E2709") %>% 
  mutate(reporter_level = case_when(median_mTurq >= 2000 ~ "positive",
                                    median_mTurq < 2000 ~ "negative"))


clones_by_pos_E2709 %>% 
  filter(!sample %in% c("ctrl_E2709_23_34A_untreated", "mC-mT-cl12", "WT", "F121", "23_41")) %>%
  mutate(sample = factor(sample, levels = sample[order(end_del)])) %>%
  ggplot(aes(x = end_del, y = sample, col = reporter_level, txt1 = sample)) +

  # Genomic annotations
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +

  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = red) +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = red) +

  geom_vline(xintercept = start(landingPad_23), col = 'darkgrey') +
  # geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', linetype = 'dotted') +

  # Deletion segments
  geom_segment(aes(x = start_del,
                   xend = end_del,
                   y = sample,
                   yend = sample),
               linetype = "dashed", color = "black", linewidth = 0.3) +

  geom_point(size = 2) +

  labs(x = 'Genomic coordinate (Mb)', y = 'sample (ordered by deletion end)') +
  theme_classic(base_size = 16) +
  theme(legend.position = "bottom") +

  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 6),
    labels = scales::unit_format(scale = 1E-6, accuracy = 0.1, unit = NULL),
    limits = c(34590000, 35200000)
  ) +
  ggtitle("manually curated clones")

```


### Plot fluorescence values of deletion clones
```{r S3F_fluorescence_del_clones_E2709}
#E2709 clones

samples_E2709_clones_idx = pData(flowset_fluo)$experiment %in% c("E2709") & pData(flowset_fluo)$sample != "mC-mT-cl12" & pData(flowset_fluo)$sample != "ctrl_E2709_23_34A_untreated" & pData(flowset_fluo)$sample != "23_41"


pE2709_mT = 
ggplot(flowset_fluo[samples_E2709_clones_idx], aes(x = mTurq, fill = expected_clonetype)) +
  geom_density_ridges(aes(
    y = factor(sample, levels = c("WT", "F121", "mC-mTctrl2", "mC-mTctrl1", "mC-mT-cl5", "mC-mT-cl9", "mC-mT-cl7", "mC-mT-cl2"))
  ),
  size = 0.3,
  scale = 1.5,
  alpha = 0.75,
  quantile_lines = TRUE) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(0, 10^3, 10^4, 10^5),
                               pos = 4.0) +
  scale_y_discrete(expand = expansion(add = c(0, 2))) +
  theme_bw() +
  theme(legend.position = 'bottom',
        axis.title.y = element_blank())

pE2709_mT

```


