---
title: "Figure5_CM20250325_Sox2_CDS"
author: "Mathias & Christine"
date: "2025-03-25"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---
# Goal
Make the manuscript figures about the Sox2 CDS swapping (currently figure 5) and hopping for the revised manuscript.

Figure panel S5A (reduction in mCherry expression in SoxP insertion clones, originally 5A) is generated in the code for figure 2.


## Notes for re-running
To re-run this code, change the paths in 'File paths' to the correct location of datasets.
The expression scores are bootstrapped 5000 times, which takes a while to run. Therefore, these chunks are commented out and the bootstrapped data is loaded from a local RDS file. To re-run those, run the bootstrapped expression score calculation once and save the result locally. The bootstrapped expression score for the Sox2P reporter is calculated (and saved) in the scripts for figure 2.


# Setup
```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
# library(dtplyr)
library(dplyr)
# library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(RcppRoll)
# library(plotly)
library(readxl)
library(ggpubr)
library(ggpmisc)
library(ggridges)
library(import)
library(ggh4x)
library(rstatix)
# library(flowCore)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(flowDensity)
library(cowplot)
library(GENOVA)
library(Signac)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}


```

```{r paths_parameters}
#datatag
datetag = paste0("CM",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/Sox2/Figure_Sox2_CDS/analysis_", datetag)
dir.create(output_dir, showWarnings = FALSE)

library(knitr)
opts_chunk$set(
               cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```

## File paths
```{r}
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"

path_annotation_E2410 = "/DATA/usr/m.eder/projects/FACS_data/E2410/E2410_me20231219_plate_reader/me20232323_annotation_file.tsv"
path_fcs_files_E2410 = '/DATA/usr/m.eder/projects/FACS_data/E2410/E2410_me20231219_plate_reader/export_fcs_single_cells/'

path_annotation_E2427 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2427_Sox2_CDS_swap/CM20240226_E2427_annotation_table_clones_pools.xlsx"
path_fcs_files_E2427 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2427_Sox2_CDS_swap/fcs_for_R'

path_annotation_E2447 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2447_C9_mCh+_41_clones/CM20240226_E2447_annotation_table_clones.xlsx"
path_fcs_files_E2447 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2447_C9_mCh+_41_clones/fcs_for_R'


path_annotation_E2742 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2742_23_mChpos_CDS_flipped_clones/CM20250225_E2742_final_PCR_genotyping.txt'
path_fcs_files_E2742 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2742_23_mChpos_CDS_flipped_clones/fcs_for_R'


path_annotation_E2759 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2759_mChneg_CDS_flipped_clones/CM20250324_E2759_genotyping.xlsx' 
path_fcs_files_E2759_round3 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2759_mChneg_CDS_flipped_clones/fcs_for_R/fcs_round3' 


#hopping
path_tib_C9_34_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-161kb_Sox2P_pools.txt"
path_tib_C9_41_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-161kb_Sox2P_CDS_pools.txt"

path_median_mTurq_relevant_cell_lines =  "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250318_median_mTurq_sorting_for_paper.tsv"
path_sorting_stats = "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250327_stats_for_expression_score_for_paper.xlsx"

path_expr_score_bootstrap_C9_mCh_34 = "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_sum_filt_C9_mCh_34_5kb.rds"

#sorting data hopping
diva_xml_file = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2534_hopping_C9_LP/FACS/Mathias (BvS)20240501_E2534_rep1.xml"
diva_xml_file2 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2534_hopping_C9_LP/FACS/Mathias (BvS)20240502_E2534_rep2.xml"
path_FACS_sorting_E2555 =  "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2534_hopping_C9_LP/FACS"

#RNA stability
path_qPCR_RNA_stability = "/DATA/projects/Sox2/Figure_Sox2_CDS/me20241217_E2682_all_rep_combined.xlsx"

#ChIP
path_JASPAR_motifs = '/DATA/projects/Sox2/Figure_Sox2_CDS/JASPAR2024_CORE_vertebrates_redundant_pfms_jaspar.txt'
path_JASPAR_mapped_motifs_Sox2 = '/DATA/projects/Sox2/Figure_Sox2_CDS/JASPAR2024_chr3_34630000-34680000'
path_cistrome_overview_table = '/DATA/projects/Sox2/cistrome_mm10/mouse_factor_full_QC.txt'
path_cistrome_peak_folder = "/DATA/projects/Sox2/cistrome_mm10/mouse_factor/"
path_cistrome_bw_folder ='/DATA/projects/Sox2/cistrome_bw/'
path_H3K27ac_ChIP = "/DATA/usr/c.moene/public_data/ChIP/Histons/H3K27ac_ENCFF583WVZ_signal_pval.bigWig"


#functions
source("/DATA/projects/Sox2/General_functions/CM20250318_new_expression_score_functions.R")


```

# Annotations
## color settings
```{r color_settings}
#colors for hopped integrations (P1-P6)
myCols_blue = c('#525252', "#080E5B",  "#212B71", "#3A4987", "#465893", "#5A719A", "#6C8AA0")

myCols_orange = c('#525252' ,'#a0430d', "#bb4e10", '#ec6518',  '#EE7733', "#f1925c", '#f3a070', '#e8aa86')
colScale7_orange <- scale_colour_manual(name = "population", values = myCols_orange, aesthetics = c("colour", "fill"))

#colors for the different inserts
myCols_inserts_named = c(`minimal insert` = "grey30", HyTK= "grey50", none = "black",
                   `Sox2P` = '#0077BB', `Sox2P_CDS_UTR`='#bd4d0e' , `Sox2P_CDS`= '#EE7733' )
col_scale_CDS_named = scale_colour_manual(name = "reporter", values = myCols_inserts_named,
                                          aesthetics = c("colour", "fill"))

myCols_insert_groups = c(`minimal insert` = "grey30", none = "black",
                   `Sox2P` = '#0077BB', `Sox2P_CDS_UTR`='#bd4d0e' , `Sox2P_CDS`= '#EE7733' )
col_scale_CDS_groups = scale_colour_manual(name = "insert_group", values = myCols_insert_groups,
                                          aesthetics = c("colour", "fill"))


myCols_insert_cell_lines = c(`C9_mCh_34` = '#0077BB', `C9_mCh_41` = '#EE7733'  )
col_scale_CDS_cl = scale_colour_manual(name = "cell_line", values = myCols_insert_cell_lines,
                                    aesthetics = c("colour", "fill"))

myCols_insert_cell_lines_FACS = myCols_insert_cell_lines
names(myCols_insert_cell_lines_FACS) = c('C9_34_mCh+' , 'C9_41_mCh+')
col_scale_CDS_cl_FACS = scale_colour_manual(name = "cell_line", values = myCols_insert_cell_lines_FACS,
                                    aesthetics = c("colour", "fill"))



## SCR
brown = "#ffb000"
```

## Locations
```{r genomic_annotations}

# Location of enhancer / gene
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")
Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")

landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")

landingPad_C9 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34598479,
                                     end = 34598480),
                    strand = "+")

LP_tib = tibble(landing_pad = c("23", "C9"), 
                start = c( start(landingPad_23),start(landingPad_C9)))

#CTCF sites based on our own mapping
CTCF_mm10 <- import.bed(path_CTCF_sites)
CTCF_mm10.chr3 <- CTCF_mm10[seqnames(CTCF_mm10)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)

# Plotting ranges
prange_zoom = c(34.5E6, 35E6)
prange_zoom_paper = c(34.59E6, 34.83E6)

```
## others
```{r}
N_bootstrap = 500
```


# Load data 
## process annotations
### E2410
```{r load_E2410}

annotation_file_E2410 = read_tsv(path_annotation_E2410)


fcs_files_E2410_plateReader = dir(path_fcs_files_E2410, full.names = T)
names(fcs_files_E2410_plateReader) = str_remove(str_remove(str_remove(fcs_files_E2410_plateReader, '.*/'),"_Single Cells.fcs"), "export_")


matched_annotation_tib_E2410 = tibble(filenames = names(fcs_files_E2410_plateReader),
                                      full_filename = fcs_files_E2410_plateReader,
                                      # N_char = str_length(filenames),
                                      well_tmp = str_extract(filenames, "..._...$"),
                                      row = substr(well_tmp, 1, 1),
                                      col = as.numeric(substr(well_tmp, 2, 3)),
                                      well = paste0(row, col),
                                      WELL = substr(well_tmp, 1, 3)) %>%
  select(-c(well_tmp)) %>%
  left_join(annotation_file_E2410) %>%
  mutate(
    date = "ME20231219",
    # case_when(col %in% c(10, 11, 12) & row %in% c("E", "F", "G") ~ "control", 
    experiment = "E2410",
    type = case_when(col %in% c(1,2,3) ~ "pool",
                     col %in% c(4,5,6) & row %in% c("E", "F", "G", "H") ~ "pool",
                     col %in% c(10, 11, 12) & row %in% c("E", "F", "G") ~ "control",
                     .default = "clone"),
    expected_insert = case_when(constructs %in% c("34", "40", "41") ~ paste0("pME0", constructs),
                                constructs == "UT" ~ "HyTK",
                                constructs == "ctrl" ~ "HyTK",
                                constructs == "WT" ~ "none",
                                .default = "pME034"),
    landing_pad = case_when(cell_line %in% c("8", "23", "C9","C10") ~ cell_line,
                            cell_line == "WT" ~ "none"),
    GFP_state = case_when(cell_line == "WT" ~ "GFP-",
                          .default = "GFP+"),
    mCh_state = case_when(cell_line == "WT" ~ "mCh-",
                          .default = "mCh+"),
    selection = case_when(GV_selected == "yes" ~ "Gv",
                          col %in% c(1,2,3) ~ "FACS"),
    clone_number = case_when(selection == "Gv" & type == "clone" ~ split_string(clone_names, "_", 3))) %>%
  mutate(
    name = paste0(case_when(!is.na(selection) & type == "pool" ~ paste0("E2410_orig_", selection, "_pool_", landing_pad, "_", mCh_state, "_", expected_insert),
                            !is.na(selection) & type == "clone" ~ paste0("E2410_orig_", selection, "_clone_", landing_pad, "_", mCh_state, "_", expected_insert, "_", clone_number),
                            type == "control" ~ case_when(constructs == "WT" ~ "WT",
                                                                constructs == "23_34A_ctrl" ~ "23_34A",
                                                                constructs == "8_34_8_ctrl" ~ "8_34-8"))
                  )) %>%
  select(-c("WELL", "cell_line", "GV_selected", "clone_names", "constructs")) %>%
  #select only the clones (we don't use the pools anymore)
  filter(type %in% c("clone", "control")) %>%
  mutate(insert_genotyped = "good") %>% #see E2568, all clones in this experiment have the correct insert
  select(-selection, clone_number)
  
```

### E2427
```{r load_E2427}
annotation_tib_E2427 = read_xlsx(path_annotation_E2427)

fcs_files_E2427 = dir(path_fcs_files_E2427, full.names = T)
names(fcs_files_E2427) = str_remove(str_remove(fcs_files_E2427, '.*/'),"_Single Cells2.fcs")

matched_annotation_tib_E2427 = tibble(filenames = names(fcs_files_E2427),
                                full_filename = fcs_files_E2427,
                                date = "CM20240206",
                                plate = split_string(filenames, "_", 2),
                                # N_char = str_length(filenames),
                                well_tmp = str_extract(filenames, "...$"),
                                row = substr(well_tmp, 1, 1),
                                col = as.numeric(substr(well_tmp, 2, 3)),
                                well = paste0(row, col)) %>%
  select(-c(row, col, well_tmp)) %>%
  left_join(annotation_tib_E2427) %>%
  #remove all remeasurements from E2410 (don't need them, just confusing) as well as the remeasured clones from E2350/ pools from E2263
  filter(!experiment %in% c("E2410", "E2350","E2263")) %>%
  mutate(expected_insert = case_when(expected_insert %in% c("34","37", "40", "41") ~ paste0("pME0", expected_insert),
                                     .default = expected_insert),
         type = case_when(experiment == "control" ~ 'control',
                          .default = type),
         experiment = "E2427") %>%
 #select only the clones (we don't use the pools anymore)
  filter(type == "clone" & insert_genotyped == "good" | type == "control")%>%
  select(-c(deletion, note_genotyping, selection, details))


``` 

### E2742
```{r load_E2742}
annotation_file_E2742 = read_tsv(path_annotation_E2742) %>%
  replace_na(list(Selection_mCh  = "", Selection_PCR = "", Comments = "")) %>%
  #select genotyped clones
  filter(Selection_mCh != "mCh_neg" & !Selection_PCR %in% c("no_band")) %>%
  mutate(insert_genotyped = 'good',
         filenames = str_remove( filename, '.fcs'))%>%
  select(-c(Comments, Selection_mCh, Selection_PCR, filename)) 

#there was an error in the export of the controls, 3 filenames got shifted
#here I correct the filenames (which I checked with the exported pdf)
annotation_E2742_controls = tibble(filenames = c("23_34A_G3_G03", "C9_34_G4_G04", "WT_G1_G01"),
                                   sample_corrected = c("F121/9", "23_34A", "WT"))

annotation_file_E2742 = bind_rows(annotation_file_E2742, annotation_E2742_controls)

fcs_files_E2742 = dir(path_fcs_files_E2742, full.names = T, recursive = T)
names(fcs_files_E2742) = str_remove(str_remove(fcs_files_E2742, '.*/export_'),"_Single Cells2.fcs")

filename_E2742_tib = tibble(filenames = names(fcs_files_E2742),
                                full_filename = fcs_files_E2742,
                                date = "CM20250204",
                                expected_insert = case_when(grepl('pCM', filenames)|grepl('pME', filenames) ~ str_remove(filenames, "_.*$")),
                                well = case_when(grepl('pCM', filenames)|grepl('pME', filenames) ~ str_remove(str_remove(filenames, "^.*_"), '.fcs'))) 

matched_annotation_tib_E2742 = annotation_file_E2742 %>% 
  left_join(filename_E2742_tib) %>%
  mutate(expected_insert = case_when(sample_corrected %in% c("WT", "F121/9" )~ "none",
                            sample_corrected == "23_34A" ~ "pME034",
                            .default = insert),
         type = case_when(sample_corrected %in% c("WT", "F121/9", "23_34A") ~ "control",
                          .default = "clone"),
         landing_pad = case_when(sample_corrected %in% c("WT", "F121/9") ~ "none",
                                 .default = "23"),
         mCh_state = case_when(sample_corrected %in% c("WT" ) ~ "mCh-",
                               .default = "mCh+"),
         GFP_state = case_when(sample_corrected %in% c("WT" ) ~ "GFP-",
                               .default = "GFP+"),
         experiment = "E2742",
         name = sample_corrected
  ) %>%
  select(-insert)


```

### E2759
```{r load_E2759}
annotation_file_E2759 = read_xlsx(path_annotation_E2759)

fcs_files_E2759_round3 = dir(path_fcs_files_E2759_round3, full.names = T)
names(fcs_files_E2759_round3) = str_remove(str_remove(str_remove(fcs_files_E2759_round3, '.*/'),"_Single Cells2.fcs"), "export_")

file_loc_tib_E2759 = enframe(fcs_files_E2759_round3)%>%
  dplyr::rename(full_filename = value) %>%
  mutate(Round = 'round3',
         sample_name_link = paste0(Round, "_", name)) %>%
  select(-name)

matched_annotation_tib_E2759 = annotation_file_E2759 %>%
  filter(Round == "round3") %>%
  mutate(sample_name_link = paste0(Round, "_",str_remove(Sample, '.fcs'))) %>%
  left_join(file_loc_tib_E2759) %>%
  mutate(date = "ME20250320") %>%
  dplyr::rename(filenames = sample_name_link) %>%
  dplyr::mutate(landing_pad = factor(landing_pad, levels = c("C9", "23")),
         expected_insert = case_when(expected_insert == "M34" ~ "pME034",
                            expected_insert == "M41" ~ "pME041",
                            expected_insert == "C38" ~ "pCM038",
                            expected_insert == "C39" ~ "pCM039",
                            expected_insert == "none" ~ "none"),
         experiment = "E2759")



```


## combine experiments
```{r combine_experiments}
#combine and rename inserts/landing pads
matched_annotation_comb = bind_rows(matched_annotation_tib_E2427, matched_annotation_tib_E2410,matched_annotation_tib_E2742, matched_annotation_tib_E2759) %>%
  mutate(insert_name = factor(expected_insert, levels = c( "none", "HyTK", "pME037", "pME034", "pME041", 
                                                           paste0("pCM0", 38:41), "pME040",paste0("pCM0", 43:45)),
                                     labels = c(  "none","HyTK", "minimal insert", "Sox2P_mTurq", "Sox2P_CDS_mTurq",
                                                  "Sox2P_mTurq_CDS", "Sox2P_mTurq_stop_CDS", "Sox2P_mTurq_PAS_CDS", "Sox2P_mTurq_revCDS", 
                                                   "Sox2P_CDS_mTurq_UTR",
                                                  "Sox2P_mTurq_CDS_UTR", "Sox2P_mTurq_stop_CDS_UTR", "Sox2P_mTurq_PAS_CDS_UTR")),
         landing_pad = factor(landing_pad, levels = c("none", "C9", "23", "8"),
                              labels = c("no insert", "-161 kb", "-116 kb", "-39 kb")),
         insert_group = factor(case_when(expected_insert  == "pME037" ~ "minimal insert",
                                  expected_insert  == "pME034" ~ "Sox2P",
                                  expected_insert  %in% c("pME041", paste0("pCM0", 38:41)) ~ "Sox2P_CDS",
                                  expected_insert  %in% c("pME040", paste0("pCM0", 43:45)) ~ "Sox2P_CDS_UTR",
                                  .default = expected_insert)),
         insert_variant = case_when(expected_insert %in% c("pME037", "pME034", "pME040", "pME041") ~ "original",
                                    expected_insert %in% c("pCM038", "pCM043") ~ "P2A",
                                    expected_insert %in% c("pCM039", "pCM044") ~ "STOP",
                                    expected_insert %in% c("pCM040", "pCM045") ~ "PAS",
                                    expected_insert == "pCM041" ~ "revCDS"))

```

## load fcs files
```{r}
#make dataframe
matched_annotation_df_comb = data.frame(matched_annotation_comb)
rownames(matched_annotation_df_comb) = matched_annotation_df_comb$filenames

#read data
flowset_comb = flowCore::read.flowSet(files = matched_annotation_df_comb$full_filename, alter.names = T, truncate_max_range = F, ignore.text.offset = T)

flowCore::sampleNames(flowset_comb) = matched_annotation_df_comb$filenames
flowCore::pData(flowset_comb) = matched_annotation_df_comb

flowset_fluo_comb = flowset_comb[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A')]
colnames(flowset_fluo_comb) = c("GFP", "mCherry", "mTurq")

```


```{r preprocess_fcs_data_comb}
total_cell_count_tmp = flowCore::keyword(flowset_fluo_comb, "$TOT")[,'$TOT']
total_cell_count = as.numeric(total_cell_count_tmp)
names(total_cell_count) = names(total_cell_count_tmp)

median_mTurq = sapply(1:length(flowset_fluo_comb), function(i){
  median(flowCore::exprs(flowset_fluo_comb[[i]])[,'mTurq'])
})
names(median_mTurq) = sampleNames(flowset_fluo_comb)

median_mCherry = sapply(1:length(flowset_fluo_comb), function(i){
  median(flowCore::exprs(flowset_fluo_comb[[i]])[,'mCherry'])
})
names(median_mCherry) = sampleNames(flowset_fluo_comb)


median_GFP = sapply(1:length(flowset_fluo_comb), function(i){
  median(flowCore::exprs(flowset_fluo_comb[[i]])[,'GFP'])
})
names(median_GFP) = sampleNames(flowset_fluo_comb)

flowCore::pData(flowset_fluo_comb)$median_mTurq = median_mTurq
flowCore::pData(flowset_fluo_comb)$median_mCherry = median_mCherry
flowCore::pData(flowset_fluo_comb)$median_GFP = median_GFP

flowCore::pData(flowset_fluo_comb)$total_cell_count = total_cell_count

fcs_tib = tibble(flowCore::pData(flowset_fluo_comb))
```


# Normalize expression
```{r normalize_expression_values}
#subtract autofluorescence, normalize to GFP
autofluo_tib_comb = fcs_tib  %>%
  group_by(date) %>%
  filter(name == "WT") %>%
  summarize(mTurq_WT = mean(median_mTurq),
           mCherry_WT = mean(median_mCherry),
            GFP_WT = mean(median_GFP) )

expression_tib_comb = fcs_tib %>%
  left_join(autofluo_tib_comb) %>%
  mutate(median_mTurq_cor = median_mTurq - mTurq_WT,
         median_mCherry_cor = median_mCherry - mCherry_WT,
         median_GFP_cor = median_GFP - GFP_WT) %>%
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)

#normalize to 23_34A (to compare between measurement days)
standard_tib = expression_tib_comb %>%
  group_by(date) %>%
  filter(name == "23_34A") %>%
  summarize(mTurq_norm_standard = mean(mTurq_norm),
           mCherry_norm_standard = mean(mCherry_norm),
            GFP_standard = mean(median_GFP_cor) )

expression_tib_comb = expression_tib_comb %>%
  left_join(standard_tib) %>%
  mutate(rel_mTurq = mTurq_norm / mTurq_norm_standard,
         rel_mCherry = mCherry_norm / mCherry_norm_standard,
         rel_GFP_cor = median_GFP_cor / GFP_standard)

``` 

# Plot clones
To compare the expression values, we use the Welch's t test, this assumes that the values are normally distributed, but does *not* assume equal variances.


## relative mTurq 
### mCh+ clones, old CDS reporters
Plot mChpos swapped clones with the original CDS reporters (from the first 2 experiments). 
I realized in the original plot we did not filter for >= 500 recorded cells, when we add that filter (which is also used in other plots), we lose 2 more clones for LP8. Below is the new, more correctly filtered plot.

```{r, fig_5B_rel_mT_mChpos_oldCDS, fig.width=6, fig.height = 4}

filter(expression_tib_comb, type %in% c("clone")  & experiment %in% c("E2410", "E2427") &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%

  #plotting
  ggplot(aes(x = insert_name, y = rel_mTurq, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
    comparisons = list(c("Sox2P_mTurq", "Sox2P_CDS_mTurq"), c("Sox2P_mTurq", "Sox2P_CDS_mTurq_UTR"),
                                        c("Sox2P_CDS_mTurq", "Sox2P_CDS_mTurq_UTR")),
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
    tip.length = 0.02) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq, clones from both experiments") +
  ylab("relative reporter expression") +
  col_scale_CDS_groups
``` 


Also slightly altered numbers because we now filter for >= 500 recorded cells
```{r, rel_mT_mChpos_numbers_oldCDS}
median_expr_values_tib = filter(expression_tib_comb, type %in% c("clone")  & experiment %in% c("E2410", "E2427") &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
 
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%
  
  #calculate median per condition
  group_by(landing_pad, insert_name) %>%
  summarize(median_rel_mTurq = median(rel_mTurq),
            median_rel_mCherry = median(rel_mCherry))
  
#calculate fold change mTurq Sox2P vs CDS
median_expr_values_tib %>%
  select(-median_rel_mCherry) %>%
  pivot_wider(names_from = "insert_name", values_from = "median_rel_mTurq")%>%
  mutate(fc_Sox2P_CDS = Sox2P_CDS_mTurq / Sox2P_mTurq,
         fc_Sox2P_CDS_UTR = Sox2P_CDS_mTurq_UTR / Sox2P_mTurq)
  
#calculate fold change mCherry each vs minimal insert
median_expr_values_tib %>%
  select(-median_rel_mTurq ) %>%
  pivot_wider(names_from = "insert_name", values_from = "median_rel_mCherry")%>%
  mutate(fc_Sox2P = Sox2P_mTurq / `minimal insert`,
         fc_Sox2P_CDS = Sox2P_CDS_mTurq / `minimal insert`,
         fc_Sox2P_CDS_UTR = Sox2P_CDS_mTurq_UTR / `minimal insert`) %>%
  mutate(fc_Sox2P_perc = 100*(1-fc_Sox2P),
         fc_Sox2P_CDS_perc = 100*(1-fc_Sox2P_CDS),
         fc_Sox2P_CDS_UTR_perc = 100*(1-fc_Sox2P_CDS_UTR)
  )
  
``` 

### Fig 5G - mCh+ clones, new reporters
```{r, fig_5G_rel_mT_mChpos_newCDS, fig.width=5.4, fig.height = 4}

filter(expression_tib_comb, type %in% c("clone")  & experiment == "E2742" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" & #C9_mCh- and 23_mCh- are not relevant for this comparison
         insert_group %in% c("minimal insert", "Sox2P", "Sox2P_CDS")) %>% #leave UTR for supplement
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%

  #plotting
  ggplot(aes(x = insert_name, y = rel_mTurq, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +


  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq, new reporters") +
  ylab("relative reporter expression") +
  col_scale_CDS_groups

``` 

### Fig S6I - rel_mT_mChpos_newCDS_UTR
```{r, fig_S6I_rel_mT_mChpos_newCDS_UTR, fig.width=6, fig.height = 4}

tib_for_plot = filter(expression_tib_comb, type %in% c("clone")  & experiment == "E2742" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-"  #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>% #leave UTR for supplement
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") 
 

#plotting
ggplot(tib_for_plot, aes(x = insert_name, y = rel_mTurq, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 2) +
  # geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_mCh), linetype = 'dashed')+
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, new reporters") +
  ylab("relative Sox2::mCherry expression") +
  col_scale_CDS_groups


``` 

### Fig 5H - mCh- clones, old
```{r, fig_5H_rel_mT_mChneg_oldCDS, fig.width=4.5, fig.height = 4}

filter(expression_tib_comb, type == "clone"  & 
         experiment == "E2427" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state == "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%
  
  #plotting
  ggplot(aes(x = insert_name, y = rel_mTurq, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
    comparisons = list(c("Sox2P_mTurq", "Sox2P_CDS_mTurq"), c("Sox2P_mTurq", "Sox2P_CDS_mTurq_UTR"),
                                        c("Sox2P_CDS_mTurq", "Sox2P_CDS_mTurq_UTR")),
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE), # t.test var.equal = F gives a Welch's t-test
    tip.length = 0.02) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq for mCh- clones") +
  ylab("relative reporter expression") +
  col_scale_CDS_groups

``` 

```{r, rel_mT_mChneg_numbers_oldCDS}
mean_expr_values_tib_mChneg = filter(expression_tib_comb, type == "clone"  & 
         experiment == "E2427" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state == "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%

  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%
  
  #calculate mean per condition
  group_by(landing_pad, insert_name) %>%
  summarize(mean_rel_mTurq = mean(rel_mTurq))
  
#calculate fold change mTurq Sox2P vs CDS
mean_expr_values_tib_mChneg %>%
  pivot_wider(names_from = "insert_name", values_from = "mean_rel_mTurq")%>%
  mutate(fc_Sox2P_CDS = Sox2P_CDS_mTurq / Sox2P_mTurq,
         fc_Sox2P_CDS_UTR = Sox2P_CDS_mTurq_UTR / Sox2P_mTurq)

``` 

### Fig S6M - mCh- clones, new
```{r, fig_S6M_rel_mT_mChneg_newCDS, fig.width=4.5, fig.height = 4}

filter(expression_tib_comb, type == "clone"  & 
         experiment == "E2759" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state == "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%
  
  #plotting
  ggplot(aes(x = insert_name, y = rel_mTurq, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 2) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
      ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq for mCh- clones") +
  ylab("relative reporter expression") +
  col_scale_CDS_groups


``` 

```{r, rel_mT_mChneg_numbers_newCDS}
mean_expr_values_tib_mChneg = filter(expression_tib_comb, type == "clone"  & 
         experiment == "E2759" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state == "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%

  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%
  
  #calculate mean per condition
  group_by(landing_pad, insert_name) %>%
  summarize(mean_rel_mTurq = mean(rel_mTurq))
  
#calculate fold change mTurq Sox2P vs CDS
mean_expr_values_tib_mChneg %>%
  pivot_wider(names_from = "insert_name", values_from = "mean_rel_mTurq")%>%
  mutate(fc_Sox2P_CDS_mTurq = Sox2P_CDS_mTurq / Sox2P_mTurq,
         fc_Sox2P_mTurq_CDS = Sox2P_mTurq_CDS / Sox2P_mTurq,
         fc_Sox2P_mTurq_stop_CDS = Sox2P_mTurq_stop_CDS / Sox2P_mTurq,
         )

``` 

## relative mCherry
### Fig 5B - mCh+ clones, old CDS rep
```{r, fig_5B_rel_mCh_mChpos_oldCDS, fig.width=6, fig.height = 4}

tib_for_plot = filter(expression_tib_comb, type %in% c("clone")  & experiment %in% c("E2410", "E2427") &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good")
 
ref_dash_lines = tib_for_plot %>%
  filter(insert_group == "minimal insert") %>%
  group_by(landing_pad) %>%
  summarize(ref_rel_mCh = mean(rel_mCherry))

#plotting
ggplot(tib_for_plot, aes(x = insert_name, y = rel_mCherry, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 2) +
  geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_mCh), linetype = 'dashed')+
  # geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
    comparisons = list( c("minimal insert", "Sox2P_mTurq"), c("minimal insert", "Sox2P_CDS_mTurq"), c("minimal insert", "Sox2P_CDS_mTurq_UTR"),
                        c("Sox2P_CDS_mTurq", "Sox2P_CDS_mTurq_UTR")),
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
    tip.length = 0.02) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, clones from both experiments") +
  ylab("relative Sox2::mCherry expression") +
  col_scale_CDS_groups


``` 

### Fig 5G - mCh+ clones, new reporters
```{r, fig_5G_rel_mCh_mChpos_newCDS, fig.width=5.4, fig.height = 4}
tib_for_plot = filter(expression_tib_comb, type %in% c("clone")  & experiment == "E2742" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" & #C9_mCh- and 23_mCh- are not relevant for this comparison
         insert_group %in% c("minimal insert", "Sox2P", "Sox2P_CDS")) %>% #leave UTR for supplement
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") 
 
ref_dash_lines = tib_for_plot %>%
  filter(insert_group == "minimal insert") %>%
  group_by(landing_pad) %>%
  summarize(ref_rel_mCh = mean(rel_mCherry))

#plotting
  ggplot(tib_for_plot, aes(x = insert_name, y = rel_mCherry, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 2) +
  geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_mCh), linetype = 'dashed')+
  # geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
     ref.group = "minimal insert",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.90, size = 3) +

  #layout
    scale_y_continuous(limits = c(0, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, new reporters") +
  ylab("relative Sox2::mCherry expression") +
  col_scale_CDS_groups


``` 

### Fig S6I - rel_mCh_mChpos_newCDS_UTR
```{r , fig_S6I_rel_mCh_mChpos_newCDS_UTR, fig.width=8, fig.height = 4}
tib_for_plot = filter(expression_tib_comb, type %in% c("clone")  & experiment == "E2742" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-"  #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>% #leave UTR for supplement
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") 
 
ref_dash_lines = tib_for_plot %>%
  filter(insert_group == "minimal insert") %>%
  group_by(landing_pad) %>%
  summarize(ref_rel_mCh = mean(rel_mCherry))

#plotting
ggplot(tib_for_plot, aes(x = insert_name, y = rel_mCherry, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 2) +
  geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_mCh), linetype = 'dashed')+
  # geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
     ref.group = "minimal insert",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.95, size = 3) +

  #layout
  scale_y_continuous(limits = c(0, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry, new reporters") +
  ylab("relative Sox2::mCherry expression") +
  col_scale_CDS_groups


``` 

## absolute expression values
still need to compare between experiments, so: subtract autofluorescence and divide by 23_34A

```{r}
standard_tib_raw = expression_tib_comb %>%
  group_by(date) %>%
  filter(name == "23_34A") %>%
  summarize(mTurq_raw_standard = mean(median_mTurq_cor),
           mCherry_raw_standard = mean(median_mCherry_cor),
            GFP_raw_standard = mean(median_GFP_cor) )

expression_tib_comb_notnorm = expression_tib_comb %>%
  left_join(standard_tib_raw) %>%
  mutate(rel_mTurq_raw = median_mTurq_cor / mTurq_raw_standard,
         rel_mCherry_raw = median_mCherry_cor / mCherry_raw_standard,
         rel_GFP_raw = median_GFP_cor / GFP_raw_standard)
```

calculate fold change
```{r raw_fold_changes_oldCDS}

#calculate fold change mTurq Sox2P vs CDS
median_expr_values_tib %>%
  select(-median_rel_mCherry) %>%
  pivot_wider(names_from = "insert_name", values_from = "median_rel_mTurq")%>%
  mutate(fc_Sox2P_CDS = Sox2P_CDS_mTurq / Sox2P_mTurq,
         fc_Sox2P_CDS_UTR = Sox2P_CDS_mTurq_UTR / Sox2P_mTurq)
  
median_expr_values_tib_notnorm = filter(expression_tib_comb_notnorm, type == "clone"  &  experiment %in% c("E2410", "E2427") &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
 
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%

  #calculate median per condition
  group_by(landing_pad, insert_name) %>%
  summarize(median_rel_mTurq_raw = median(rel_mTurq_raw),
            median_rel_mCherry_raw = median(rel_mCherry_raw))


#calculate fold change mTurq Sox2P vs CDS
median_expr_values_tib_notnorm %>%
  select(-median_rel_mCherry_raw) %>%
  pivot_wider(names_from = "insert_name", values_from = "median_rel_mTurq_raw")%>%
  mutate(fc_Sox2P_CDS = Sox2P_CDS_mTurq / Sox2P_mTurq,
         fc_Sox2P_CDS_UTR = Sox2P_CDS_mTurq_UTR / Sox2P_mTurq)

``` 

```{r raw_fold_changes_newCDS}

#calculate fold change mTurq Sox2P vs CDS
median_expr_values_tib %>%
  select(-median_rel_mCherry) %>%
  pivot_wider(names_from = "insert_name", values_from = "median_rel_mTurq")%>%
  mutate(fc_Sox2P_CDS = Sox2P_CDS_mTurq / Sox2P_mTurq,
         fc_Sox2P_CDS_UTR = Sox2P_CDS_mTurq_UTR / Sox2P_mTurq)
  
median_expr_values_tib_notnorm = filter(expression_tib_comb_notnorm, type == "clone"  &  experiment %in% c("E2742") &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
 
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") %>%

  #calculate median per condition
  group_by(landing_pad, insert_name) %>%
  summarize(median_rel_mTurq_raw = median(rel_mTurq_raw),
            median_rel_mCherry_raw = median(rel_mCherry_raw))


#calculate fold change mTurq Sox2P vs CDS
median_expr_values_tib_notnorm %>%
  mutate(Sox2P_ref = pull(filter(median_expr_values_tib_notnorm , insert_name == "Sox2P_mTurq"), median_rel_mTurq_raw ),
         fc_to_Sox2P = median_rel_mTurq_raw/Sox2P_ref) %>%
  select(-c(median_rel_mCherry_raw, Sox2P_ref))

``` 

### Fig S7A - mCh+ clones, old
```{r, fig_S7A_abs_mChpos_old, fig.width=8, fig.height = 4}

tib_for_plot = filter(expression_tib_comb_notnorm, type == "clone"  & experiment %in% c("E2410", "E2427") & 
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good")

ref_dash_lines = tib_for_plot %>%
  filter(insert_group == "minimal insert") %>%
  group_by(landing_pad) %>%
  summarize(ref_rel_mCh = mean(rel_mCherry_raw),
            ref_rel_GFP = mean(rel_GFP_raw))


#plotting mTurq
plot_mTurq = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mTurq_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
    geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +

  #stats
  stat_compare_means(
    comparisons = list(c("Sox2P_mTurq", "Sox2P_CDS_mTurq"), c("Sox2P_mTurq", "Sox2P_CDS_mTurq_UTR"),
                                        c("Sox2P_CDS_mTurq", "Sox2P_CDS_mTurq_UTR")),
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
    tip.length = 0.02) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq") +
  ylab("relative expression, not normalized") +
  col_scale_CDS_named


#plotting mCherry
plot_mCherry = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mCherry_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
   geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_mCh), linetype = 'dashed')+
  geom_hline(yintercept = 0) + 
   geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +

  #stats
  stat_compare_means(
    comparisons = list( c("minimal insert", "Sox2P_mTurq"), c("minimal insert", "Sox2P_CDS_mTurq"), c("minimal insert", "Sox2P_CDS_mTurq_UTR"),
                        c("Sox2P_CDS_mTurq", "Sox2P_CDS_mTurq_UTR")),
     label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
    tip.length = 0.02) +

  #layout
  scale_y_continuous(limits = c(0, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry") +
  ylab("relative expression") +
  col_scale_CDS_named

#plotting GFP
plot_GFP = ggplot(tib_for_plot, aes(x = insert_name, y = rel_GFP_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_GFP), linetype = 'dashed')+
  geom_hline(yintercept = 0) + 
   geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
 
  #stats
  stat_compare_means(
    comparisons = list( c("minimal insert", "Sox2P_mTurq"), c("minimal insert", "Sox2P_CDS_mTurq"), c("minimal insert", "Sox2P_CDS_mTurq_UTR"),
                        c("Sox2P_CDS_mTurq", "Sox2P_CDS_mTurq_UTR")),
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
    tip.length = 0.02) +

  #layout
  scale_y_continuous(limits = c(0, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("GFP") +
  ylab("relative expression") +
  col_scale_CDS_named

cowplot::plot_grid(plotlist = list(plot_mTurq, plot_mCherry, plot_GFP), 
                   align = 'h', axis = c('tb'), nrow = 1)

``` 

### Fig S7B - mCh+ clones, new
```{r, fig_S7B_abs_mChpos_new, fig.width=8, fig.height = 4}

tib_for_plot = filter(expression_tib_comb_notnorm, type == "clone"  & experiment == "E2742" & 
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good")

ref_dash_lines = tib_for_plot %>%
  filter(insert_group == "minimal insert") %>%
  group_by(landing_pad) %>%
  summarize(ref_rel_mCh = mean(rel_mCherry_raw),
            ref_rel_GFP = mean(rel_GFP_raw))


#plotting mTurq
plot_mTurq = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mTurq_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
    geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +

  #stats
  stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq") +
  ylab("relative expression, not normalized") +
  col_scale_CDS_named


#plotting mCherry
plot_mCherry = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mCherry_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
   geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_mCh), linetype = 'dashed')+
  geom_hline(yintercept = 0) + 
   geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +

  #stats
  stat_compare_means(
     ref.group = "minimal insert",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  scale_y_continuous(limits = c(0,2.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry") +
  ylab("relative expression") +
  col_scale_CDS_named

library(scales)
#plotting GFP
plot_GFP = ggplot(tib_for_plot, aes(x = insert_name, y = rel_GFP_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_GFP), linetype = 'dashed')+
  geom_hline(yintercept = 0) + 
   geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
 
  #stats
  stat_compare_means(
     ref.group = "minimal insert",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +


  #layout
  scale_y_continuous(limits = c(0, 2.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("GFP") +
  ylab("relative expression") +
  col_scale_CDS_named

cowplot::plot_grid(plotlist = list(plot_mTurq, plot_mCherry, plot_GFP), 
                   align = 'h', axis = c('tb'), nrow = 1)

``` 

```{r, fig_5supE_abs_mChpos_new_wobars, fig.width=8, fig.height = 4}
tib_for_plot = filter(expression_tib_comb_notnorm, type == "clone"  & experiment == "E2742" & 
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state != "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good")

ref_dash_lines = tib_for_plot %>%
  filter(insert_group == "minimal insert") %>%
  group_by(landing_pad) %>%
  summarize(ref_rel_mCh = mean(rel_mCherry_raw),
            ref_rel_GFP = mean(rel_GFP_raw))


#plotting mTurq
plot_mTurq = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mTurq_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
    geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +

  #stats
  stat_compare_means(
        # comparisons = list(c("Sox2P_mTurq", "Sox2P_CDS_mTurq"), c("Sox2P_mTurq", "Sox2P_mTurq_CDS"),
        #                c("Sox2P_mTurq", "Sox2P_mTurq_stop_CDS"), c("Sox2P_mTurq", "Sox2P_mTurq_PAS_CDS"),
        #                                 c("Sox2P_mTurq", "Sox2P_mTurq_revCDS"),
        #                 c("Sox2P_mTurq", "Sox2P_CDS_mTurq_UTR"), c("Sox2P_mTurq", "Sox2P_mTurq_CDS_UTR"),
        #                c("Sox2P_mTurq", "Sox2P_mTurq_stop_CDS_UTR"), c("Sox2P_mTurq", "Sox2P_mTurq_PAS_CDS_UTR")),
    ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq") +
  ylab("relative expression, not normalized") +
  col_scale_CDS_named


#plotting mCherry
plot_mCherry = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mCherry_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
   geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_mCh), linetype = 'dashed')+
  geom_hline(yintercept = 0) + 
   geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +

  #stats
  stat_compare_means(
       # comparisons = list(c("minimal insert", "Sox2P_mTurq"), c("minimal insert", "Sox2P_CDS_mTurq"), c("minimal insert", "Sox2P_mTurq_CDS"),
       #                 c("minimal insert", "Sox2P_mTurq_stop_CDS"), c("minimal insert", "Sox2P_mTurq_PAS_CDS"),
       #                                  c("minimal insert", "Sox2P_mTurq_revCDS"),
       #                  c("minimal insert", "Sox2P_CDS_mTurq_UTR"), c("minimal insert", "Sox2P_mTurq_CDS_UTR"),
       #                 c("minimal insert", "Sox2P_mTurq_stop_CDS_UTR"), c("minimal insert", "Sox2P_mTurq_PAS_CDS_UTR")),
    ref.group = "minimal insert",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
    angle = 90, label.y.npc = 0.8, size = 3) +
  
  #layout
  scale_y_continuous(limits = c(0,2.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry") +
  ylab("relative expression") +
  col_scale_CDS_named

library(scales)
#plotting GFP
plot_GFP = ggplot(tib_for_plot, aes(x = insert_name, y = rel_GFP_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_hline(data = ref_dash_lines, aes(yintercept = ref_rel_GFP), linetype = 'dashed')+
  geom_hline(yintercept = 0) + 
   geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
 
  #stats
  stat_compare_means(
       # comparisons = list(c("minimal insert", "Sox2P_mTurq"), c("minimal insert", "Sox2P_CDS_mTurq"), c("minimal insert", "Sox2P_mTurq_CDS"),
       #                 c("minimal insert", "Sox2P_mTurq_stop_CDS"), c("minimal insert", "Sox2P_mTurq_PAS_CDS"),
       #                                  c("minimal insert", "Sox2P_mTurq_revCDS"),
       #                  c("minimal insert", "Sox2P_CDS_mTurq_UTR"), c("minimal insert", "Sox2P_mTurq_CDS_UTR"),
       #                 c("minimal insert", "Sox2P_mTurq_stop_CDS_UTR"), c("minimal insert", "Sox2P_mTurq_PAS_CDS_UTR")),
    ref.group = "minimal insert",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
    angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  scale_y_continuous(limits = c(0, 2.5)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("GFP") +
  ylab("relative expression") +
  col_scale_CDS_named

cowplot::plot_grid(plotlist = list(plot_mTurq, plot_mCherry, plot_GFP), 
                   align = 'h', axis = c('tb'), nrow = 1)

``` 


### Fig S7C - mCh- clones, old
```{r fig_S7C_abs_mChneg_old, fig.width=8, fig.height = 4}

tib_for_plot = filter(expression_tib_comb_notnorm, type == "clone"  & 
         experiment == "E2427" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state == "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") 

#plotting mTurq
plot_mTurq = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mTurq_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
    stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq") +
  ylab("relative expression, not normalized") +
  col_scale_CDS_groups


#plotting mCherry
plot_mCherry = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mCherry_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  scale_y_continuous(limits = c(-0.01, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry") +
  ylab("relative expression") +
  col_scale_CDS_groups

#plotting GFP
plot_GFP = ggplot(tib_for_plot, aes(x = insert_name, y = rel_GFP_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
    stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  scale_y_continuous(limits = c(-0.01, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("GFP") +
  ylab("relative expression") +
  col_scale_CDS_groups

cowplot::plot_grid(plotlist = list(plot_mTurq, plot_mCherry, plot_GFP), 
                   align = 'h', axis = c('tb'), nrow = 1)

``` 

### Fig S7D - mCh- clones, new
```{r fig_S7D_abs_mChneg_new, fig.width=8, fig.height = 4}

tib_for_plot = filter(expression_tib_comb_notnorm, type == "clone"  & 
         experiment == "E2759" &
         total_cell_count >= 500 &
         GFP_state != "GFP-" & #exclude WT (scaling to 0 GFP is meaningless)
         mCh_state == "mCh-" #C9_mCh- and 23_mCh- are not relevant for this comparison
         ) %>%
  
  #genotyping
  filter(is.na(insert_genotyped) | insert_genotyped == "good") 

#plotting mTurq
plot_mTurq = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mTurq_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
    stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mTurq") +
  ylab("relative expression, not normalized") +
  col_scale_CDS_groups


#plotting mCherry
plot_mCherry = ggplot(tib_for_plot, aes(x = insert_name, y = rel_mCherry_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
  stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  scale_y_continuous(limits = c(-0.025, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("mCherry") +
  ylab("relative expression") +
  col_scale_CDS_groups

#plotting GFP
plot_GFP = ggplot(tib_for_plot, aes(x = insert_name, y = rel_GFP_raw, col = insert_group)) +
  facet_nested(. ~ landing_pad , scales = "free_x", space = "free_x") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge(jitter.width = 1.8), size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed") + 
  geom_hline(yintercept = 0) + 
  
  #stats
    stat_compare_means(
     ref.group = "Sox2P_mTurq",
    label = "p.format", method = "t.test", method.args = list(var.equal = FALSE),
   angle = 90, label.y.npc = 0.8, size = 3) +

  #layout
  scale_y_continuous(limits = c(-0.025, 2)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "none",
        axis.title.x = element_blank()) +
  ggtitle("GFP") +
  ylab("relative expression") +
  col_scale_CDS_groups

cowplot::plot_grid(plotlist = list(plot_mTurq, plot_mCherry, plot_GFP), 
                   align = 'h', axis = c('tb'), nrow = 1)

``` 

# C9_mCh flipped clones
Show density plots for the clones we derived to use for the hopping
From E2447
## Pre-process data
```{r}
annotation_tib_E2447 = read_xlsx(path_annotation_E2447)

```

Load all FCS files
```{r load_set_of_fcs_files}
fcs_files = dir(path_fcs_files_E2447, full.names = T)
names(fcs_files) = str_remove(str_remove(str_remove(fcs_files, '.*/'),"_Single Cells2.fcs"), "export_")

matched_annotation_tib = tibble(filenames = names(fcs_files),
                                full_filenames = fcs_files,
                                # N_char = str_length(filenames),
                                well_tmp = str_extract(filenames, "...$"),
                                row = substr(well_tmp, 1, 1),
                                col = as.numeric(substr(well_tmp, 2, 3)),
                                well = paste0(row, col)) %>%
  select(-c(row, col, well_tmp)) %>%
  left_join(annotation_tib_E2447) %>%
  filter(experiment == "E2447" & insert_genotyped == "good" | experiment == "control") %>%
  mutate(insert_name = factor(expected_insert, 
                              levels = c( "none", "HyTK", "37", "34", "41", "40"),
                              labels = c(  "none","HyTK", "minimal insert", "Sox2P", "Sox2P_CDS", "Sox2P_CDS_UTR")))

matched_annotation_df = data.frame(matched_annotation_tib)
rownames(matched_annotation_df) = matched_annotation_df$filenames

flowset = flowCore::read.flowSet(files = matched_annotation_tib$full_filenames, alter.names = T, truncate_max_range = F, ignore.text.offset = T)

flowCore::sampleNames(flowset) = matched_annotation_tib$filenames
pData(flowset) = matched_annotation_df

flowset_fluo = flowset[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A')]
colnames(flowset_fluo) = c("GFP", "mCherry", "mTurq")

```


```{r preprocess_fcs_data}
total_cell_count_tmp = flowCore::keyword(flowset_fluo, "$TOT")[,'$TOT']
total_cell_count = as.numeric(total_cell_count_tmp)
names(total_cell_count) = names(total_cell_count_tmp)

median_mTurq = sapply(1:length(flowset_fluo), function(i){
  median(exprs(flowset_fluo[[i]])[,'mTurq'])
})
names(median_mTurq) = sampleNames(flowset_fluo)

median_mCherry = sapply(1:length(flowset_fluo), function(i){
  median(exprs(flowset_fluo[[i]])[,'mCherry'])
})
names(median_mCherry) = sampleNames(flowset_fluo)


median_GFP = sapply(1:length(flowset_fluo), function(i){
  median(exprs(flowset_fluo[[i]])[,'GFP'])
})
names(median_GFP) = sampleNames(flowset_fluo)

pData(flowset_fluo)$median_mTurq = median_mTurq
pData(flowset_fluo)$median_mCherry = median_mCherry
pData(flowset_fluo)$median_GFP = median_GFP

pData(flowset_fluo)$total_cell_count = total_cell_count

```

## Fig S6C - plot density plot
Sorted by mTurq level
```{r}
#set up the transformations I use in the plots
fwd_transf  = flowWorkspace::flowjo_biexp(widthBasis = -100, pos = 4.42, inverse = F)
inv_transf = flowWorkspace::flowjo_biexp(widthBasis = -100, pos = 4.42, inverse = T)

#function to draw a line at the two top peaks
#important: the values for the geom_density_ridges quantile lines need to be sorted non-decreasingly!
fun_high_peaks2 = function(x, ...){
  peaks_obj = getPeaks(fwd_transf(x))
  N_peaks_to_pick = min(2, length(peaks_obj$Peaks))
  # N_peaks_to_pick = 1
  top_2_peaks = order(peaks_obj$P.h, decreasing = T)[1:N_peaks_to_pick] 
  linear_peak = inv_transf(getPeaks(fwd_transf(x))$Peaks[top_2_peaks])
  sort(linear_peak)
}

```

```{r Fig_S6C_mChpos_41clones,  fig.width= 4, fig.height = 5}
sel_idx = pData(flowset_fluo)$total_cell_count >= 1000 & 
  !(pData(flowset_fluo)$insert_name == "Sox2P")

ggplot(flowset_fluo[sel_idx], aes(x = mTurq, fill = insert_name)) + 
  facet_grid(factor(insert_name, levels = c("Sox2P", "Sox2P_CDS", "Sox2P_CDS_UTR", "none"))
             ~ . , scales = 'free_y', space = 'free_y') +
  geom_density_ridges(aes(y = fct_reorder(name, median_mTurq)),
                      size =0.2, # rel_min_height  = 0.005,
                      scale = 2,
                      alpha = 0.75,
                      quantile_lines = TRUE, quantile_fun = fun_high_peaks2, width_vline = 2) + #, 
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw() +
  theme(legend.position = 'none',
        axis.title.y = element_blank()) +
  col_scale_CDS_named
```

# Fig S6L - Density plots -116 kb mCh- flipped clones
```{r fig_S6L_dens_mChneg, fig.width= 5, fig.height = 5}

samples_23_mChneg_names = 
  filter(fcs_tib, (type == "clone"  & mCh_state == "mCh-" & landing_pad == "-116 kb") | (type == "control" & landing_pad == "no insert")) %>%
  filter(date == "CM20240206") %>%
  filter(insert_genotyped == "good" | is.na(insert_genotyped)) %>%
  pull(filenames)



ggplot(flowset_fluo_comb[samples_23_mChneg_names], aes(x = mTurq, fill = insert_group)) + 
  facet_grid(factor(insert_name, levels = c("Sox2P", "Sox2P_CDS", "Sox2P_CDS_UTR", "none"))
             ~ . , scales = 'free_y', space = 'free_y') +
  geom_density_ridges(aes(y = fct_reorder(name, median_mTurq)),
                      size =0.2, # rel_min_height  = 0.005,
                      scale = 2,
                      alpha = 0.75,
                      quantile_lines = TRUE, quantile_fun = fun_high_peaks2, width_vline = 2) + #, 
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme_bw() +
  theme(legend.position = 'none',
        axis.title.y = element_blank()) +
  col_scale_CDS_groups

```



# Hopping experiment
## Load data
```{r}

#load the -161kb mCh+ Sox2P_CDS data (C9_mCh_41)
tib_C9_41 = read_tsv(path_tib_C9_41_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  ungroup() %>%
  #mark hopped integrations
  mutate(hopped = start != 34598479)

```

combine and filter
```{r}
tib_filtered <- tib_C9_41  %>%
    #add variables used in the plotting (old names of cell lines etc)
  mutate(population = case_when(sample_type == "control pool" ~ "ctrl",
                                .default = sorted_gate),
         cell_line = case_when(launch_pad_location == "-161 kb" & insert == "Sox2P" & genotype == "WT" ~ "C9_mCh_34",
                               launch_pad_location == "-161 kb" & insert == "Sox2P_CDS" & genotype == "WT" ~ "C9_mCh_41"
                               ),
         landing_pad = case_when(launch_pad_location == "-116 kb" ~ "23",
                                 launch_pad_location == "-161 kb" ~ "C9")) %>%
  #annotate confidence of mapping (one or both ITRs)
  mutate(mapped_arms = case_when(read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only",
                                 read_count_1 > 0 & read_count_2 > 0 ~ "both_arms")) %>%

  #filter the data
  filter(population == "ctrl" | read_count >= 2) %>% #for ctrl distribution 1 mapping read is enough
  filter(!(start >= 34721183 & start <= 34721192)) %>% #remove contamination from LP8
  filter(! start %in% c(25018987, 35232946) ) %>% #remove contamination clone C1
  filter(landing_pad  == "C9" | start != 34598479) %>% #remove contamination from LP C9
  filter(landing_pad  == "23" | start != 34643961) %>% #remove contamination from LP 23
  filter(landing_pad  == "C10" | start != 35478413) %>% #remove contamination from LP C10
  filter(strand %in% c("+", "-"))


tib_filtered_P1_strict = tib_filtered %>%
  filter(!(population == "P1" & mapped_arms != "both_arms")) #only keep P1 integrations with 2 mapped arms

all_exp = unique(tib_filtered_P1_strict$experiment)
```


### Add expression data
For C9_mCh+_41 we just use the FACS data from that cell line
```{r}
#load median expression values (from sorting pdfs)
median_sorted_mTurq = read_tsv(path_median_mTurq_relevant_cell_lines)

#add mTurq values to the integration tibble
tib_filtered_mTurq_vals = left_join(tib_filtered_P1_strict, median_sorted_mTurq) %>%
  filter(population != "ctrl") %>%
   dplyr::rename(mTurq_expr_gate = mTurq) 

#check that all integrations have a fluorescence value:
tib_filtered_mTurq_vals %>% filter(is.na(mTurq_expr_gate))
```

### Sorting statistics
For the new expression score we need to know, per population and replicate:
1. number of cells sorted
2. fraction of the cells in that gate (take fraction of the mChpos/mChneg total cells, whichever is relevant for that cell_line)

We add a psuedocount of 1 to all the 'recorded_count' when calculating the 'recorded_fraction', to deal with the 0 recorded cells in P1. 
Got the sorting stats from the sorting pdfs & the labguru notes of the experiments

```{r}
sorting_stats_tib_raw = read_xlsx(path_sorting_stats) 

sorting_stats_tib = sorting_stats_tib_raw %>%
  mutate(recorded_fraction = (recorded_count + 1)/total_recorded) %>% #recorded fraction with psuedocount of 1
  mutate(weigh_factor = recorded_fraction / sorted_count, 
         weigh_factor_1M = weigh_factor * 1E6) # weigh factor 1M is just to make the factor more easily human readable
```
For C9_mCh+_41 all populations have exactly 2 replicates, so no need to adjust for replicate count


## Fig 5C - Plot beeswarm
```{r}
plot_beeswarm_orange = function(TIB, CL = "C9_mCh_41", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = F, PLOT_ANNOT = T,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom",
                              NEXT_RANGE = NULL){
  tib_for_plot = filter(TIB, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP & start >= P_RANGE[1] & start <= P_RANGE[2])
  LP_tib_cl = select(tib_for_plot, c(cell_line, landing_pad)) %>%
    distinct() %>% left_join(LP_tib)

  ggplot(filter(tib_for_plot, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP),
         aes(x = start, y = population, col = population)) +
    {if(length(CL)>1) facet_grid(cell_line ~ .)} +  #only facet when necessary

    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    {if(PLOT_ANNOT & length(CL)== 1)
      geom_vline(xintercept = pull(
        filter(LP_tib, landing_pad %in% unique(tib_for_plot$landing_pad)),
        start),
        col = 'darkgrey')} +
    {if(PLOT_ANNOT & length(CL) > 1)
       geom_vline(data = LP_tib_cl, aes(xintercept = start), col = 'darkgrey')} +

    # # annotate CTCFs
    {if(PLOT_CTCF)geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty="11")}+
    {if(!is.null(NEXT_RANGE)) geom_rug(data = tibble(range = NEXT_RANGE),
                                       aes(x = range),
                                       fill = 'grey',
                                       inherit.aes = F)}+
     #annotate landing pad
    labs(x='Genomic coordinate (Mb)',y='sorted gate') +

    #datapoints
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = T, bandwidth = 0.8,
                     # shape = 16, #changes to point without border, so alpha applies to the whole point
                     method = RANDOM_METHOD,
                     stroke = NA) +
    geom_text(data = plyr::count(tib_for_plot, vars = c("cell_line", "population")), aes(label = paste("n = ", freq, sep = ""),
                                                                                         x = Inf),
              hjust = "inward", vjust = "top",
              col = 'black') +
    #layout:
    theme_classic(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                       limits=P_RANGE, expand=c(0,0)) +
    scale_y_discrete(limits = rev, drop = F) +
    ggtitle(TITLE) +
    colScale7_orange
}
```

```{r Fig_5C_beeswarm_only_prange, fig.width = 6, fig.height = 3}
tib_for_plot = filter(tib_filtered_P1_strict, hopped & chr == "chr3" &  experiment %in% all_exp ) %>%
  mutate(population = factor(population, levels = c("ctrl", paste0("P", 1:6))))

C9_41_beesw = plot_beeswarm_orange(TIB = tib_for_plot,
                   CL = c("C9_mCh_41"),
                   PLOT_ANNOT = T,
                   PLOT_CTCF = T,
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1.5
)
C9_41_beesw
```

## Fig 5D - Expression score
```{r}
#C9_mCh+_34 5kb window
set.seed(20240327)
min_ints_window = 3
N_bootstrap = 5000
BinSize = 500
CL = "C9_mCh_41"
all_exp = unique(tib_filtered_mTurq_vals$experiment)
tib_filtered_mTurq_vals = tib_filtered_mTurq_vals %>% left_join(sorting_stats_tib)


# # calculate running mean on random sample of same size, 500 times
# expr_score_tib = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 10,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = CL, EXP = all_exp)},
#                                             TIB = tib_filtered_mTurq_vals,
#                                             N_resamples = N_bootstrap)
# saveRDS(expr_score_tib, "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_C9_mCh_41_5kb.rds")

expr_score_tib = readRDS("/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_C9_mCh_41_5kb.rds")



expr_score_sum_tib = bootstrap_summary_fun(expr_score_tib,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real =
  expr_score_function_new_weigh(tib_filtered_mTurq_vals,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = CL,
                                EXP = all_exp)
  
expr_score_sum_tib_filt_C9_41 = left_join(expr_score_sum_tib,  expr_score_real) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA)) %>%
  mutate(cell_line = CL)
  

```

```{r}
#load the bootstrapped data from C9_mCh_34 (generated in fig2)
expr_score_sum_tib_filt_C9_34 = readRDS(path_expr_score_bootstrap_C9_mCh_34)

expr_sum_tib_filt_comb = bind_rows(expr_score_sum_tib_filt_C9_34, expr_score_sum_tib_filt_C9_41)
```


###  Fig 5D - Plot
```{r Fig_5D_expr_plot_34_41, fig.width=6, fig.height = 4}

expr_plot_C9 = ggplot(expr_sum_tib_filt_comb, aes(x = mid_interval)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_C9), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel,
                  fill = cell_line), alpha = 0.3) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel, col = cell_line)) +

  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none') +
  col_scale_CDS_cl
expr_plot_C9

```

### Fig 5CD - Plot
```{r Fig_5CD_beesw_with_expr_plot_34_41, fig.width=6, fig.height = 6}
cowplot::plot_grid(plotlist = list(C9_41_beesw, expr_plot_C9),
                   align = 'v', axis = c('lr'), ncol = 1)

```

# FACS data sorter
## load FACSDiva file
```{r}
library(openCyto)
library(CytoML)
library(flowWorkspace)


diva_ws <- open_diva_xml(diva_xml_file)


gs <- diva_to_gatingset(diva_ws,
                        name = 1, #the group to import
                        path = path_FACS_sorting_E2555,
                        #swapped columns are a known diva 'bug', for some files you need to 'unswap' them, default is T (as necessary for this file)
                        # swap_cols = F,
                        worksheet = "global",
                        verbose = F,
                        execute = T)

pdata_tib = pData(gs) %>%
  as_tibble() %>%
  mutate(name_short = str_remove(str_remove(name, "ME2024050._E2534_rep._"), "_([0-9]){3}.fcs"),
         replicate = str_extract(name_short, "rep.$"),
         cell_line = case_when(grepl("WT", name_short) ~ "WT",
                               grepl("F121_9", name_short) ~ "F121_9",
                               grepl("23_34A", name_short) ~ "23_34A",
                               .default = str_extract(name_short, "C9_.._mCh.")),

         treatment = case_when(grepl("PB", name_short) ~ "PB",
                               grepl("SB", name_short) ~ "SB",
                               .default = "untreated"),
         recording = case_when(grepl("_SORT", name_short) ~ "sorting",
                               .default = "pre-recording"),
         sample_type = case_when(cell_line %in% c("WT", "F121_9", "23_34A") | treatment == "untreated" ~ "control",
                                 .default = "sample")
  ) %>%
  mutate(construct = case_when(cell_line %in% c("WT", "F121_9") ~ "none",
                                cell_line == c("23_34A") ~ "34",
                                .default = str_extract(cell_line, "([0-9]){2}")),
         mCh_status = case_when(cell_line == "WT" ~ "mCh-",
                                cell_line %in% c("23_34A", "F121_9") ~ "mCh+",
                                .default = str_extract(cell_line, "mCh.")))


pdata_df = as.data.frame(pdata_tib)
rownames(pdata_df) = pdata_df$name

pData(gs) = pdata_df
```

Load replicate 2:
```{r}
diva_ws2 <- open_diva_xml(diva_xml_file2)


gs2 <- diva_to_gatingset(diva_ws2,
                        name = 1, #the group to import
                        path = path_FACS_sorting_E2555,
                        swap_cols = F, #apparently we don't need to swap the FSC/SSC-H/W here
                        worksheet = "global",
                        verbose = F,
                        execute = T)

pdata_tib2 = pData(gs2) %>%
  as_tibble() %>%
  mutate(name_short = str_remove(str_remove(name, "ME2024050._E2534_rep._"), "_([0-9]){3}.fcs"),
         replicate = str_extract(name_short, "rep.$"),
         cell_line = case_when(grepl("WT", name_short) ~ "WT",
                               grepl("F121_9", name_short) ~ "F121_9",
                               grepl("23_34A", name_short) ~ "23_34A",
                               .default = str_extract(name_short, "C9_.._mCh.")),

         treatment = case_when(grepl("PB", name_short) ~ "PB",
                               grepl("SB", name_short) ~ "SB",
                               .default = "untreated"),
         recording = case_when(grepl("_SORT", name_short) ~ "sorting",
                               .default = "pre-recording"),
         sample_type = case_when(cell_line %in% c("WT", "F121_9", "23_34A") | treatment == "untreated" ~ "control",
                                 .default = "sample")
  ) %>%
  mutate(construct = case_when(cell_line %in% c("WT", "F121_9") ~ "none",
                                cell_line == c("23_34A") ~ "34",
                                .default = str_extract(cell_line, "([0-9]){2}")),
         mCh_status = case_when(cell_line == "WT" ~ "mCh-",
                                cell_line %in% c("23_34A", "F121_9") ~ "mCh+",
                                .default = str_extract(cell_line, "mCh.")))


pdata_df2 = as.data.frame(pdata_tib2)
rownames(pdata_df2) = pdata_df2$name

pData(gs2) = pdata_df2
```

## Fig S6F - mCh+ cell lines
```{r, sup_Fig_5F_plot_FACS_41, height = 3, width = 6}

gs_oi_mChpos = subset(gs, sample_type == "sample"  & mCh_status == "mCh+" & construct == "41" &
                        (recording == "sorting" | treatment == "PB")) #sorting only, combining 2 data sets doesn't give the right percentages on the gates

ggcyto::ggcyto(gs_oi_mChpos, aes(x = "GFP", y = "mTurqoise2"), subset = "mCh_pos") +
  #facet by cell line / treatment
  facet_grid(. ~ treatment) +

  #plot cells
  geom_hex(bins = 128, aes( fill = after_stat(ncount))) +
  scale_fill_distiller(palette = 'Spectral') + #manually adding a fill scale also forces the scale to be linear again.

  #plot_gates
  ggcyto::geom_gate(paste0("P", 1:6, "_mCh+"), col = 'black') +
  ggcyto::geom_stats(adjust = c(-0.3, 0.5), digits = 2, size = 5) +

  #layout
  theme_bw(base_size = 14) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.position = "none",
        aspect.ratio = 1) +
  ggtitle(element_blank())+
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  ggcyto::ggcyto_par_set(limits = list(x = c(1.2, 4), y = c(0, 4.2))) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0))+
  ggcyto::labs_cyto(labels = "marker")

```

## Normalization
There is a default function to get the median fluorescent intensity (MFI) per fluorophore per gate:

```{r}

#here I get the single cells, that gate is called 'P4', don't confuse with the mTurq expression gate P4 (which is 'P4_mCh+' or 'P4_mCh-')
ctrl_samples_MFI = gs_pop_get_stats(gs, paste0("P4"), type = pop.MFI, inverse.transform = T)
ctrl_samples_MFI_tib1 = as_tibble(ctrl_samples_MFI) %>%
  left_join(pData(gs), by = join_by(sample == name)) %>%
  filter(treatment == "untreated" &
           !cell_line %in% c("C9_34_mCh-", "C9_41_mCh-"))%>%
  mutate(rel_mCh = mCherry/GFP,
         rel_mT = mTurqoise2/GFP,
         replicate = "rep1",
         pop = "single_cells")

ctrl_samples_MFI2 = gs_pop_get_stats(gs2, paste0("P4"), type = pop.MFI, inverse.transform = T)
ctrl_samples_MFI_tib2 = as_tibble(ctrl_samples_MFI2) %>%
  left_join(pData(gs2), by = join_by(sample == name)) %>%
  filter(treatment == "untreated" &
           !cell_line %in% c("C9_34_mCh-", "C9_41_mCh-"))%>%
  mutate(rel_mCh = mCherry/GFP,
         rel_mT = mTurqoise2/GFP,
         replicate = "rep2",
         pop = "single_cells")

ctrl_samples_MFI_tib = bind_rows(ctrl_samples_MFI_tib1, ctrl_samples_MFI_tib2)

# normalization
MFI_WT_ctrl = ctrl_samples_MFI_tib %>%
  filter(cell_line == 'WT') %>%
  mutate(GFP_WT = GFP, mTurq_WT = mTurqoise2, mCh_WT = mCherry) %>%
  select(replicate, GFP_WT, mTurq_WT, mCh_WT)

MFI_ctrl_tib2 = ctrl_samples_MFI_tib %>%
  left_join(MFI_WT_ctrl) %>%
  mutate(GFP_cor = GFP - GFP_WT,
         mTurq_cor = mTurqoise2 - mTurq_WT,
         mCh_cor = mCherry - mCh_WT) %>%
  mutate(mTurq_norm = mTurq_cor / GFP_cor,
         mCh_norm = mCh_cor / GFP_cor)

MFI_23_34A = MFI_ctrl_tib2 %>%
  filter(cell_line == "23_34A") %>%
  mutate(mTurq_norm_ctrl = mTurq_norm,
         mCh_norm_ctrl = mCh_norm) %>%
  select(replicate, mTurq_norm_ctrl, mCh_norm_ctrl)

```

## Plot individual cells P2-P4
extract the individual data for P2-P4 (mCh+ sorting data)
```{r}
gs_oi_mChpos = subset(gs, sample_type == "sample"  & treatment == "SB"  & mCh_status == "mCh+")
gs_oi_mChpos2 = subset(gs2, sample_type == "sample"  & treatment == "SB" & mCh_status == "mCh+")

#replicate 1
exprs_P24_ls = lapply(1:2, function(cl){
  exprs_per_rep_ls = lapply(c("P2", "P3", "P4"), function(pop){
    gatename = paste0(pop, "_mCh+")
    data_pop = gh_pop_get_data(gs_oi_mChpos[[cl]], gatename, inverse.transform = T)
    as_tibble(flowCore::exprs(data_pop))
  })
  names(exprs_per_rep_ls) = c("P2", "P3", "P4")
  exprs_rep_tib = bind_rows(exprs_per_rep_ls, .id = "population") %>% as_tibble()
})
names(exprs_P24_ls) = c("C9_34_mCh+", "C9_41_mCh+")
exprs_P24_tib = bind_rows(exprs_P24_ls, .id = "cell_line") %>% as_tibble() %>% mutate(replicate = "rep1")

#replicate 2
exprs_P24_ls2 = lapply(1:2, function(cl){
  exprs_per_rep_ls = lapply(c("P2", "P3", "P4"), function(pop){
    gatename = paste0(pop, "_mCh+")
    data_pop = gh_pop_get_data(gs_oi_mChpos2[[cl]], gatename, inverse.transform = T)
    as_tibble(flowCore::exprs(data_pop))
  })
  names(exprs_per_rep_ls) = c("P2", "P3", "P4")
  exprs_rep_tib = bind_rows(exprs_per_rep_ls, .id = "population") %>% as_tibble()
})
names(exprs_P24_ls2) = c("C9_34_mCh+", "C9_41_mCh+")
exprs_P24_tib2 = bind_rows(exprs_P24_ls2, .id = "cell_line") %>% as_tibble() %>% mutate(replicate = "rep2")

# combine and normalize
exprs_P24_tib_comb = bind_rows(exprs_P24_tib, exprs_P24_tib2)

exprs_P24_tib_norm = exprs_P24_tib_comb %>%
  left_join(MFI_WT_ctrl) %>%
    mutate(GFP_cor = `BL[B] 530/30-A` - GFP_WT,
         mTurq_cor = `V[F] 450/50-A` - mTurq_WT,
         mCh_cor = `YG[D] 610/20-A` - mCh_WT) %>%
  mutate(mTurq_norm = mTurq_cor / GFP_cor,
         mCh_norm = mCh_cor / GFP_cor) %>%
  left_join(MFI_23_34A) %>%
  mutate(rel_mTurq_norm = mTurq_norm / mTurq_norm_ctrl,
         rel_mCh_norm = mCh_norm / mCh_norm_ctrl)
```

## Fig 5E - rel_mCh_hopping_cut_axis
```{r Fig_5E_rel_mCh_hopping_cut_axis, fig.width = 5, fig.height = 5}
#here I flip the axis to make it more interpretable, since we want to see the effect of mTurq (x) on mCherry (y)

ggplot(exprs_P24_tib_norm, aes(x = rel_mTurq_norm, y = rel_mCh_norm, col = cell_line, fill = cell_line)) +
  # facet_grid(replicate ~ cell_line) +
  geom_point(alpha = 0.5) +
  coord_cartesian(xlim = c(0, 30), ylim = c(0, 1.5), expand = F) +
  geom_smooth(method = "lm")+
  stat_cor(method = 'spearman',label.x.npc = 1, label.y.npc = 0, hjust = 1)+
  labs(x = "relative reporter expression (mTurq)", y = "relative Sox2::mCherry expression") +
  theme_classic(base_size = 14) +
  col_scale_CDS_cl_FACS +
  theme(legend.position = "bottom",
        aspect.ratio = 1)
```

## Fig S6G - rel_mCh_hopping_all_data
```{r sup_F_rel_mCh_hopping_all_data, fig.width = 5, fig.height = 5}

ggplot(exprs_P24_tib_norm, aes(x = rel_mTurq_norm, y = rel_mCh_norm, col = cell_line, fill = cell_line)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm")+
  stat_cor(method = 'spearman',label.x.npc = 1, label.y.npc = 1, hjust = 1)+
  labs(x = "relative reporter expression (mTurq)", y = "relative Sox2::mCherry expression") +
  theme_classic(base_size = 14) +
  col_scale_CDS_cl_FACS +
  theme(legend.position = "bottom",
        aspect.ratio = 1)
```


## plot SCR-Sox2 interaction
Code for generating Figure S6B can be found in the github of Mikhail Magnitov (https://github.com/magnitov/sox2_rcmc).


# RNA stability CDS constructs
We wanted to check whether the addition of the CDS changes the stability of the produced RNA.
For this we did RNA degradation timecourses, followed by qPCR.

```{r}
#load the raw cq values 
tib_qPCR_raw = read_xlsx(path_qPCR_RNA_stability )

#preprocessing
tib_qPCR_all = tib_qPCR_raw %>%
  #calculate mean of 3 technical replicates
  mutate(Cq_mean = rowMeans(select(., Cq_1, Cq_2, Cq_3), na.rm = TRUE)) %>%
  #fold change over t0
  group_by(Cell_line, target, rep) %>%
  mutate(delta_Cq = Cq_mean - Cq_mean[time == 0]) %>%
  ungroup() %>%
  #convert to linear scale
  mutate(abundance = 2^(-delta_Cq)) %>%
  #sd on the technical replicates
  mutate(Cq_sd = apply(select(., Cq_1, Cq_2, Cq_3), 1, sd, na.rm = TRUE))


```

## Fig S6D - RNA degradation curve
Calculate and plot SD of replicates for delta Ct, normalized to T0
```{r}
# Step 1: Compute mean and SD for Fold Change across replicates
plot_data_deltaCt <- tib_qPCR_all %>%
  group_by(Cell_line, target, time) %>%
  summarise(
    abundance_Mean = mean(abundance),
    abundance_Change_SD = sd(abundance),
    .groups = "drop"
  )
```

```{r Fig_S6D_RNA_degradation_curve, fig.width=5, fig.height=5}

plot_data_deltaCt %>%
  mutate(target_spec = case_when(target == "mTurq" ~ paste0(Cell_line, "_", target),
                                 .default = target)) %>%
  mutate(interesting_lines = case_when(target == "mTurq" ~ "interesting",
                                    .default = "control"),
         interesting_lines = factor(interesting_lines, levels =c("interesting", "control"))) %>%
  # mutate(line_thickness = case_when(target == "mTurq" ~ "thicker",
  #                                   .default = "normal")) %>%
  filter(target != "sox2") %>%
  dplyr::arrange(match(target, c("gfp", "mCherry", "actin", "gapdh", "mTurq")) ) %>%
  ggplot(aes(x = time, y = abundance_Mean, group = interaction(Cell_line, target_spec), 
             col = target_spec)) + #linetype = Cell_line, 
  geom_line(aes(linetype = interesting_lines, linewidth = interesting_lines))+
  # geom_line(aes(linewidth = line_thickness)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = abundance_Mean - abundance_Change_SD, 
                    ymax = abundance_Mean + abundance_Change_SD),  
                width = 0.2) +
  # facet_wrap(~Cell_line) +
  scale_y_log10() + # Log scale for better visualization
  labs(
    # title = "RNA abundance relative to T0",
    x = "Time (hours)",
    y = "relative RNA abundance (log scale)"
  ) +
  theme_bw() +
  theme(#legend.position = "none",
        aspect.ratio = 0.8) +
  scale_color_manual(
    values = c(
      "gfp" = "#6ed17c",
      "mCherry" = "#ff3fb1",
      # "gfp" = "#39B54A",
      # "mCherry" = "#EC008C",
      "SoxP_mTurq" = '#0077BB',
      "SoxP CDS_mTurq" = "#EE7733",
      "SoxP UTR_mTurq" = "#bd4d0e",
      "actin" = "lightgrey",
      "gapdh" = "darkgrey")
    ) +
    scale_linewidth_manual(values = c("interesting" = 1.2, "control" = 0.7))
```

## Fig S6E - RNA half live
### Fitting a non linear model (exponential decay) to data, calculate half-live
```{r}

tib_model = tib_qPCR_all %>% group_by(Cell_line, target, rep) %>%
  select(Cell_line, target, rep, time, abundance) %>%
  arrange(time)

# Fit the model for each group
fit_results <- tib_model %>%
  nest(data = c(time, abundance)) %>%
  mutate(
    model = map(data, ~nls(abundance ~ Y0 * exp(-k * time), 
                           data = .x, 
                           start = list(Y0 = 1, k = 0.1))),
    params = map(model, broom::tidy)  # Extract model parameters
  )

# Extract and view fitted parameters
params <- fit_results %>%
  unnest(params)

params


# Add half-life calculation
params <- params %>%
  filter(term == "k") %>%
  mutate(half_life = log(2) / estimate)

params
```

### plot
Make again, just with faceting
```{r S6E_half_lives_Sox2, fig.width=3, fig.height=3}

tib_half_live_mTurq = params %>% 
  #filter(target == "mTurq") %>% 
  select(Cell_line, half_life, target, rep) %>%
  group_by(Cell_line, target) %>% 
  summarise(mean_half_life = mean(half_life),
            half_live_sd = sd(half_life),
            .groups = "drop")


tib_for_plot = params %>% filter(target %in% c("gfp", "mCherry", "mTurq")) %>%
    mutate(target_spec = case_when(target == "mTurq" ~ paste0(Cell_line, "_", target),
                                 .default = target))

means_for_plot = tib_for_plot %>% group_by(Cell_line, target, target_spec) %>%
      summarise(mean_half_life = mean(half_life), .groups = "drop")
  
ggplot(tib_for_plot, aes(x = target, 
           y = half_life, 
           color = target_spec)) +
  facet_grid(.~Cell_line, scales = "free_x")+
  # Points for each (target, cell line) combination
  geom_jitter(width = 0.2, size = 1) +
  
    # Add horizontal line for mean
  geom_errorbar(data = means_for_plot,
   aes(ymin = mean_half_life, ymax = mean_half_life, x = target, col = target_spec),
    width = 0.7,
   size = 1,  # Thickness of the line
    inherit.aes = FALSE) +  # Prevents ggplot from using global aesthetics
 
  scale_y_continuous(limits = c(0, NA),
                     name = "Half-life (hours)") +
  theme_bw(base_size = 14) +

  # ggtitle("Comparing RNA Half-lives Across Cell Lines") +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        aspect.ratio = 1.5,
        legend.position = 'none') +
  scale_color_manual(
    values = c(
      "gfp" = "#39B54A",
      "mCherry" = "#EC008C",
      "SoxP_mTurq" = '#0077BB',
      "SoxP CDS_mTurq" = "#EE7733",
      "SoxP UTR_mTurq" = "#bd4d0e")
    ) 

```


# Fig S6J - ChIP on CDS
Explore whether there is evidence for TF binding on the CDS.
Steps:
1. Use the JASPAR called motifs (downloaded from the UCSC genome browser track) to find all motifs on the CDS
2. Overlay with cistrome ChIP datasets, to find the motifs that have a peak overlaying it.
3. Manually check the ChIP datasets, remove all sets that are wrongly annotated, not actually mESCs or perturbation conditions.
4. Plot 1 dataset per found TF.

## 1. JASPAR

Download the full list of nonredundant, vertebrate core JASPAR motifs from the JASPAR website (to get the TF names).
```{r}
jaspar_motifs_file = readLines(path_JASPAR_motifs) 
jaspar_motifs_names_lines = jaspar_motifs_file[grepl("^>", jaspar_motifs_file)]

jaspar_motifs_names = tibble(Motif_ID = str_remove(str_remove(jaspar_motifs_names_lines, "^>"), "\t.*$"),
                             Factor = toupper(str_remove(jaspar_motifs_names_lines, "^.*\t")))
```


Exported the JASPAR mapped motifs from the prange_Sox2_gene_gr (chr3 34630000-34680000) from the UCSC JASPAR track (and saved).
Filter for a score of 300, this corresponds to a maximum p-value of 10^-3
```{r}
motifs_on_roi = read_tsv(path_JASPAR_mapped_motifs_Sox2, col_names = c("seqnames", 'start','end', 'Motif_ID', 'score', 'strand'))

motifs_on_roi_filt = motifs_on_roi %>% 
  filter(score >= 300) %>%
  left_join(jaspar_motifs_names)

#split the double motifs, chip of either factor needs to be matched
double_motifs = motifs_on_roi_filt %>% 
  as_tibble() %>% 
  filter(grepl("::", Factor))%>%
  mutate(Factor1 = str_remove(Factor, "::.*$"),
         Factor2 = str_remove(Factor, "^.*::")) %>%
  dplyr::rename(full_factor = Factor) %>%
  pivot_longer(cols = c(Factor1, Factor2), names_to = 'halfmotif', values_to = 'Factor')

motifs_on_roi_filt_all = as_tibble(motifs_on_roi_filt)  %>% bind_rows(double_motifs)

```

Define regions
```{r}

Sox2_CDS_gr = GRanges('chr3:34650415-34651375')

```

extract motifs
```{r}
motifs_on_roi_filt_gene_all  = subsetByOverlaps(GRanges(motifs_on_roi_filt_all), Sox2_CDS_gr, ignore.strand = T)


```

## 2. Cistrome 

Load the overview table of all datasets
```{r}
cistrome_all_datasets_tib = read_tsv(path_cistrome_overview_table)

cistrome_mESC_tib = cistrome_all_datasets_tib %>% filter(Cell_type == "Embryonic Stem Cell")

```

Filter for the factors with a motif on the CDS
```{r}

cistrome_mESC_tfoi = cistrome_mESC_tib %>% filter(Factor %in% unique(motifs_on_roi_filt_gene_all$Factor))


N_motif_roi = length(unique(motifs_on_roi_filt_gene_all$Factor))
N_motif_roi
N_ChIP_roi = length(unique(cistrome_mESC_tfoi$Factor))
N_ChIP_roi
```


Overlap cistrome peaks with the motifs
```{r}

extraCols_narrowPeak <- c(singnalValue = "numeric", pValue = "numeric",
                            qValue = "numeric", peak = "integer")

peaks_on_motifs_ls = lapply(cistrome_mESC_tfoi$DCid, function(id){
  # print(id)
  #load the tf motifS
  tf = cistrome_mESC_tfoi$Factor[cistrome_mESC_tfoi$DCid == id]
  tf_motifs = motifs_on_roi_filt_gene_all[motifs_on_roi_filt_gene_all$Factor == tf,] %>% GRanges()
  
  #load the tf peaks
  file_tf = paste0(path_cistrome_peak_folder, id, "_sort_peaks.narrowPeak.bed")
  tf_narrowPeak <- import.bed(file_tf,
                              extraCols = extraCols_narrowPeak)
  
  #find overlaps
  overlaps = findOverlaps(tf_narrowPeak, tf_motifs, ignore.strand = T)
  
  #get the motifs that have overlaps, with annotation
  motif_oi_tib = as_tibble(tf_motifs[subjectHits(overlaps)]) %>%
    mutate(peak_hit_number = queryHits(overlaps),
           DCid = id)
  
  #combine with the peak information
  peak_oi_tib = as_tibble(tf_narrowPeak[queryHits(overlaps)]) %>%
    mutate(peak_hit_number = queryHits(overlaps)) %>%
    unique()%>%
    full_join(motif_oi_tib, join_by(peak_hit_number), suffix = c('.peak', '.motif'))
  
  peak_oi_tib
})

peaks_on_motifs = bind_rows(peaks_on_motifs_ls)


```




## 3. check and download chip

I manually check the datasets that give a hit.
Remove the following ones (DCid)

59471:59474,#not ZFP57 chip
48327, 48329, #not SPI1
55144, #not mESC
58293, 58294, #NRF1 overexpression

449, #SMAD2: pSmad2 ChIPSeq of ES cells treated for 18 hours in 10uM SB
33779, 33721, 33762, 33794, 33755, #Pou5f1 ChIp after Sox2 or Sox17 overexpression
34041, #embryoid bodies, not mESC
35374, # Dnmt3a-/-Dnmt3b-/-(DKO)
44552, #klf4 overexpression
48489, #GSM1355156 was deleted by the GEO staff on Feb 29, 2016. (rep2 is still on geo, but not showing up in our list) Manually checked, rep1 and 2 look completely different so I suspect there is something wrong with the rep1 (which showed a 'peak' in our data)
52799, #treatment:  1um Retinoic Acid (RA) for 48hrs (other RXRA sample is the ES sample, that one is okay)
53113, # ESC+SB to induce endoderm differentiation
53680, # c-Jun was induced by dox for 36h, using c-Jun antibody performed ChIP-seq
73465, #PARP-1KO (and the WT ES samples do not show up in our analysis)
76355, # mitotic arrest, the asynchronous samples do not show up in our analyiss
81307, 81310, 82115 #one of these samples it the input samples, the other the actual IP samples, they look the same and both have peaks, so no actual signal


Programmatically download the ChIP tracks

```{r}
peaks_oi = peaks_on_motifs %>% filter(end.motif >= 34650415 & start.motif <= 34652461)  %>%
  filter(!DCid %in% c(
    59471:59474,#not ZFP57 chip
    48327, 48329, #not SPI1
    55144, #not mESC
    58293, 58294, #NRF1 overexpression
    449, #SMAD2: pSmad2 ChIPSeq of ES cells treated for 18 hours in 10uM SB
    33779, 33721, 33762, 33794, 33755, #Pou5f1 ChIp after Sox2 or Sox17 overexpression
    34041, #embryoid bodies, not mESC
    35374, # Dnmt3a-/-Dnmt3b-/-(DKO)
    44552, #klf4 overexpression
    48489, #GSM1355156 was deleted by the GEO staff on Feb 29, 2016. (rep2 is still on geo, but not showing up in our list) Manually checked, rep1 and 2 look completely different so I suspect there is something wrong with the rep1 (which showed a 'peak' in our data)
    52799, #treatment:  1um Retinoic Acid (RA) for 48hrs (other RXRA sample is the ES sample, that one is okay)
    53113, # ESC+SB to induce endoderm differentiation
    53680, # c-Jun was induced by dox for 36h, using c-Jun antibody performed ChIP-seq
    73465, #PARP-1KO (and the WT ES samples do not show up in our analysis)
    76355, # mitotic arrest, the asynchronous samples do not show up in our analyiss
    81307, 81310, 82115 #one of these samples it the input samples, the other the actual IP samples, they look the same and both have peaks, so no actual signal
  )) 
files_oi = peaks_oi %>% select(DCid, Factor) %>% unique()

DCid_to_download = files_oi$DCid
TF_name_DCid = files_oi$Factor


# lapply(1:length(DCid_to_download), function(i){
#   url_cistrome = paste0('http://dc2.cistrome.org//genome_browser/bw/', DCid_to_download[i],'_treat.bw')
#   folder_to_make = paste0(path_cistrome_bw_folder, TF_name_DCid[i])
#   filename_to_make = paste0(folder_to_make, "/", TF_name_DCid[i], "_", DCid_to_download[i], ".bw")
# 
#   dir.create(folder_to_make, showWarnings = FALSE)
#   if (!file.exists(filename_to_make)) {
#     download.file(url_cistrome, filename_to_make)
#   }
# })

cistrome_mESC_tfoi %>% filter(DCid %in% peaks_oi$DCid)


```

## 4. Plot filtered 
We want to plot only 1 track per TF. When there are 2 from the same experiment, just use rep1
For Pou5F1 (oct4) there are two different datasets. I take 52773  (the most recent dataset), which is completely untreated samples whereas the other one is a control shLuc transfection.

```{r}
data_oi_filt = peaks_oi %>% 
  filter(end.motif >= 34650415 & start.motif <= 34651375)  %>%
  select(DCid, Factor) %>% unique() %>%
  filter(!DCid %in% c(51365, 53133, #only keep rep1
                      38632 #only keep 1 Pou5f1 experiment
                      ))

tf_file_filtered_ls = lapply(1:length(data_oi_filt$DCid), function(i){
  paste0(path_cistrome_bw_folder, data_oi_filt$Factor[i], "/", data_oi_filt$Factor[i], "_", data_oi_filt$DCid[i], ".bw")
})
names(tf_file_filtered_ls) = paste0(1:length(data_oi_filt$DCid), "_", data_oi_filt$Factor)


```

```{r fig.width = 6, fig.height = 5}

Sox2_gene_annot = tibble(txStart =34649994 , 
                         txEnd = 34652461 ,
                         cdsStart= 34650415,
                         cdsEnd = 34651375)




plot_limits_ChIP =  c(34646500, 34656000)

plot_ls = lapply(order(data_oi_filt$Factor), #this takes the Factors in alphabetical order
                 function(i) {
                   BigwigTrack(GRanges('chr3:34646500-34656000'), 
                               paste0(path_cistrome_bw_folder, data_oi_filt$Factor[i], "/", data_oi_filt$Factor[i], "_", data_oi_filt$DCid[i], ".bw"),
                               smooth = 100) +
                     scale_x_continuous(expand= c(0,0), limits = plot_limits_ChIP) +
                     ylab(data_oi_filt$Factor[i]) +
                     # geom_vline(data = filter(motifs_on_roi_filt_all, Factor == data_oi_filt$Factor[i] & score >300),
                     #            aes(xintercept = start), col = 'green4') +
                     geom_rug(data = filter(motifs_on_roi_filt_all, Factor == data_oi_filt$Factor[i] & score >300),
                                aes(x = start), col = 'black', length = unit(0.2, "npc"), inherit.aes = F) +
                     annotate("rect", xmin = Sox2_gene_annot$cdsStart, xmax = Sox2_gene_annot$cdsEnd, ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red')+
                     theme(legend.position = 'none',
                           axis.title.x = element_blank(),
                           axis.line.x = element_blank(),
                           axis.ticks.x = element_blank(),
                           axis.text.x = element_blank())
                 })

plot_H3K27ac = BigwigTrack(GRanges('chr3:34646500-34656000'), 
                               path_H3K27ac_ChIP,
                               smooth = 100) +
                     scale_x_continuous(expand= c(0,0), limits = plot_limits_ChIP) +
                     ylab('H3K27ac') +
                     
                     annotate("rect", xmin = Sox2_gene_annot$cdsStart, xmax = Sox2_gene_annot$cdsEnd, ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red')+
                     theme(legend.position = 'none',
                           axis.title.x = element_blank(),
                           axis.line.x = element_blank(),
                           axis.ticks.x = element_blank(),
                           axis.text.x = element_blank())

plot_gene_annot = ggplot(data = Sox2_gene_annot) +
  geom_rect(aes(xmin = txStart, xmax = txEnd), ymin = 0.2, ymax = 0.4) +
  geom_rect(aes(xmin = cdsStart, xmax = cdsEnd), ymin = 0, ymax = 0.6) +
  geom_vline(xintercept = start(CTCF_mm10.chr3_extra ), col = 'darkgrey',  lty="11") +
  
  scale_x_continuous(limits = plot_limits_ChIP, expand = c(0,0),
                     breaks = scales::pretty_breaks(n = 4), labels = scales::unit_format(scale = 1E-6, accuracy = 0.002, unit=NULL)) +
  scale_y_continuous(limits = c(0,0.6)) +
  theme_classic() +
  theme(legend.position = 'none')

cowplot::plot_grid(plotlist = c(list(plot_gene_annot),
                                list(plot_H3K27ac),
                                plot_ls),
                   ncol = 1, align="v", axis="lr", heights = c(1, rep(0.5,7)))

```