---
title: "Figure2_CM20250514_reporter_only_hopping"
author: "Mathias & Christine"
date: "2025-03-31"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

#Goal
Plot the integrations (beeswarm plots), expression scores and actual expression for the Sox2P reporter hopping pools and clones. This is figure 2, S2 and S8 of the manuscript.
Because we load all the clonal data here, we also plot figure S6A (which is based on the same clones) 

## Notes for re-running
To re-run this code, change the paths in 'File paths' to the correct location of datasets.
The expression scores are bootstrapped 5000 times, which takes a while to run. Therefore, these chunks are commented out and the bootstrapped data is loaded from a local RDS file. To re-run those, run the bootstrapped expression score calculation once and save the result locally.  

```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
# library(gtools)
library(tidyverse)
# library(dtplyr)
library(dplyr)
# library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(cowplot)
library(RcppRoll)
library(viridis)
library(ggh4x)
library(ggrastr)
library(caTools)
library(readxl)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(GGally)
library(GENOVA)
library(ggpubr)
library(ggridges)
library(ggpmisc)
library(Signac)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}

```

# Paths & parameters
```{r paths_parameters}
#datatag
datetag = paste0("CM",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/Sox2/Figure_reporter_only_hopping/analysis_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}

library(knitr)
opts_chunk$set(cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/"),
               cache.lazy = F) 
pdf.options(useDingbats = FALSE)
```

```{r}
N_bootstrap = 5000
```

## File paths
```{r}
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"

# path_tib_23_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-116kb_Sox2P_pools.txt"
path_tib_23_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/still_to_upload/mapped_integrations_-116kb_Sox2P_pools.txt"

path_tib_C9_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-161kb_Sox2P_pools.txt"

path_tib_C10_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/still_to_upload/mapped_integrations_+715kb_Sox2P_pools.txt"

path_median_mTurq_relevant_cell_lines =  "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250318_median_mTurq_sorting_for_paper.tsv"
path_sorting_stats = "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250327_stats_for_expression_score_for_paper.xlsx"

path_RCMC_WT_file = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/CM20230718_GSM6281849_RCMC_BR1_merged_allCap_WT_mm39.merged.50.mcool"

path_mm10_to_mm39 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm39.over.chain"
path_mm39_to_mm10 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm39ToMm10.over.chain"

paths_ext_data_files = c(
  ATAC = '/DATA/usr/c.moene/public_data/ATAC/GSE98390_E14_ATAC_MERGED.DANPOS.bw',
  H3K27ac = '/DATA/usr/c.moene/public_data/ChIP/Histons/H3K27ac_ENCFF583WVZ_signal_pval.bigWig',
  CTCF = '/DATA/projects/Sox2/CTCF_ChIP_analysis/results/bigwig/6764_2_ME_E2_CGATGT_S2_R1_001.bw' #new, aligned to mm10 only
)

# FACS sorting
diva_xml_file_E2555_sorting = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2534_hopping_C9_LP/FACS/Mathias (BvS)20240501_E2534_rep1.xml"
path_FACS_sorting_E2555 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2534_hopping_C9_LP/FACS"

diva_xml_file_E2096_sorting = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2096_hopping_mTurq_6gates/fcs_for_R_sorting/Diva_export/Mathias (BvS) 20221117_mTurq/Mathias (BvS) 20221117_mTurq.xml"
path_FACS_sorting_E2096 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2096_hopping_mTurq_6gates/fcs_for_R_sorting/Diva_export/Mathias (BvS) 20221117_mTurq"

diva_xml_file_E2687_round2_sorting = "/DATA/projects/Sox2/Zenodo_collection/E2687/Mathias (BvS)20241031_E2687_round2.xml"
path_FACS_sorting_E2687_round2 = "/DATA/projects/Sox2/Zenodo_collection/E2687/Mathias (BvS)20241031_E2687_round2"

#Flowcytometry clones
path_insertion_clones_E2221 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2344_mapping_clones/CM20231218_E2221_locations_all_mapped_clones.rds"
path_link_well_clonename_E2221 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2217_FACS_clones/CM20230309_E2204_link_wellindex_clonename.txt"

path_fcs_files_E2221 = '/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2217_FACS_clones/gated_fcs_files/plate_reader'

path_WT_ctrl_E2250 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2250_CRISPR_Sox2_delBclones_pools/fcs_for_R/export_ME20230418_E2250_pools_WT_001_Single Cells.fcs"

path_fcs_files_E2762 = '/DATA/projects/Sox2/Figure_FACS_copmensation/fcs_for_R'


#functions
source("/DATA/projects/Sox2/General_functions/CM20240621_general_plotting_functions.R")
source("/DATA/projects/Sox2/General_functions/CM20250318_new_expression_score_functions.R")

```



# Load data
From the combined results table
```{r}
#load the -116kb data (LP23)
tib_23_all_data = read_tsv(path_tib_23_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) 

#the control pool in E2285 (rep3) was sequenced as 10 separate libraries, combine them into one sample
tib_23_E2285_ctrls = tib_23_all_data %>%
  filter(replicate == "rep3" & insert == "Sox2P" & sample_type == "control pool") %>%
  mutate(sample_name = "Sox2P-reporter -116 kb, unsorted control pool rep3")%>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") 

tib_23 = tib_23_all_data %>%
  #select the rest of the data
  filter(!(replicate == "rep3" & insert == "Sox2P" & sample_type == "control pool")) %>%
  select(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name,
         read_count, read_count_1, read_count_2)%>%
  #add merged controls back in
  bind_rows(tib_23_E2285_ctrls) %>%
  #mark hopped integrations
  mutate(hopped = start != 34643961)

#load the -161kb data (LP C9)
tib_C9 = read_tsv(path_tib_C9_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  #combine the 2 libraries of "Sox2P-reporter -161 kb, population P1 pool1 rep2"
  mutate(to_merge = case_when(replicate == "rep2" & insert == "Sox2P" & sorted_gate == "P1" ~ "merge",
                              .default = sample_name)) %>%
  mutate(sample_name = case_when(to_merge == "merge" ~ "Sox2P-reporter -161 kb, population P1 pool1 rep2",
                             .default = sample_name)) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  #mark hopped integrations
  mutate(hopped = start != 34598479)
  
#load the +715kb data (LP C10)
tib_C10 = read_tsv(path_tib_C10_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  #mark hopped integrations
  mutate(hopped = start != 35478413)
  



#filter the integrations
tib_filtered <- bind_rows(tib_23,tib_C9, tib_C10)  %>%
  #add variables used in the plotting (old names of cell lines etc) 
  mutate(population = case_when(sample_type == "control pool" ~ "ctrl",
                                .default = sorted_gate),
         cell_line = case_when(launch_pad_location == "-116 kb" & insert == "Sox2P" ~ "23_34A",
                               launch_pad_location == "-161 kb" & insert == "Sox2P" ~ "C9_mCh_34",
                               launch_pad_location == "+715 kb" & insert == "Sox2P" ~ "C10_34"),
         landing_pad = case_when(launch_pad_location == "-116 kb" ~ "23",
                                 launch_pad_location == "-161 kb" ~ "C9",
                                 launch_pad_location == "+715 kb" ~ "C10")) %>%
  #annotate confidence of mapping (one or both ITRs)
  mutate(mapped_arms = case_when(read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only",
                                 read_count_1 > 0 & read_count_2 > 0 ~ "both_arms")) %>%
  
  #filter the data
  filter(population == "ctrl" | read_count >= 2) %>% #for ctrl distribution 1 mapping read is enough
  filter(!(start >= 34721183 & start <= 34721192)) %>% #remove contamination from LP8
  filter(! start %in% c(25018987, 35232946) ) %>% #remove contamination clone C1
  filter(landing_pad  == "C9" | start != 34598479) %>% #remove contamination from LP C9
  filter(landing_pad  == "23" | start != 34643961) %>% #remove contamination from LP 23
  filter(landing_pad  == "C10" | start != 35478413) %>% #remove contamination from LP C10
  filter(strand %in% c("+", "-")) 

tib_filtered_P1_strict = tib_filtered %>%
  filter(!(population == "P1" & mapped_arms != "both_arms")) %>% #only keep P1 integrations with 2 mapped arms
  mutate(population = factor(population)) 
  
```


## Expression data
Because of the large number of unhopped, expressed cells, the median expression in P2-P5 of 23_34A (-116kb) is mostly determined by the 23_34A expression and not by the hopped cells.
In contrast, P1-P4 for C9_34 (-161kb) and C10_34 (+715kb) are only unhopped cells so they give a more meaningful median_mTurq. For P1 C9_34 has no observations, so we use the 23_34A (mCh+ sorted) instead
For P5, none of the cell lines have clean, hopped only observations, but C9_34 is more representative then 23_34A.
For P6, the C9_34 and C10_34 are dominated by the unhopped cells, while 23_34A contains mostly hopped cells. Therefore, we will use the fluorescence of 23_34A for P6.

The median_sorted_mTurq file contains the mT values from the sorter. For C9_mCh+_34 we did not measure any cells in P1, so we used the average of the 23_34A mCh+ sorted replicates (that value is already in the table).
```{r}
#load median expression values (from sorting pdfs)
median_sorted_mTurq = read_tsv(path_median_mTurq_relevant_cell_lines)

#get a general mTurq level per gate, for the mCh+ hopping
general_mTurq_mChpos = median_sorted_mTurq %>% 
  filter(mCherry_selection == "mCh+" & construct == "34") %>%
  mutate(cell_line_group = case_when(cell_line == "23_34A" ~ "expressed",
                                     .default = "silent")) %>%
  group_by(cell_line_group, mCh_status, construct, mCherry_selection, population) %>%
  # mean mTurq per gate per group of cell lines
  summarize(mTurq_mean = mean(mTurq)) %>%
  pivot_wider(names_from = cell_line_group, values_from = mTurq_mean) %>%
  # pick mTurq level to use
  mutate(mTurq_general = case_when(population == "P6" ~ expressed,
                           .default = silent)) %>% 
  ungroup() %>%
  select(construct, population, mTurq_general)


#add mTurq values to the integration tibble
tib_filtered_mTurq_vals = left_join(tib_filtered_P1_strict, general_mTurq_mChpos) %>%
  filter(population != "ctrl") %>%
   dplyr::rename(mTurq_expr_gate = mTurq_general) 

#check that all integrations have a fluorescence value:
tib_filtered_mTurq_vals %>% filter(is.na(mTurq_expr_gate))
``` 

## Sorting statistics
For the new expression score we need to know, per population and replicate:
1. number of cells sorted
2. fraction of the cells in that gate (take fraction of the mChpos/mChneg total cells, whichever is relevant for that cell_line)

We add a psuedocount of 1 to all the 'recorded_count' when calculating the 'recorded_fraction', to deal with the 0 recorded cells in P1. 
Got the sorting stats from the sorting pdfs & the labguru notes of the experiments

```{r}
sorting_stats_tib_raw = read_xlsx(path_sorting_stats) 

sorting_stats_tib = sorting_stats_tib_raw %>%
  mutate(recorded_fraction = (recorded_count + 1)/total_recorded) %>% #recorded fraction with psuedocount of 1
  mutate(weigh_factor = recorded_fraction / sorted_count, 
         weigh_factor_1M = weigh_factor * 1E6) # weigh factor 1M is just to make the factor more easily human readable
```

Due to the 50K pools, 23_34A has more replicates for P3-P5 than for the other populations. We reduce the weight of those integrations, so the total weight counts up to the same number of replicates for each population.

```{r}
# number or replications per population
sorting_stats_tib_repN = sorting_stats_tib %>%
  filter(cell_line %in% c("C9_mCh_34", "C10_34", "23_34A")) %>%
  group_by(cell_line, population) %>%
  summarize(N_reps = n()) %>% #number of reps per population per cell line
  group_by(cell_line) %>%
  mutate(min_reps_CL = min(N_reps) ) %>% #least replicated population
  mutate(rep_weigh_factor = min_reps_CL/N_reps) #scale the other populations to the least replicated pop

# adjust weigh factor by the replicate weigh factor
sorting_stats_tib_repcor = 
  sorting_stats_tib %>%
  left_join(sorting_stats_tib_repN) %>%
  mutate(weigh_factor_old = weigh_factor,
         weigh_factor = weigh_factor_old * rep_weigh_factor)

```

Rough overview of the stats
```{r}
tib_filtered_mTurq_vals %>% 
  filter(cell_line %in% c("C9_mCh_34", "C10_34", "23_34A")) %>%
  group_by(cell_line, experiment, replicate, population) %>%
  filter(hopped) %>%
  summarize(N_ints = n()) %>%
  left_join(sorting_stats_tib)

```

#some setups
```{r}
all_exp = unique(tib_filtered_P1_strict$experiment)


# Location of enhancer / gene
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")
Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")

landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")

landingPad_C9 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34598479,
                                     end = 34598480),
                    strand = "+")
landingPad_C10 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 35478413,
                                     end = 35478414      ),
                    strand = "+")

LP_tib = tibble(landing_pad = c("23", "C9", "C10"), 
                start = c( start(landingPad_23),start(landingPad_C9), start(landingPad_C10)))


#CTCF sites based on our own mapping
# new analysis (based on mapping to mm10 only)
CTCF_mm10_new <- import.bed(path_CTCF_sites)

#filtering:
CTCF_mm10.chr3 <- CTCF_mm10_new[seqnames(CTCF_mm10_new)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)


```

General colors and ranges
```{r}
# Plotting ranges
prange_zoom = c(34.5E6, 35E6)
prange_zoom_paper = c(34.59E6, 34.83E6)
prange_SCR = c(34.725E6, 34.785E6)


# Color annotations
## SCR 
brown = "#ffb000"

#color schemes
myCols_blue = c('#525252', "#080E5B",  "#212B71", "#3A4987", "#465893", "#5A719A", "#6C8AA0")
names(myCols_blue) = c("ctrl", paste0("P", 1:6))
colScale7_blue <- scale_colour_manual(name = "population", values = myCols_blue)
fillScale7_blue <- scale_fill_manual(name = "population", values = myCols_blue)

myCols_blue_strand = c( "#080E5B", "#5A719A")
names(myCols_blue_strand) = c("+","-")
colScale2_blue <- scale_colour_manual(name = "strand", values = myCols_blue_strand)
fillScale2_blue <- scale_fill_manual(name = "strand", values = myCols_blue_strand)
```


# Functions
## Plotting beeswarm
```{r}
plot_beeswarm_only = function(TIB, CL = "23_34A", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = F, PLOT_ANNOT = T,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom",
                              NEXT_RANGE = NULL){
  tib_for_plot = filter(TIB, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP & start >= P_RANGE[1] & start <= P_RANGE[2])
  LP_tib_cl = select(tib_for_plot, c(cell_line, landing_pad)) %>% 
    distinct() %>% left_join(LP_tib)
  
  ggplot(filter(tib_for_plot, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP),
         aes(x = start, y = population, col = population)) +
    {if(length(CL)>1) facet_grid(cell_line ~ .)} +  #only facet when necessary

    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    
     #annotate landing pad
    {if(PLOT_ANNOT & length(CL)== 1)
      geom_vline(xintercept = pull(
        filter(LP_tib, landing_pad %in% unique(tib_for_plot$landing_pad)), 
        start),
        col = 'darkgrey')} +
    {if(PLOT_ANNOT & length(CL) > 1)
       geom_vline(data = LP_tib_cl, aes(xintercept = start), col = 'darkgrey')} +

    # # annotate CTCFs
    {if(PLOT_CTCF)geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty="11")}+
    {if(!is.null(NEXT_RANGE)) geom_rug(data = tibble(range = NEXT_RANGE),
                                       aes(x = range),
                                       fill = 'grey',
                                       inherit.aes = F)}+
    
    
    #datapoints
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = T, bandwidth = 0.8,
                     # shape = 16, #changes to point without border, so alpha applies to the whole point
                     method = RANDOM_METHOD,
                     stroke = NA) +
    geom_text(data = plyr::count(tib_for_plot, vars = c("cell_line", "population")), aes(label = paste("n = ", freq, sep = ""),
                                                                                         x = Inf),
              hjust = "inward", vjust = "top",
              col = 'black') +
    
    #layout:
    labs(x='Genomic coordinate (Mb)',y='sorted gate') +
    theme_classic(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                       limits=P_RANGE, expand=c(0,0)) +
    scale_y_discrete(limits = rev, drop=F) +
    ggtitle(TITLE) +
    colScale7_blue
}
```


# Beeswarm
## Panel C: zoom-out
```{r C_beeswarm_23_34A_zoomout, fig.width = 4, fig.height = 2.5}


range_2Mb = c(33.643960E6, 35.643960E6)
plot_beeswarm_only(TIB = filter(tib_filtered_P1_strict, hopped & chr == "chr3" & experiment %in% all_exp),
                   CL = "23_34A",
                   PLOT_ANNOT = T,
                   PLOT_CTCF = F,
                   P_RANGE = range_2Mb,
                   POINT_SIZE = 1,
                   NEXT_RANGE = prange_zoom_paper
)


```

```{r S2C_beeswarm_23_34A_zoomout_50Kpools, fig.width = 6, fig.height = 3}

range_2Mb = c(33.643960E6, 35.643960E6)
beesw_rep4 = plot_beeswarm_only(TIB = filter(tib_filtered_P1_strict, hopped & chr == "chr3" & experiment %in% c("E2712", "E2716")),
                   CL = "23_34A",
                   PLOT_ANNOT = T,
                   PLOT_CTCF = F,
                   P_RANGE = range_2Mb,
                   POINT_SIZE = 1
)

beesw_rep123 = plot_beeswarm_only(TIB = filter(tib_filtered_P1_strict, hopped & chr == "chr3" & !experiment %in% c("E2712", "E2716")),
                   CL = "23_34A",
                   PLOT_ANNOT = T,
                   PLOT_CTCF = F,
                   P_RANGE = range_2Mb,
                   POINT_SIZE = 1
)

cowplot::plot_grid(beesw_rep123,
                   beesw_rep4,
                   nrow = 1, align="h", axis="tb")

```


```{r S2E_beeswarm_C10_34_zoomout,  fig.width = 4, fig.height = 2.5}
plot_beeswarm_only(TIB = filter(tib_filtered_P1_strict, hopped & chr == "chr3" & experiment %in% all_exp),
                   CL = "C10_34",
                   PLOT_ANNOT = T,
                   PLOT_CTCF = F,
                   P_RANGE = c(33.5E6, 36E6),
                   POINT_SIZE = 1
)



```

## Panel D-E: prange zoom
LP23, C9 and C10 (sup)

```{r DG_beeswarm_only_prange, fig.width = 6, fig.height = 6}
plot_beeswarm_only(TIB = filter(tib_filtered_P1_strict, hopped & chr == "chr3" &  experiment %in% all_exp ),
                   CL = c("23_34A", "C9_mCh_34"),
                   PLOT_ANNOT = T,
                   PLOT_CTCF = T,
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1.5
)

```


```{r S2F_C10_beeswarm_prange, fig.width = 6, fig.height = 3}
plot_beeswarm_only(TIB = filter(tib_filtered_P1_strict, hopped & chr == "chr3" &  experiment %in% all_exp ),
                   CL = c("C10_34"),
                   PLOT_ANNOT = T,
                   PLOT_CTCF = T,
                   P_RANGE = prange_zoom_paper,
                   POINT_SIZE = 1.5
)

```



## Orientation CTCF sites
Plotting triangles in a pdf is a nightmare, so instead 1 just make one plot with the CTCF lines colored by orientation and based on that add the CTCF triangles manually to the paper figures.

```{r}
ggplot() +
   # annotate enhancer
  # annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown)+
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown)+
  # annotate gene
  # annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red')+
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') +
  
  
  geom_vline(data = as_tibble(CTCF_mm10.chr3_extra), aes(xintercept = start, col = strand), linewidth = 2, lty = 'dashed') +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  theme_bw()

```




# Expression scores


5000x, per cell line, save result

## Main: combined score

```{r}
#all combined, 5kb window
set.seed(20240331)
min_ints_window = 3
N_bootstrap = 5000
BinSize = 500
CL = "34_mChpos"
all_exp = unique(tib_filtered_mTurq_vals$experiment)
tib_filtered_mTurq_vals_comb = tib_filtered_mTurq_vals %>% left_join(sorting_stats_tib_repcor) %>%
  filter(cell_line %in% c("23_34A", "C9_mCh_34", "C10_34")) %>%
  mutate(cell_line = "34_mChpos")


# # calculate running mean on random sample of same size, 500 times
# expr_score_tib = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 10,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = CL, EXP = all_exp)},
#                                             TIB = tib_filtered_mTurq_vals_comb,
#                                             N_resamples = N_bootstrap)
# saveRDS(expr_score_tib, "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250331_new_expression_score_34_mChpos_5k.rds")

expr_score_tib = readRDS("/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250331_new_expression_score_34_mChpos_5k.rds")



expr_score_sum_tib = bootstrap_summary_fun(expr_score_tib,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real =
  expr_score_function_new_weigh(tib_filtered_mTurq_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = CL,
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb = left_join(expr_score_sum_tib,  expr_score_real) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA)) %>%
  mutate(cell_line = CL)
  


```


```{r H_expression_score_prange_comb, fig.width = 8.5, fig.height = 5}
expr_score_plot_34_mChpos = ggplot(expr_score_sum_tib_filt_comb, aes(x = mid_interval)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel), alpha = 0.2, fill = "#080E5B") +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel), col = "#080E5B") +

  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')
expr_score_plot_34_mChpos
```

Calculate how far activation spreads downstream of SCR
```{r get stats for upstream activation}
# Get stats
## Look what the levels of P5 and P6 is for the three different cell lines in that plot C9, C10, 23

tib_P_stats = median_sorted_mTurq %>% 
  filter(cell_line %in% c("23_34A", "C10_34", "C9_mCh_34") & population %in% c("P5", "P6")) %>% 
  group_by(population) %>% 
  summarise(mean_P = mean(mTurq))


expr_score_sum_tib_filt_comb %>% filter(expr_score_window_median_sel <= 525.5 & mid_interval >= end(enhancer)) %>% arrange(mid_interval)


ggplot(expr_score_sum_tib_filt_comb, aes(x = mid_interval)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +
  
  # Line downstream SCR spread of activity
  geom_hline(yintercept = 525.5) +
  geom_vline(xintercept = 34785250) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel), alpha = 0.2, fill = "#080E5B") +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel), col = "#080E5B") +

  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')

```



## Supplementary: Score per cell line
Calculate (bootstrapped) expression score

```{r}
#23_34A 10kb window
set.seed(20240327)
min_ints_window = 3
N_bootstrap = 5000
BinSize = 1000
CL = "23_34A"
all_exp = unique(tib_filtered_mTurq_vals$experiment)
tib_filtered_mTurq_vals_repcor = tib_filtered_mTurq_vals %>% left_join(sorting_stats_tib_repcor)


# # calculate running mean on random sample of same size, 500 times
# expr_score_tib = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 10,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = CL, EXP = all_exp)},
#                                             TIB = tib_filtered_mTurq_vals_repcor,
#                                             N_resamples = N_bootstrap)
# saveRDS(expr_score_tib, "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_23_34A_10kb.rds")

expr_score_tib = readRDS("/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_23_34A_10kb.rds")



expr_score_sum_tib = bootstrap_summary_fun(expr_score_tib,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real =
  expr_score_function_new_weigh(tib_filtered_mTurq_vals_repcor,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = CL,
                                EXP = all_exp)
  
expr_score_sum_tib_filt_23_34A = left_join(expr_score_sum_tib,  expr_score_real) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA)) %>%
  mutate(cell_line = CL)
  



```




```{r}
#C10_34 10kb window
set.seed(20240327)
min_ints_window = 3
N_bootstrap = 5000
BinSize = 1000
CL = "C10_34"
all_exp = unique(tib_filtered_mTurq_vals$experiment)
tib_filtered_mTurq_vals_repcor = tib_filtered_mTurq_vals %>% left_join(sorting_stats_tib_repcor)

# 
# # calculate running mean on random sample of same size, 500 times
# expr_score_tib = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 10,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = CL, EXP = all_exp)},
#                                             TIB = tib_filtered_mTurq_vals_repcor,
#                                             N_resamples = N_bootstrap)
# saveRDS(expr_score_tib, "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_C10_34_10kb.rds")

expr_score_tib = readRDS("/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_C10_34_10kb.rds")



expr_score_sum_tib = bootstrap_summary_fun(expr_score_tib,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real =
  expr_score_function_new_weigh(tib_filtered_mTurq_vals_repcor,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = CL,
                                EXP = all_exp)
  
expr_score_sum_tib_filt_C10 = left_join(expr_score_sum_tib,  expr_score_real) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA)) %>%
  mutate(cell_line = CL)
  

```



```{r}
#C9_mCh+_34 5kb window
set.seed(20240327)
min_ints_window = 3
N_bootstrap = 5000
BinSize = 500
CL = "C9_mCh_34"
all_exp = unique(tib_filtered_mTurq_vals$experiment)
tib_filtered_mTurq_vals_repcor = tib_filtered_mTurq_vals %>% left_join(sorting_stats_tib_repcor)


# # calculate running mean on random sample of same size, 500 times
# expr_score_tib = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 10,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = CL, EXP = all_exp)},
#                                             TIB = tib_filtered_mTurq_vals_repcor,
#                                             N_resamples = N_bootstrap)
# saveRDS(expr_score_tib, "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_C9_mCh_34_5kb.rds")

expr_score_tib = readRDS("/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_C9_mCh_34_5kb.rds")



expr_score_sum_tib = bootstrap_summary_fun(expr_score_tib,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real =
  expr_score_function_new_weigh(tib_filtered_mTurq_vals_repcor,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = CL,
                                EXP = all_exp)
  
expr_score_sum_tib_filt_C9_34 = left_join(expr_score_sum_tib,  expr_score_real) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA)) %>%
  mutate(cell_line = CL)
  
# #save filtered table, to plot in fig 4&5
# saveRDS(expr_score_sum_tib_filt, "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_sum_filt_C9_mCh_34_5kb.rds")
```


```{r S8E_expr_score_per_cel, fig.width = 8.5, fig.height = 5}

expr_score_sum_per_cl = bind_rows(C9_mCh_34 = expr_score_sum_tib_filt_C9_34, 
                                  `23_34A` = expr_score_sum_tib_filt_23_34A, 
                                  C10_34 = expr_score_sum_tib_filt_C10, 
                                  .id = 'cell_line') %>%
  mutate(cell_line = factor(cell_line, levels = c("C9_mCh_34", "23_34A", "C10_34")))


#combine cell line specific with combined expr scores
comb_expr_score_for_plot = 
  expr_score_sum_tib_filt_comb %>% ungroup() %>% 
  select(mid_interval, expr_score_window_median_sel) %>%
  dplyr::rename(expr_score_comb = expr_score_window_median_sel)

comb_expr_score_for_plot_long = bind_rows(C9_mCh_34 = comb_expr_score_for_plot,
                                          `23_34A` = comb_expr_score_for_plot,
                                          C10_34 = comb_expr_score_for_plot, .id = 'cell_line')
LP_tib_for_plot = LP_tib %>%
  mutate(cell_line = case_when(landing_pad == "23" ~ "23_34A",
                               landing_pad == "C9" ~ "C9_mCh_34",
                               landing_pad == "C10" ~ "C10_34"))

#plot expression scores
expr_score_sum_per_cl %>%
  ggplot(aes(x = mid_interval)) +
  facet_grid(.~ factor(cell_line, levels = c("C9_mCh_34", "23_34A", "C10_34"))) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(data = LP_tib_for_plot, aes(xintercept = start), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +
  #plot combined data
  geom_line(data = comb_expr_score_for_plot_long, aes(y = expr_score_comb ), col = 'grey30', linewidth = 0.4) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel), alpha = 0.2, fill = "#080E5B") +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel), col = "#080E5B", linewidth = 0.6) +
  

  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')
  

```


## Panel H: annotations


### Virtual 4C
Load RCMC data
```{r}

RCMC_1kb = load_contacts(signal_path = path_RCMC_WT_file,
                      sample_name = 'WT',
                      resolution = 1000)

```

Calculate virtual 4C in 1kb bins, from the SCR
```{r}
# download.file('https://hgdownload.soe.ucsc.edu/goldenPath/mm39/liftOver/mm39ToMm10.over.chain.gz',
#               "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm39ToMm10.over.chain.gz")
# download.file('https://hgdownload.soe.ucsc.edu/goldenPath/mm10/liftOver/mm10ToMm39.over.chain.gz',
#               "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm39.over.chain.gz")

# library(R.utils)
# gunzip("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm39ToMm10.over.chain.gz")
# gunzip("/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/general_analyses/region_capture_microC/mm10ToMm39.over.chain.gz")


chain_mm10_to_mm39 = import.chain(path_mm10_to_mm39)
chain_mm39_to_mm10 = import.chain(path_mm39_to_mm10)

#ranges on mm39
prange_paper_mm39_gr = liftOver(GRanges(seqnames = "chr3", IRanges(start = prange_zoom_paper[1], end = prange_zoom_paper[2]) ), chain_mm10_to_mm39)
prange_paper_mm39 = c(start(unlist(prange_paper_mm39_gr)), end(unlist(prange_paper_mm39_gr)))

prange_2MB_mm39_gr = liftOver(GRanges(seqnames = "chr3", IRanges(start = range_2Mb[1], end = range_2Mb[2]) ), chain_mm10_to_mm39)
prange_2Mb_mm39 = c(start(unlist(prange_2MB_mm39_gr)), end(unlist(prange_2MB_mm39_gr)))
                      
#core SCR
core_SCR_mm10 = GRanges("chr3:34757573-34762064") #brosh et al bins 24 & 26, the 2 essential elements in the SCR
core_SCR_mm39 = liftOver(core_SCR_mm10, chain_mm10_to_mm39) 
core_SCR_mm39_v = c(start(unlist(core_SCR_mm39)), end(unlist(core_SCR_mm39)))

# SCR_virt4C_mm39_core = virtual_4C(RCMC_1kb, unlist(core_SCR_mm39), 
#                              prange_paper_mm39)

# virtual 4C xlim: give the number of bases flanking the VP that you want to calculate
xlim_2Mb_region = c(core_SCR_mm39_v[1]-prange_2Mb_mm39[1], prange_2Mb_mm39[2]-core_SCR_mm39_v[2])
  
#calculate virtual 4C
SCR_virt4C_mm39_core = SCR_virt4C_mm39_core = virtual_4C(RCMC_1kb, unlist(core_SCR_mm39), 
                             xlim = xlim_2Mb_region) 

#clean up tibble
SCR_virt4C_mm39_GR_core= SCR_virt4C_mm39_core$data %>% mutate(start = mid - 500,
                             end = mid + 500) %>% GRanges()

#lift back to mm10
SCR_virt4C_mm10_core = liftOver(GRanges(SCR_virt4C_mm39_GR_core), chain_mm39_to_mm10) %>% unlist()


```

Plot virtual 4C track
```{r I_expr_scores_virt4C, width = 8, height = 5}
virt_4c_plot_core = 
  as_tibble(SCR_virt4C_mm10_core) %>%
  mutate(mid_interval = (start + end)/2) %>%
  select(-mid) %>%
  ggplot(aes(x = mid_interval, y = WT)) + 
  
  # annotate enhancer
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer),
           ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), 
           ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red') +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') +
  # annotate CTCFs
  geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', linetype = 'dotted') +
  geom_col(fill = 'grey30', col = 'grey30') + 
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 6), 
    labels = scales::unit_format(scale = 1E-6, accuracy = 0.05, unit=NULL), 
    limits = prange_zoom_paper,
    expand = c(0,0)) +
  labs(x='Genomic coordinate (Mb)',y='contact with SCR') +
  theme_classic(base_size = 16) +
  coord_cartesian(ylim = c(0, 2000), expand = F) +
  theme(legend.position = 'none') +
  geom_rug(data = tibble(pos = c(start(core_SCR_mm10), end(core_SCR_mm10))), 
           inherit.aes = F, aes(x = pos), col = 'gold')

cowplot::plot_grid(plotlist = list(expr_score_plot_34_mChpos, virt_4c_plot_core),
                   ncol = 1, align="v", axis="lr", rel_heights = c(3,2))

```

### Combined plot

The function BigwigTrack from the Signac package gives nicer plots than our own code, so use that instead.
One disadvantage: if we want to scale the y-axis per track, the values are removed (scaled to 1), so to get the actual y-axis back I need to plot the tracks one by one.
Is not a problem, just need to remove a lot of unnecessary axis labels etc in illustrator

```{r I_expr_scores_virt4c_ChIP, fig.width = 6, fig.height = 8}

prange_zoom_paper_gr = GRanges(seqnames = "chr3", IRanges(start = prange_zoom_paper[1], end = prange_zoom_paper[2]))
ext_data_files_ls = split(unname(paths_ext_data_files),names(paths_ext_data_files))

# plot_bw_signac = BigwigTrack(prange_zoom_paper_gr, ext_data_files_ls[c(1,3,2)], bigwig.scale = 'separate', smooth = 500)+
#   theme(legend.position = 'none')
# 
# cowplot::plot_grid(plotlist = c(list(expr_score_plot_34_mChpos, virt_4c_plot_core, plot_bw_signac)),
#                    ncol = 1, align="v", axis="lr", rel_heights = c(1, 0.6, 0.6))

cowplot::plot_grid(plotlist = c(list(expr_score_plot_34_mChpos, virt_4c_plot_core, 
                                     BigwigTrack(prange_zoom_paper_gr, ext_data_files_ls[1], smooth = 500) + 
  scale_x_continuous(expand= c(0,0)),
                                     BigwigTrack(prange_zoom_paper_gr, ext_data_files_ls[3], smooth = 500) + 
  scale_x_continuous(expand= c(0,0)),
                                     BigwigTrack(prange_zoom_paper_gr, ext_data_files_ls[2], smooth = 500) + 
  scale_x_continuous(expand= c(0,0))
                                     )),
                   ncol = 1, align="v", axis="lr", rel_heights = c(1, 0.6, 0.3, 0.3, 0.3))

```

## Panel J: correlation with contact - all data

Calculate binned expression score
```{r cor_contact_expr_score_all_data, width = 4, height = 4}
# 1kb resolution
exp_score_mm10_bins = expr_score_function_new_weigh(tib_filtered_mTurq_vals_comb,
                      binsize = 1000,
                      N_bins_window = 1,
                      calculation_window =  c(min(start(SCR_virt4C_mm10_core)), max(end(SCR_virt4C_mm10_core))), #c(34590000, 34830000),
                      CL = c("34_mChpos"), EXP = all_exp)

contact_expr_tib_core = left_join(exp_score_mm10_bins, as_tibble(SCR_virt4C_mm10_core), by = join_by("start_interval" == "start")) %>%
  mutate( WT = replace_na(WT, 0))

```

Color the points that are between gene and SCR in blue, the rest black.
It is hard in R to give only the correlation for one subset of the data, so instead give it for both and the whole data (purple here) and then remove the black line and recolor the purple in illustrator, to clearly show the blue subset of the data and the whole interval

```{r J_cor_virt4C_34_mChpos_alldata, width = 8, height = 4}
tib_for_plot_RCMC = filter(contact_expr_tib_core, 
              # #exclude the bins on the SCR), 
              (start_interval < start(core_SCR_mm10)-1000 | start_interval > end(core_SCR_mm10)) &
                !is.na(total_ints_bin) & total_ints_bin >= 3 &
                WT > 0 ) %>%
  filter(!is.na(mid_interval)) %>%
  mutate(between_gene_SCR = start_interval >= end(Sox2_gene) + 5000 & start_interval <= start(enhancer) - 6000) %>%
  mutate(dist_SCR = abs(start(resize(core_SCR_mm10, width = 1, fix = 'center')) - mid_interval)) %>%
  arrange(between_gene_SCR)


cor_plot_RCMC_contact = 
  ggplot(tib_for_plot_RCMC, 
       aes(x = WT, y = expr_score_window, col = between_gene_SCR)) +
  #data
  geom_point(alpha = 0.5, size = 2.5) +
  stat_cor(method = "spearman", 
           label.x.npc = "right", hjust = 1,
           label.y.npc = 'bottom') +
  stat_cor(method = "spearman", 
           label.x.npc = "left", 
           label.y.npc = 'top', col = 'purple') +
   #data
  stat_poly_line(col = 'purple', method = 'lm') +
  stat_poly_line(method = 'lm') +
  
  #layout:
  scale_x_log10()+
  scale_y_log10(limits = c(200, NA),expand = c(0,0.1))+
  
  labs(x = "contact with SCR", y = "expression score bin") +
 
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, legend.position = 'none') +
  scale_color_manual(values = c(`TRUE` = '#0077BB', `FALSE` = 'black'))


cor_plot_genom_dist = 
  ggplot(tib_for_plot_RCMC, aes(x = dist_SCR/1000, y = expr_score_window, col = between_gene_SCR)) +
  #data
  geom_point(alpha = 0.5, size = 2.5) +
  stat_cor(method = "spearman", 
           label.x.npc = "left", hjust = 0,
           label.y.npc = 'bottom') +
  stat_cor(method = "spearman", 
           label.x.npc = "right", hjust = 1,
           label.y.npc = 'top', col = 'purple') +
  #data
  stat_poly_line(col = 'purple', method = 'lm') +
  stat_poly_line(method = 'lm') +

  #layout:
  scale_x_log10()+
  scale_y_log10(limits = c(200, NA), expand = c(0,0.1))+
  # 
  labs(x = "genomic distance from SCR (kb)", y = "expression score bin") +

  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, legend.position = 'none') +
  scale_color_manual(values = c(`TRUE` = '#0077BB', `FALSE` = 'black'))

cowplot::plot_grid(plotlist = list(cor_plot_genom_dist, cor_plot_RCMC_contact),
                   align = 'h', nrow = 1, axis = 'tb')

```


## supp: correlation with contact - C9

Calculate binned expression score
```{r cor_contact_expr_score_C9, width = 4, height = 4}
# 1kb resolution
exp_score_mm10_bins_C9 = expr_score_function_new_weigh( left_join(tib_filtered_mTurq_vals, sorting_stats_tib_repcor),
                      binsize = 1000,
                      N_bins_window = 1,
                      calculation_window =  c(min(start(SCR_virt4C_mm10_core)), max(end(SCR_virt4C_mm10_core))), #c(34590000, 34830000),
                      CL = c("C9_mCh_34"), EXP = all_exp)

contact_expr_tib_core_C9 = left_join(exp_score_mm10_bins_C9, as_tibble(SCR_virt4C_mm10_core), by = join_by("start_interval" == "start")) %>%
  mutate( WT = replace_na(WT, 0))

```

Color the points that are between gene and SCR in blue, the rest black.
It is hard in R to give only the correlation for one subset of the data, so instead give it for both and the whole data (purple here) and then remove the black line and recolor the purple in illustrator, to clearly show the blue subset of the data and the whole interval

```{r S2K__cor_virt4C_34_mChpos_C9, width = 8, height = 4}
tib_for_plot_RCMC_C9 = filter(contact_expr_tib_core_C9, 
              # #exclude the bins on the SCR), 
              (start_interval < start(core_SCR_mm10)-1000 | start_interval > end(core_SCR_mm10)) &
                !is.na(total_ints_bin) & total_ints_bin >= 3 &
                WT > 0 ) %>%
  filter(!is.na(mid_interval)) %>%
  mutate(between_gene_SCR = start_interval >= end(Sox2_gene) + 5000 & start_interval <= start(enhancer) - 6000) %>%
  mutate(dist_SCR = abs(start(resize(core_SCR_mm10, width = 1, fix = 'center')) - mid_interval)) %>%
  arrange(between_gene_SCR)

cor_plot_RCMC_contact = 
  ggplot(tib_for_plot_RCMC_C9, 
       aes(x = WT, y = expr_score_window, col = between_gene_SCR)) +
  #data
  geom_point(alpha = 0.5, size = 2.5) +
  stat_cor(method = "spearman", 
           label.x.npc = "right", hjust = 1,
           label.y.npc = 'bottom') +
  stat_cor(method = "spearman", 
           label.x.npc = "left", 
           label.y.npc = 'top', col = 'purple') +
   #data
  stat_poly_line(col = 'purple', method = 'lm') +
  stat_poly_line(method = 'lm') +
  
  #layout:
  scale_x_log10()+
  scale_y_log10(limits = c(200, NA),expand = c(0,0.1))+
  
  labs(x = "contact with SCR", y = "expression score bin") +
  
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, legend.position = 'none') +
  scale_color_manual(values = c(`TRUE` = '#0077BB', `FALSE` = 'black'))


cor_plot_genom_dist = 
  ggplot(tib_for_plot_RCMC_C9, aes(x = dist_SCR/1000, y = expr_score_window, col = between_gene_SCR)) +
  #data
  geom_point(alpha = 0.5, size = 2.5) +
  stat_cor(method = "spearman", 
           label.x.npc = "left", hjust = 0,
           label.y.npc = 'bottom') +
  stat_cor(method = "spearman", 
           label.x.npc = "right", hjust = 1,
           label.y.npc = 'top', col = 'purple') +
  #data
  stat_poly_line(col = 'purple', method = 'lm') +
  stat_poly_line(method = 'lm') +
   
  #layout:
  scale_x_log10()+
  scale_y_log10(limits = c(200, NA), expand = c(0,0.1))+
  
  labs(x = "genomic distance from SCR (kb)", y = "expression score bin") +
  
  theme_classic(base_size = 14) +
  theme(aspect.ratio = 1, legend.position = 'none') +
  scale_color_manual(values = c(`TRUE` = '#0077BB', `FALSE` = 'black'))

cowplot::plot_grid(plotlist = list(cor_plot_genom_dist, cor_plot_RCMC_contact),
                   align = 'h', nrow = 1, axis = 'tb')

```



# Supplementary:
## Reproducibility integration count
Calculate (per replicate) fraction of all P1, P2 etc integrations in each bin across the locus, using 5kb bins in the 'prange_zoom_paper'.
Plot these against each other (R1 vs R2, R1 vs R3, R2 vs R3)

### 23_34A
```{r S8A_replicates_23_34A_int_count, fig.width= 5, fig.height=5}
bin_borders = seq(from = prange_zoom_paper[1], to = prange_zoom_paper[2], by = 5000)
bin_borders_tib = tibble(start_window = bin_borders) %>%
  mutate(interval_number = row_number())

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = tib_filtered_P1_strict %>% 
  filter(cell_line == "23_34A" & hopped) %>%
  group_by(experiment, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = tib_filtered_P1_strict %>% 
  filter(chr == "chr3" & cell_line == "23_34A" & hopped) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(experiment, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = experiment,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

#plot pairs
filter(frac_ints_bins_wide_full, 
               population != 'ctrl' & 
                 interval_number != c(0, max(interval_number))) %>% #remove the outer bins, which are bigger than 5kb
  dplyr::rename(replicate_1 = E2138,
                replicate_2 = E2282,
                replicate_3 = E2285,
                replicate_4 = E2712) %>%
  mutate(replicate_4 = case_when(population %in% c("P1", "P2", "P6") ~ NA, #replicate 4 only contained P3-P5
                                 .default = replicate_4)) %>%
  ggpairs(aes(col = population),
        columns = paste0('replicate_',1:4),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  theme_bw(base_size = 12) 

``` 
I think it is best to use these arbitrary colors from R, because the blue scale from the beeswarms has too similar colors for this type of plot.
For some reason the text for rep4 does not follow the correct colors, manually adjust those in illustrator

### C9_mCh_34
C9 is split into 'replicates', not experiments, and no problems with replicates of only some populations. Otherwise same code as for 23_34A
```{r S8B_replicates_C9_int_count, fig.width= 3, fig.height=3}

#count the number of integrations per cell line, population and bin
N_ints_pop_tib = tib_filtered_P1_strict %>% 
  filter(cell_line == "C9_mCh_34" & hopped) %>%
  group_by(replicate, population) %>%
  summarize(N_ints_pop = n())

N_ints_bins_tmp = tib_filtered_P1_strict %>% 
  filter(chr == "chr3" & cell_line == "C9_mCh_34" & hopped) %>%
  mutate(interval_number = findInterval(start, bin_borders)) %>%
  group_by(replicate, population, interval_number) %>%
  summarize(N_ints_bin = n()) %>%
  left_join(N_ints_pop_tib) %>%
  mutate(fract_ints_bin = N_ints_bin/N_ints_pop)

frac_ints_bins_wide = N_ints_bins_tmp %>%
  pivot_wider(id_cols = c(population, interval_number),
              names_from = replicate,
              values_from = fract_ints_bin,
              values_fill = 0)

frac_ints_bins_wide_full = 
  #make all combinations of interval_number and population
  expand_grid(interval_number = c(0, bin_borders_tib$interval_number),
              population = unique(N_ints_bins_tmp$population)) %>%
  #add the fractions and replace all NAs with zeros
  left_join(frac_ints_bins_wide) %>%
  mutate(across(everything(), ~replace_na(.x, 0))) 

#plot pairs
filter(frac_ints_bins_wide_full, 
               population != 'ctrl' &
                 interval_number != c(0, max(interval_number))) %>% #remove the outer bins, which are bigger than 5kb
  dplyr::rename(replicate_1 = rep1,
                replicate_2 = rep2) %>%
  ggpairs(aes(col = population),
        columns = paste0('replicate_',1:2),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  # coord_cartesian(xlim = c(0, 0.25), ylim = c(0, 0.25)) +
  theme_bw(base_size = 12) 

``` 
## Reproducibility expression score

###23_34A
reproducibility on the expression score can only be done for rep1-3, because rep4 does not cover all populations.
```{r S8C_reproducibility_expr_score_23_34A, fig.width = 4, fig.height = 4}
exp_score_per_experiment_ls = lapply(paste0("rep", 1:3), function(replicateN){
  expr_score_function_new_weigh(filter(left_join(tib_filtered_mTurq_vals, sorting_stats_tib_repcor), replicate == replicateN),
                      binsize = 5000,
                      N_bins_window = 1,
                      calculation_window =  prange_zoom_paper,
                      CL = "23_34A", EXP = all_exp)
})
names(exp_score_per_experiment_ls) = paste0("replicate_", 1:3)

exp_score_per_experiment_tib = bind_rows(exp_score_per_experiment_ls, .id = "replicate")

exp_score_per_experiment_tib_wide = exp_score_per_experiment_tib %>%
  select(replicate, interval_number, start_interval, mid_interval, expr_score_window) %>%
  pivot_wider(names_from = replicate, values_from = expr_score_window)

exp_score_per_experiment_tib_wide %>%

  ggpairs(columns = paste0('replicate_',1:3),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  theme_bw(base_size = 12)

```


### C9_mCh+_34
```{r reproducibility_expr_score_C9_34, fig.width = 4, fig.height = 4}
exp_score_per_experiment_ls = lapply(paste0("rep", 1:2), function(replicateN){
  expr_score_function_new_weigh(filter(left_join(tib_filtered_mTurq_vals, sorting_stats_tib_repcor), replicate == replicateN),
                      binsize = 5000,
                      N_bins_window = 1,
                      calculation_window =  prange_zoom_paper,
                      CL = "C9_mCh_34", EXP = all_exp)
})
names(exp_score_per_experiment_ls) = paste0("replicate_", 1:2)

exp_score_per_experiment_tib = bind_rows(exp_score_per_experiment_ls, .id = "replicate")

exp_score_per_experiment_tib_wide = exp_score_per_experiment_tib %>%
  select(replicate, interval_number, start_interval, mid_interval, expr_score_window) %>%
  pivot_wider(names_from = replicate, values_from = expr_score_window)

exp_score_per_experiment_tib_wide %>%

  ggpairs(columns = paste0('replicate_',1:2),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  theme_bw(base_size = 12)

```

### between launch pads
```{r S8F_reproducibility_between_LPs, fig.width = 4, fig.height = 4}
exp_score_per_LP_binned =  expr_score_function_new_weigh(left_join(tib_filtered_mTurq_vals, sorting_stats_tib_repcor),
                      binsize = 5000,
                      N_bins_window = 1,
                      calculation_window =  prange_zoom_paper,
                      CL = c("23_34A", "C9_mCh_34", "C10_34"), EXP = all_exp)


exp_score_per_LP_tib_wide = exp_score_per_LP_binned %>%
  select(cell_line, interval_number, start_interval, mid_interval, expr_score_window) %>%
  pivot_wider(names_from = cell_line, values_from = expr_score_window)


exp_score_per_LP_tib_wide %>%
  ggpairs(columns = c("23_34A", "C9_mCh_34", "C10_34"),
        lower = list(continuous = "points", stroke = NA), #, mapping=aes(color=population
        upper = list(continuous = wrap("cor", method = "spearman"))) +
  theme_bw(base_size = 12)
```

## Stranded expression score
```{r}
#all combined, 5kb window, stranded
set.seed(20240331)
min_ints_window = 3
N_bootstrap = 5000
BinSize = 500
CL = "34_mChpos"
all_exp = unique(tib_filtered_mTurq_vals$experiment)
tib_filtered_mTurq_vals_comb = tib_filtered_mTurq_vals %>% left_join(sorting_stats_tib_repcor) %>%
  filter(cell_line %in% c("23_34A", "C9_mCh_34", "C10_34")) %>%
  mutate(cell_line = "34_mChpos")


# # calculate running mean on random sample of same size, 5000 times
# expr_score_tib = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh_stranded(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 10,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = CL, EXP = all_exp)},
#                                             TIB = tib_filtered_mTurq_vals_comb,
#                                             N_resamples = N_bootstrap)
# saveRDS(expr_score_tib, "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250401_new_expression_score_34_mChpos_5kb_stranded.rds")

expr_score_tib = readRDS("/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250401_new_expression_score_34_mChpos_5kb_stranded.rds")



expr_score_sum_tib = bootstrap_summary_fun_str(expr_score_tib,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real =
  expr_score_function_new_weigh_stranded(tib_filtered_mTurq_vals_comb,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = CL,
                                EXP = all_exp)
  
expr_score_sum_tib_filt_comb_str = left_join(expr_score_sum_tib,  expr_score_real) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA)) %>%
  mutate(cell_line = CL)
  

```


```{r S2I_expr_score_stranded_comb, fig.width = 8.5, fig.height = 5}
expr_plot_prange_zoom_str = ggplot(expr_score_sum_tib_filt_comb_str, aes(x = mid_interval, col = strand, fill = strand, linetype = strand)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower, ymax = expr_score_window_upper), alpha = 0.2, linewidth = NA) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median)) +

  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'bottom') +
  colScale2_blue +
  fillScale2_blue
expr_plot_prange_zoom_str
```


# FACS data hopping
Load E2555, for C9_mCh_34 gates and comparison of untreated C9_mCh_34, 23_34A and F121/9, WT.
## load FACSDiva file E2555
```{r}
library(openCyto) 
library(CytoML)
library(flowWorkspace)


diva_ws <- open_diva_xml(diva_xml_file_E2555_sorting)
  

gs <- diva_to_gatingset(diva_ws,
                        name = 1, #the group to import
                        path = path_FACS_sorting_E2555,
                        #swapped columns are a known diva 'bug', for some files you need to 'unswap' them, default is T (as necessary for this file)
                        # swap_cols = F, 
                        worksheet = "global",
                        verbose = F,
                        execute = T)

pdata_tib = pData(gs) %>%
  as_tibble() %>%
  mutate(name_short = str_remove(str_remove(name, "ME2024050._E2534_rep._"), "_([0-9]){3}.fcs"),
         replicate = str_extract(name_short, "rep.$"),
         cell_line = case_when(grepl("WT", name_short) ~ "WT",
                               grepl("F121_9", name_short) ~ "F121_9",
                               grepl("23_34A", name_short) ~ "23_34A",
                               .default = str_extract(name_short, "C9_.._mCh.")),
         
         treatment = case_when(grepl("PB", name_short) ~ "PB",
                               grepl("SB", name_short) ~ "SB",
                               .default = "untreated"),
         recording = case_when(grepl("_SORT", name_short) ~ "sorting",
                               .default = "pre-recording"),
         sample_type = case_when(cell_line %in% c("WT", "F121_9", "23_34A") | treatment == "untreated" ~ "control",
                                 .default = "sample")
  ) %>%
  mutate(construct = case_when(cell_line %in% c("WT", "F121_9") ~ "none",
                                cell_line == c("23_34A") ~ "34",
                                .default = str_extract(cell_line, "([0-9]){2}")),
         mCh_status = case_when(cell_line == "WT" ~ "mCh-",
                                cell_line %in% c("23_34A", "F121_9") ~ "mCh+",
                                .default = str_extract(cell_line, "mCh.")))


pdata_df = as.data.frame(pdata_tib)
rownames(pdata_df) = pdata_df$name

pData(gs) = pdata_df
```

### Density plots untreated samples
The gating set, ggcyto and cowplot don't work together nicely.
Fortifying to a flowSet doesn't help, because then I can't inverse transform the axis.
So instead I just extract all the data as a tibble and plot those.

```{r S2A_density_reporter_cell_lines, fig.width = 10, fig.height=4}
gs_oi_ctrls = subset(gs, sample_type == "control"  & cell_line %in% c("WT", "F121_9", "23_34A", "C9_34_mCh+")) #combine pre-recording & sorting to have as much data as possible

colscale_34 = scale_colour_manual(values = c(`34` = '#0077BB', none = 'grey60'), aesthetics = c("fill", "colour"))


tib_expr_val_ls = lapply(1:4, function(cl){
  data_pop = gh_pop_get_data(gs_oi_ctrls[[cl]], "P4", inverse.transform = T)
  as_tibble(flowCore::exprs(data_pop))
})

names(tib_expr_val_ls) = c(pData(gs_oi_ctrls)$name)
tib_expr_val = bind_rows(tib_expr_val_ls, .id = "name")
tib_expr_val = left_join(tib_expr_val, pData(gs_oi_ctrls))

untr_GFP = ggplot(tib_expr_val, aes(x = `BL[B] 530/30-A`, fill = construct)) +
  geom_density_ridges(aes(y = factor(cell_line, 
                                     levels = c("WT", "F121_9", "23_34A", "C9_34_mCh+"))),
                      linewidth = 0.5,
                      scale = 1.5,
                      alpha = 1,
                      quantile_lines = T, quantiles = 2) +
  theme_bw() +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme(legend.position = 'none') +
  labs(y = "cell line", x = "Sox2::eGFP")+
  colscale_34

untr_mCh = ggplot(tib_expr_val, aes(x = `YG[D] 610/20-A`, fill = construct)) +
  geom_density_ridges(aes(y = factor(cell_line, 
                                     levels = c("WT", "F121_9", "23_34A", "C9_34_mCh+"))),
                      linewidth = 0.5, 
                      scale = 1.5,
                      alpha = 1,
                      quantile_lines = T, quantiles = 2) + 
  theme_bw() +
   ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme(legend.position = 'none') +
  labs(y = "cell line", x = "Sox2::mCherry")+
  colscale_34

untr_mT = ggplot(tib_expr_val, aes(x = `V[F] 450/50-A`, fill = construct)) +
  geom_density_ridges(aes(y = factor(cell_line, 
                                     levels = c("WT", "F121_9", "23_34A", "C9_34_mCh+"))),
                      linewidth = 0.5,
                      scale = 1.5,
                      alpha = 1,
                      quantile_lines = T, quantiles = 2) + 
  theme_bw() +
   ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  theme(legend.position = 'none') +
  labs(y = "cell line", x = "reporter (mTurq)")+
  colscale_34


cowplot::plot_grid(plotlist = list(untr_GFP, untr_mCh, untr_mT),
                   align = 'h', nrow = 1, axis = 'tb')
```

### plot gates sorting
```{r, 2F_plot_gates_mChpos_C9, height = 6, width = 6}
gs_oi_mChpos = subset(gs, sample_type == "sample"  & mCh_status == "mCh+" & cell_line == "C9_34_mCh+" & 
                        (recording == "sorting" | treatment == "PB")) #plot the recording only, combining 2 data sets doesn't give the right percentages on the gates

ggcyto::ggcyto(gs_oi_mChpos, aes(x = "GFP", y = "mTurqoise2"), subset = "mCh_pos") +
  #facet by cell line / treatment
  facet_grid(cell_line ~ treatment) +
  
  #plot cells
  geom_hex(bins = 128, aes( fill = after_stat(ncount))) +
  scale_fill_distiller(palette = 'Spectral') + #manually adding a fill scale also forces the scale to be linear again.
  
  #plot_gates
  ggcyto::geom_gate(paste0("P", 1:6, "_mCh+"), col = 'black') +
  ggcyto::geom_stats(adjust = c(-0.3, 0.5), digits = 2, size = 3) +
    
  #layout
  theme_bw(base_size = 14) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.position = "none",
        aspect.ratio = 1) +
  ggtitle(element_blank())+
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  ggcyto::ggcyto_par_set(limits = list(x = c(1.2, 4.5), y = c(0, 4.2))) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0))+
  ggcyto::labs_cyto(labels = "marker")

```


## load FACSDiva file E2096
```{r}



diva_ws_E2096 <- open_diva_xml(diva_xml_file_E2096_sorting)
  

gs_E2096 <- diva_to_gatingset(diva_ws_E2096,
                        name = 1, #the group to import
                        path = path_FACS_sorting_E2096,
                        swap_cols = F, #apparently we don't need to swap the FSC/SSC-H/W here (this is a known diva 'bug', if you import the xml into flowjo the columns are swapped but apparently not here)
                        worksheet = "global",
                        verbose = F,
                        execute = T)

pdata_tib_E2096 = pData(gs_E2096) %>%
  as_tibble() %>%
  mutate(name_short = str_remove(str_remove(name, "Mathias (BvS) 20221117_mTurq_"), "_([0-9]){3}.fcs"),
         cell_line = case_when(grepl("neg_ctrl", name_short) ~ "WT",
                               grepl("F121,2f,9 parental", name_short) ~ "F121_9",
                               grepl("23_34", name_short) ~ "23_34A",
                               grepl("23_A1_34", name_short) ~ "A1_34"),
         treatment = case_when(grepl("PB", name_short) ~ "PB",
                               grepl("SB", name_short) ~ "SB",
                               .default = "untreated"),
         recording = case_when(grepl("_sort", name_short) ~ "sorting",
                               .default = "pre-recording"),
         sample_type = case_when(cell_line %in% c("WT", "F121_9") | treatment == "untreated" ~ "control",
                                 .default = "sample")
  ) 


pdata_tib_E2096_df = as.data.frame(pdata_tib_E2096)
rownames(pdata_tib_E2096_df) = pdata_tib_E2096_df$name

pData(gs_E2096) = pdata_tib_E2096_df

```

### plot gates sorting

```{r, 2B_plot_gates_mChpos_23_34A, height = 6, width = 6}
gs_oi_E2096 = subset(gs_E2096, sample_type == "sample"  & cell_line == "23_34A"  & 
                        (recording == "sorting" | treatment == "PB")) #recording only, combining 2 data sets doesn't give the right percentages on the gates

ggcyto::ggcyto(gs_oi_E2096, aes(x = "GFP", y = "mTurqoise2"), subset = "P4") + #P4 is just the single cells, don't confuse with 'gate4'
  #facet by cell line / treatment
  facet_grid(cell_line ~ treatment) +

  #plot cells
  geom_hex(bins = 128, aes( fill = after_stat(ncount))) +
  scale_fill_distiller(palette = 'Spectral') + #manually adding a fill scale also forces the scale to be linear again.

  #plot_gates
  ggcyto::geom_gate(paste0("gate", 1:6), col = 'black') +
  ggcyto::geom_stats(adjust = c(-0.3, 0.5), digits = 2, size = 3) +

  #layout
  theme_bw(base_size = 14) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.position = "none",
        aspect.ratio = 1) +
  ggtitle(element_blank())+
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  ggcyto::ggcyto_par_set(limits = list(x = c(1.2, 4.5), y = c(0, 4.2))) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0))+
  ggcyto::labs_cyto(labels = "marker")

```

## load FACSDiva file E2687
```{r}

diva_ws_E2687 <- open_diva_xml(diva_xml_file_E2687_round2_sorting)
  

gs_E2687 <- diva_to_gatingset(diva_ws_E2687,
                        name = 1, #the group to import
                        path = path_FACS_sorting_E2687_round2,
                        swap_cols = F, #apparently we don't need to swap the FSC/SSC-H/W here (this is a known diva 'bug', if you import the xml into flowjo the columns are swapped but apparently not here)
                        worksheet = "global",
                        verbose = F,
                        execute = T)

pdata_tib_E2687 = pData(gs_E2687) %>%
  as_tibble() %>%
  mutate(name_short = str_remove(str_remove(name, "me20241031_E2687_round2_"), "_([0-9]){3}.fcs"),
         cell_line = case_when(grepl("WT_ctrl", name_short) ~ "WT",
                               grepl("CM2285", name_short) ~ "F121_9",
                               grepl("CM2267", name_short) ~ "23_34A_delB_mCh-",
                               grepl("CM2268", name_short) ~ " 23_34A_delB_GFP-",
                               grepl("CM2350", name_short) ~ "F121_9",
                               grepl("23_34", name_short) ~ "23_34A",
                               grepl("C9_34_mCh-", name_short) ~ "C9_mCh-_34",
                               grepl("C9_34_P6", name_short) ~ "P6_mCh-_34",
                               grepl("C9_41_P6", name_short) ~ "P6_mCh-_41",
                               grepl("C10_P7", name_short) ~ "C10_34"),
         treatment = case_when(grepl("PB", name_short) ~ "PB",
                               grepl("SB", name_short) ~ "SB",
                               grepl("Martijn", name_short) ~ "other_experiment",
                               .default = "untreated"),
         recording = case_when(grepl("_SORT", name_short) ~ "sorting",
                               .default = "pre-recording"),
         sample_type = case_when(cell_line %in% c("WT", "F121_9") | treatment == "untreated" ~ "control",
                                 .default = "sample")
  ) 


pdata_tib_E2687_df = as.data.frame(pdata_tib_E2687)
rownames(pdata_tib_E2687_df) = pdata_tib_E2687_df$name

pData(gs_E2687) = pdata_tib_E2687_df

```
### check for cells
First check that we can see cells:
It seems arbitrary whether or not FSC-H/W and SSC-H/W are swapped in the FACSdiva file. So best approach is to plot them here, and if the gate is wrong, turn off/on the swap_cols parameter
```{r}
ggcyto::ggcyto(gs_E2687, aes(x = "FSC-A", y = "FSC-H"), subset = "P4") +
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  geom_hex(bins = 126, aes( fill = after_stat(ncount))) +
  ggcyto::geom_gate("P3", col = 'black') +
  theme_classic()

```
I see cells for P4 (= all single cells) so the columns are set up correctly

### plot gates sorting

```{r, S2D_plot_gates_mChpos_C10_34, fig.height = 6, fig.width = 6}
gs_oi_E2687 = subset(gs_E2687, sample_type == "sample"  & cell_line == "C10_34"  & 
                        (recording == "sorting" | treatment == "PB")) #sorting only, combining 2 data sets doesn't give the right percentages on the gates

ggcyto::ggcyto(gs_oi_E2687, aes(x = "GFP", y = "mTurqoise2"), subset = "mCh_pos") + 
  #facet by cell line / treatment
  facet_grid(cell_line ~ treatment) +

  #plot cells
  geom_hex(bins = 128, aes( fill = after_stat(ncount))) +
  scale_fill_distiller(palette = 'Spectral') + #manually adding a fill scale also forces the scale to be linear again.

  #plot_gates
  ggcyto::geom_gate(paste0("P", 1:6, "_mCh+"), col = 'black') +
  ggcyto::geom_stats(adjust = c(-0.3, 0.5), digits = 2, size = 3) +

  #layout
  theme_bw(base_size = 14) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.position = "none",
        aspect.ratio = 1) +
  ggtitle(element_blank())+
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  ggcyto::ggcyto_par_set(limits = list(x = c(1.2, 4.5), y = c(0, 4.2))) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0))+
  ggcyto::labs_cyto(labels = "marker")

```

# Clonal data
## Load annotation
We only use the insertion clones, so only use those

```{r}
#load the insertion data & the link with the wells 
top_insertion_clones = readRDS(path_insertion_clones_E2221)
link_well_clonename = read_tsv(path_link_well_clonename_E2221)

## Link location - FACS data
link_well_clonename_tib = link_well_clonename %>%
  mutate(sample = paste0("clone_", `clone name`,"_34"),
         plate_name =split_string(Plate_well_new, "_", 1),
         plate_number = case_when(plate_name == "plate1" ~ 1,
                                  plate_name == "plate25" ~ 2,
                                  plate_name == "plate3645" ~ 3)
         ) %>%
  dplyr::rename('well_ID'= `Well index culture plate`) %>%
  rename(clone_name = `clone name`)

linked_clones = left_join(top_insertion_clones, link_well_clonename_tib) 

``` 

## Load data 
Load all FCS files
```{r load_set_of_fcs_files_platereader}
all_fcs_files = dir(path_fcs_files_E2221, full.names = T)
names(all_fcs_files) = split_string(str_remove(all_fcs_files, '.*/export_'), "_", 1,2)

fcs_files_tib = tibble(filename = all_fcs_files, 
                       Plate_well_new = names(all_fcs_files))

fcs_files_sel = left_join(linked_clones, fcs_files_tib) %>%
  filter(!is.na(filename)) %>%#remove the clone we didn't map
  mutate(clone_name = paste0("clone_", clone_name))  

#make dataframe
fcs_files_sel_df = data.frame(fcs_files_sel)
rownames(fcs_files_sel_df) = fcs_files_sel_df$clone_name

#load data
flowset_plates = read.flowSet(files = fcs_files_sel_df$filename, alter.names = T, truncate_max_range = F)
flowCore::sampleNames(flowset_plates) = fcs_files_sel_df$clone_name
flowCore::pData(flowset_plates) = fcs_files_sel_df


flowset_fluo = flowset_plates[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A')]
colnames(flowset_fluo) = c("GFP", "mCherry", "mTurq")


```

## Normalization

calculate median
```{r}
#calculate central tendency
median_exprs = t(sapply(1:length(flowset_fluo), function(i){
  apply(exprs(flowset_fluo[[i]]),2,median)
}))
rownames(median_exprs) <- sampleNames(flowset_fluo)
colnames(median_exprs) <- paste0("median_", colnames(exprs(flowset_fluo[[1]])))

  
#pull out unhopped clones
unhopped_clones = fcs_files_sel %>% filter(start == 34643961) %>% pull(clone_name)

median_exprs[unhopped_clones,]

``` 

### With close WT control
We don't have a WT control for this clonal data. 
The unhopped reporters from the clonal data (measured on 20230228) are most similar to the measurements from E2250 (date 20230418), so we will use the WT control from that date: ME20230418_E2250_pools_WT

```{r}
flowset_WT_ctrl_E2250 = flowCore::read.flowSet(files = path_WT_ctrl_E2250, alter.names = T, truncate_max_range = F)[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A', 'FSC.A', 'SSC.A')]
colnames(flowset_WT_ctrl_E2250) = c("GFP", "mCherry", "mTurq",'FSC', 'SSC')

autofluo_est = tibble(experiment = "E2250",
                      date = "20230418",
                      GFP_WT = median(exprs(flowset_WT_ctrl_E2250[[1]])[,'GFP']),
                      mCherry_WT = median(exprs(flowset_WT_ctrl_E2250[[1]])[,'mCherry']),
                      mTurq_WT = median(exprs(flowset_WT_ctrl_E2250[[1]])[,'mTurq']))

``` 

```{r}
median_exprs_tib = as_tibble(median_exprs) %>%
  mutate(clone_name = rownames(median_exprs))

fcs_tib_clones = fcs_files_sel %>%
  left_join(median_exprs_tib) %>%
  mutate(hopped = start != 34643961) %>%
  mutate(median_mTurq_cor = median_mTurq - autofluo_est$mTurq_WT,
         median_mCherry_cor = median_mCherry - autofluo_est$mCherry_WT,
         median_GFP_cor = median_GFP - autofluo_est$GFP_WT) %>%
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)


#normalize to 23_34A (to compare between measurement days)
standard_tib = fcs_tib_clones %>%
  filter(! hopped) %>%
  group_by(start) %>%
  summarize(
    #GFP normalized
    mTurq_norm_standard = mean(mTurq_norm),
    mCherry_norm_standard = mean(mCherry_norm),
    #not GFP normalized
    GFP_standard = mean(median_GFP_cor),
    mCherry_standard = mean(median_mCherry_cor))

expression_tib_cor = fcs_tib_clones %>%
  mutate(
    #GFP normalized
    rel_norm_mTurq = mTurq_norm / standard_tib$mTurq_norm_standard,
    rel_norm_mCherry = mCherry_norm / standard_tib$mCherry_norm_standard,
    #not GFP normalized
    rel_GFP_cor = median_GFP_cor / standard_tib$GFP_standard,
    rel_mCherry_cor = median_mCherry_cor / standard_tib$mCherry_standard
         )

``` 

## Plot relative expression, with WT correction

### mTurq
```{r}
expression_tib_cor = expression_tib_cor %>%
  #group the regions
  mutate(region = case_when(start <= 34.66E6 & start >= 34.64E6  ~ "around Sox2",
                            start <= 34.9E6 & start >= prange_SCR[1]  ~ "around SCR")) %>%
  mutate(region = factor(region, levels = c("around Sox2", "around SCR")))


``` 

Compute the geom_smooth from the mean per location (so the smooth matches the points actually in the plot).
That is computed here:

```{r}
ins_for_plot = filter(expression_tib_cor, start >= prange_zoom_paper[1] & start <= prange_zoom_paper[2] & chr == "chr3" & observed_clonetype_mapping == "insertion")


ins_for_plot_mean = ins_for_plot %>% 
  group_by(chr, start, strand, region) %>%
  summarize(rel_norm_mTurq = mean(rel_norm_mTurq),
            rel_norm_mCherry = mean(rel_norm_mCherry))

clones_norm_expr_facet_cor2 = ggplot(ins_for_plot_mean,
                                aes(x = start, y = rel_norm_mTurq)) +
  facet_grid(~ region, space = "free_x", scales = "free_x") +
  # # annotate enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red') +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') +
    #annotate landing pad
  geom_vline(xintercept = start(landingPad_23), col = 'darkgrey') +
  geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted') +
  #layout
  labs(x = 'Genomic coordinate (Mb)',
       y = 'relative mTurq') +
  theme_classic(base_size = 16) +
  facetted_pos_scales(
    x = list(
      region == "around Sox2" ~ scale_x_continuous(labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                                                   limits = c(34.64E6, 34.66E6),
                                                   breaks = c(34.64E6,34.65E6, 34.66E6)),
      region == "around SCR" ~ scale_x_continuous(labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                                                  limits = c(34.735E6, 34.775E6))
    )) +
  scale_y_continuous(breaks = scales::pretty_breaks(n=6), 
                     limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.05))) +
  geom_hline(yintercept = 1, linetype = 'dashed')+
  
  #data
  geom_point(aes(shape = strand), col = 'black', size = 2) +
  # stat_summary(fun = mean, col = 'black', aes(shape = strand)) +
  #smooth
  geom_smooth(data = filter(ins_for_plot_mean, region == "around SCR"), aes(x = start, y = rel_norm_mTurq), col = 'grey30')

``` 


## Plot clones with expression score

Annotate clones plot with expression score (from combined data). 5 kb bin size, is also used in the main figure, so no need to re-calculate
```{r}
#plot regions we have clones for (in facets)
expr_score_sum_tib_comb_facet = expr_score_sum_tib_filt_comb %>% 
  mutate(region = case_when(mid_interval <= 34.66E6 & mid_interval >= 34.64E6  ~ "around Sox2",
                            mid_interval <= 34.9E6 & mid_interval >= prange_SCR[1]  ~ "around SCR")) %>%
  mutate(region = factor(region, levels = c("around Sox2", "around SCR"))) %>%
  filter(!is.na(region))

expr_score_region_clones = ggplot(expr_score_sum_tib_comb_facet, aes(x = mid_interval)) +
  facet_grid(. ~ region, space = "free_x", scales = "free_x") +
  
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_C9), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')) +

    ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower, ymax = expr_score_window_upper), alpha = 0.2) +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median)) +

  #axes
  facetted_pos_scales(
    x = list(
      region == "around Sox2" ~ scale_x_continuous(labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                                                   limits = c(34.64E6, 34.66E6),
                                                   breaks = c(34.64E6,34.65E6, 34.66E6)),
      region == "around SCR" ~ scale_x_continuous(labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                                                  # limits = c(start(SCR_50kb), end(SCR_50kb)))
                                                  limits = c(34.735E6, 34.775E6))
    )) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=4),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') + 
  theme_classic(base_size = 16) +
  theme(legend.position = 'bottom') 

```



```{r S2G_insertion_clones_with_expr_score, fig.width= 8 , fig.height= 5}

expr_score_region_clones_comb_noX = expr_score_region_clones +
    theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()
        )
clones_expr_score_plot = cowplot::plot_grid(expr_score_region_clones_comb_noX,
                                            clones_norm_expr_facet_cor2,
                   ncol = 1, align="v", axis="lr", rel_heights = c(0.9, 1.1)) #, rel_heights = c(90, 110))

clones_expr_score_plot
```



# Figure 5, clones
In figure 5 we need 1 panel based on the clones, I generate it here so we don't have to copy so much code.

### mCherry 
```{r plot_mCherry_clones}

clones_norm_expr_mCh_facet_cor = ggplot(ins_for_plot_mean,
                                aes(x = start, y = rel_norm_mCherry)) +
  facet_grid(~ region, space = "free_x", scales = "free_x") +
  # # annotate enhancer
  annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown) +
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown) +
  # annotate gene
  annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red') +
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') +
    #annotate landing pad
  geom_vline(xintercept = start(landingPad_23), col = 'darkgrey') +
  geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted') +
  #layout
  labs(x = 'Genomic coordinate (Mb)',
       y = 'relative mCherry') +
  theme_classic(base_size = 16) +
  facetted_pos_scales(
    x = list(
      region == "around Sox2" ~ scale_x_continuous(labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                                                   limits = c(34.64E6, 34.66E6),
                                                   breaks = c(34.64E6,34.65E6, 34.66E6)),
      region == "around SCR" ~ scale_x_continuous(labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                                                  limits = c(34.735E6, 34.775E6))
    )) +
  scale_y_continuous(breaks = scales::pretty_breaks(n=6), 
                     limits = c(0, NA),
                     expand = expansion(mult = c(0, 0.05))) +
  geom_hline(yintercept = 1, linetype = 'dashed')+
  
  #data
  # geom_point(aes(shape = strand, text = sample, text2 = Plate_well_new), col = 'darkgrey', stroke = NA) +
  stat_summary(fun = mean, col = 'black', aes(shape = strand)) +
  #smooth
  geom_smooth(data = filter(ins_for_plot, region == "around SCR"), aes(x = start, y = rel_norm_mCherry), col = 'grey30')

```

```{r S6A_clones_rel_mT_mCh, fig.width = 8, fig.height = 6}

clones_norm_expr_facet_cor_noX = clones_norm_expr_facet_cor2 +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank()
        )

clones_mT_mCh_comb= cowplot::plot_grid(clones_norm_expr_facet_cor_noX, clones_norm_expr_mCh_facet_cor,
                   ncol = 1, align="v", axis="lr", rel_heights = c(90, 110))

clones_mT_mCh_comb
```

Minimum level of mCherry (according to the smoothened data) is 0.75:
```{r}
clones_norm_expr_mCh_facet_cor + 
  geom_hline(yintercept = 0.75)
```

# Relative expression qPCR
For figure 5 (CDS) we generated qPCR data for an RNA degradation timecourse.
We can use t0 (untreated) of this data to estimate the relative steady-state RNA level of the reporter compared to Sox2 GFP/mCherry
The timecourse has 4 replicates. We also have 1 biological replicate of just RNA isolation - RT - qPCR on several cell lines. We will use this as well (for -116kb_mCh+_34)

Load the data
```{r}

path_qPCR_RNA_stability = "/DATA/projects/Sox2/Figure_Sox2_CDS/me20241217_E2682_all_rep_combined.xlsx"
tib_qPCR_raw_t0 = read_xlsx(path_qPCR_RNA_stability ) %>%
  filter(time == 0) %>%
  mutate(Cell_line = case_when(Cell_line == "SoxP" ~ "Sox2P",
                               Cell_line == "SoxP CDS" ~ "Sox2P_CDS",
                               Cell_line == "SoxP UTR" ~ "Sox2P_CDS_UTR",
                               ))

tib_qPCR_cell_lines = read_tsv("/DATA/usr/m.eder/projects/qPCR_data/E2682_RNA_stability/me20250110_cell_lines.tsv")
tib_rep5 = tib_qPCR_cell_lines %>% 
  filter(construct == "34" & landingPad == "LP_-116kb" & status_mCh == "mCh+") %>%
  mutate(Cell_line = "Sox2P",
         time = 0,#change T0 to 0
         rep = 'rep5',
         target = case_when(target == "cherry" ~ "mCherry",
                            target == "sox" ~ "sox2",
                            .default = target
                            )) %>% 
  select(Cell_line, rep,target, Cq_1, Cq_2, Cq_3, time, date)


tib_qPCR_23_34A_all_raw = bind_rows(tib_qPCR_raw_t0, tib_rep5) %>% 
  dplyr::rename(cell_line = Cell_line) %>%
  filter(cell_line == "Sox2P")
```

average the technical replicates
```{r}
tib_qPCR_23_34A_all = tib_qPCR_23_34A_all_raw %>%
# 1) Summarize mean and sd
  group_by(cell_line , target, rep) %>%
  summarise(
    mean_Ct = mean(c(Cq_1, Cq_2, Cq_3), na.rm = TRUE),
    sd_Ct   = sd(c(Cq_1, Cq_2, Cq_3), na.rm = TRUE),
    .groups = "drop"
  ) 


```

Calculate delta_ct for GFP/mCh vs mT
First take mean of delta_GFP and delta_mCh per replicate, then combine biological replicates
```{r}
fc_mT_23_34A = tib_qPCR_23_34A_all %>%
  select(-sd_Ct)%>%
  pivot_wider(names_from=target, values_from = mean_Ct) %>%
  rowwise()%>%
  mutate(delta_mT_GFP = mTurq - gfp,
         delta_mT_mCh = mTurq - mCherry,
         fc_mT_GFP = 2^-delta_mT_GFP,
         fc_mT_mCh = 2^-delta_mT_mCh,
         #mean of delta mT_GFP and mT_mCh
         delta_merged =  mean(c_across(c(delta_mT_GFP, delta_mT_mCh))),
         fc_merged = 2^-delta_merged) %>%
  ungroup()
         
fc_mT_23_34A %>%
  group_by(cell_line) %>%
  summarise(mean_delta_merged = mean(delta_merged),
         sd_delta_merged = sd(delta_merged)) %>%
  mutate(mean_fc = 2^-mean_delta_merged,
         #calculate confidence interval with +- 1.96*SD
         fc_min_95 = 2^-(mean_delta_merged+1.96*sd_delta_merged),
         fc_max_95 = 2^-(mean_delta_merged-1.96*sd_delta_merged),
         )
  
```
So relative mTurq is 9.7% (95%CI=[6.1%-15.5%])


# FACS spillover
On the FACS Aria Fusion2 sorter we used compensation. 
For the analytical FACS we used no compensation. 
We want to check that there is no spillover of mTurq into the GFP channel on the analytical FACS. 



## Load data
Get the file names and annotate them
```{r load_set_of_fcs_files}

##load the file names
fcs_files_E2762 = dir(path_fcs_files_E2762, full.names = T)

#wrangle sample names E2762
names(fcs_files_E2762) = split_string(str_remove(fcs_files_E2762, '.*/'), "_", 4)

pdata_tib_E2762 = tibble(files = fcs_files_E2762, sample_name = names(fcs_files_E2762)) %>%
  mutate(cell_line = case_when(sample_name == "CM1874" ~ "WT",
                               sample_name == "CM2167" ~ "23_34A",
                               sample_name == "CM2267" ~ "23_34A_mCh-",
                               sample_name == "CM2292" ~ "23_41",
                               sample_name == "CM2293" ~ "23_41_mCh-",
                               sample_name == "CM2268" ~ "23_34_GFP-",
                               sample_name == "CM2431" ~ "F121"),
         fluo_status = case_when(cell_line == "WT" ~ "none",
                               cell_line == "23_34A" ~ "GFP+_mCh+_mT+",
                               cell_line == "23_34A_mCh-" ~ "GFP+_mCh-_mT+",
                               cell_line == "23_41" ~ "GFP+_mCh-_mT+",
                               cell_line == "23_41_mCh-" ~ "GFP+_mCh-_mT+",
                               cell_line == "23_34_GFP-" ~ "GFP-_mCh+_mT+",
                               cell_line == "F121" ~ "GFP+_mCh+_mT-"),
           deletion = grepl("mCh-|GFP-", cell_line))
```

load the fcs files
```{r}
#load data
pdata_df = data.frame(pdata_tib_E2762)
rownames(pdata_df) = pdata_df$sample_name


flowset = flowCore::read.flowSet(files = pdata_tib_E2762$files, 
                                       alter.names = T, truncate_max_range = F,
                                 column.pattern = 'BL.B..530_30.A|YG.D..610_20.A|V.F..450_50.A|FSC.A|SSC.A')
flowCore::sampleNames(flowset) = pdata_tib_E2762$sample_name
pData(flowset) = pdata_df


total_cell_count = as.numeric(flowCore::keyword(flowset, "$TOT")[,'$TOT'])
pData(flowset)$cell_count = total_cell_count

flowset_fluo = flowset[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A')]
colnames(flowset_fluo) = c("GFP", "mCherry", "mTurq")
rm(flowset)
```

## Plot histograms
```{r}
# Filter relevant cell lines
flowset_fluo_idx = pData(flowset_fluo)$cell_line %in% c("WT", "F121", "23_34A_mCh-", "23_34_GFP-")

pFluo_GFP = 
ggplot(flowset_fluo[flowset_fluo_idx], aes(x = GFP)) +
  geom_density_ridges(aes(y =  fluo_status),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 1.5,
                      alpha = 0.75,
                      quantile_lines = TRUE,
                      fill = 'green3') +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  scale_y_discrete(expand = expansion( add = c(0, 2))) +
  theme_bw() +
  theme(legend.position = 'none',
        axis.title.y = element_blank())


pFluo_mTurq = 
ggplot(flowset_fluo[flowset_fluo_idx], aes(x = mTurq)) +
  geom_density_ridges(aes(y =  fluo_status),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 1.5,
                      alpha = 0.75,
                      quantile_lines = TRUE,
                      fill = 'skyblue') +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  scale_y_discrete(expand = expansion( add = c(0, 2))) +
  theme_bw() +
  theme(legend.position = 'none',
        axis.title.y = element_blank())




plot_grid(pFluo_GFP, pFluo_mTurq, ncol = 2)
```
