---
title: "CM20250514_Figure4_Sox2_CRISPR"
author: "Mathias & Christine"
date: "2025-04-01"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---


# Goal 
Plot the figure panels relating to the CRISPR deletion of Sox2, from the sorting data as well as the sorted pools & clones. 


## Notes for re-running
To re-run this code, change the paths in 'File paths' to the correct location of datasets.
The file paths of the fcs files are hardcoded in the annotation of the clones (). To re-run the analysis of those clones, adapt the chunk 'load_fcs_data_clones' to point directly to the path you stored these fcs files (paste0(/your/own/path/, fcs_files_E2350$filename)).

# setup
## libraries
```{r load_libraries}

# Load dependencies
library(GenomicRanges)
library(rtracklayer)
library(tidyverse)
# library(dtplyr)
library(dplyr)
# library(tidyr)
library(ggbeeswarm)
library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
library(RColorBrewer)
library(RcppRoll)
# library(plotly)
library(readxl)
library(ggpubr)
library(ggpmisc)
library(ggridges)
library(import)
library(ggh4x)
import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
import::from(ggcyto, 'fortify_fs')
library(flowDensity)
library(flowWorkspace)
library(CytoML)
#to install this you need to manually install RFOC first:
#install_url('https://cran.r-project.org/src/contrib/Archive/RFOC/RFOC_3.4-6.tar.gz')
library(caTools)
library(ggrastr)
library(cowplot)
library(Signac)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}
```

# Paths & parameters
```{r paths_parameters}
#datatag
datetag = paste0("CM",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/Sox2/Figure_Sox2_CRISPR/analysis_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}

library(knitr)
opts_chunk$set(cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/")) 
pdf.options(useDingbats = FALSE)
```

## File paths
```{r}
#CTCF motifs
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"

#Sox2 CRISPR experiment
path_fcs_E2736 = '/DATA/usr/m.eder/projects/FACS_data/E2736/me20250206_E2736_sox2_deletion/singlecells_for_R'

#clones
path_annotation_clones_E2350 = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2350_generating_CRISPR_clones/CM20240221_E2350_fcs_files_tib.rds"

#hopping data
path_tib_23_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-116kb_Sox2P_pools.txt"
path_tib_C9_mChpos_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-161kb_Sox2P_pools.txt"
path_tib_C9_mChneg_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/still_to_upload/mapped_integrations_-161kb_Sox2mCherrydel_Sox2P_pools.txt"

path_median_mTurq_relevant_cell_lines =  "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250318_median_mTurq_sorting_for_paper.tsv"
path_sorting_stats = "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250327_stats_for_expression_score_for_paper.xlsx"

path_expr_score_bootstrap_C9_mCh_34 = "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250327_new_expression_score_sum_filt_C9_mCh_34_5kb.rds"


#sorting data hopping
diva_xml_file = "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2534_hopping_C9_LP/FACS/Mathias (BvS)20240501_E2534_rep1.xml"
path_FACS_sorting_E2555 =  "/DATA/usr/c.moene/projects/Hopping_Sox2_mTurq/E2534_hopping_C9_LP/FACS"


diva_xml_file_E2687_round2_sorting = "/DATA/projects/Sox2/Zenodo_collection/E2687/Mathias (BvS)20241031_E2687_round2.xml"
path_FACS_sorting_E2687_round2 = "/DATA/projects/Sox2/Zenodo_collection/E2687/Mathias (BvS)20241031_E2687_round2"

#ChIP and ATAC data
paths_ChIP_files = c(
  ATAC = '/DATA/usr/c.moene/public_data/ATAC/GSE98390_E14_ATAC_MERGED.DANPOS.bw',
  H3K27ac = '/DATA/usr/c.moene/public_data/ChIP/Histons/H3K27ac_ENCFF583WVZ_signal_pval.bigWig',
  CTCF = '/DATA/projects/Sox2/CTCF_ChIP_analysis/results/bigwig/6764_2_ME_E2_CGATGT_S2_R1_001.bw',
  H3K27me3 = '/DATA/projects/Sox2/ChIP_seq_tracks/histons/H3K27me3_ENCFF595SIA_signal_pval.bigWig',
  Ring1B = '/DATA/projects/Sox2/ChIP_seq_tracks/other/Ring1B_GSE96107.bigWig'
)

#functions
source("/DATA/projects/Sox2/General_functions/CM20250318_new_expression_score_functions.R")

```


# Annotations

gene etc
```{r}
Sox2_gene <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34649995,
                                     end = 34652461),
                    strand = "+")
                    
enhancer <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34753415,
                                     end = 34766401),
                    strand = "*")

#CTCF sites based on our own mapping
CTCF_mm10 <- import.bed(path_CTCF_sites)
CTCF_mm10.chr3 <- CTCF_mm10[seqnames(CTCF_mm10)== 'chr3']
# add the missing site after the SCR
CTCF_mm10.chr3_extra = sort(c(CTCF_mm10.chr3,
                         GRanges(seqnames = "chr3",
                                 IRanges(start = 34772210, end = 34772210),
                                 strand = "+")),
                         ignore.strand = T)


```



# Color setup
```{r}

colors_gates = c("black", "grey60", "grey60", "#1cbdc2")
colscale_gates = scale_colour_manual(name = "gate_for_color", values = colors_gates,
                                     aesthetics = c("color", "fill"))

colors_clones = c(`negative control` = "grey60", untreated = "black", delA = '#067276', delB = '#1cbdc2')
colscale_clones = scale_colour_manual(name = "treatment_for_color", values = colors_clones,
                                     aesthetics = c("color", "fill"))

## blue colors beeswarm (mChpos_34)
myCols_blue = c('#525252', "#080E5B",  "#212B71", "#3A4987", "#465893", "#5A719A", "#6C8AA0")
names(myCols_blue) = c("ctrl", paste0("P", 1:6))
colScale7_blue <- scale_colour_manual(name = "population", values = myCols_blue)
fillScale7_blue <- scale_fill_manual(name = "population", values = myCols_blue)

## pink colors beeswarm (mChneg_34)
myCols_pink = c('#525252', rep("#1cbdc2",3), rep("#18a5aa",3), "#158e92", "#117679", "#0e5f61", "#0a4749")
names(myCols_pink) =  c("ctrl", "P1_H", "P1", "P1_L",
                        "P2_H", "P2", "P2_L",
                        paste0("P", 3:6))

colScale7_pink =  scale_colour_manual(name = "population", values = myCols_pink, aesthetics = c("colour", "fill"))

cols_CL_pop = c(myCols_blue, myCols_pink)
names(cols_CL_pop) = c(paste0("mChpos_34_", names(myCols_blue)),
                       paste0("C9_mCh-_34_" , names(myCols_pink) ))
colScale_pop_CL = scale_colour_manual(name = "CL_population", values = cols_CL_pop)

## SCR
brown = "#ffb000"
```

Exact coordinates landing pads
```{r}

landingPad_23 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34643960,
                                     end = 34643962),
                    strand = "+")

landingPad_C9 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 34598479,
                                     end = 34598480),
                    strand = "+")
landingPad_C10 <- GRanges(seqnames = "chr3", 
                    ranges = IRanges(start = 35478413,
                                     end = 35478414      ),
                    strand = "+")

LP_tib = tibble(landing_pad = c("23", "C9", "C10"), 
                start = c( start(landingPad_23),start(landingPad_C9), start(landingPad_C10)))

```

```{r}
# Plotting ranges
prange_zoom = c(34.5E6, 35E6)
prange_zoom_paper = c(34.59E6, 34.83E6)
prange_SCR = c(34.725E6, 34.785E6)

```

# Load data, sorting

## wrangle E2736
```{r load_E2736}
fcs_files_E2736 = dir(path_fcs_E2736, 
                      full.names = T)

fcs_files_E2736_tib = tibble(file = fcs_files_E2736) %>%
  mutate(sample_name_tmp = split_string(str_remove(file, '.*/'), "_", 4, 7)) %>%
  mutate(cell_line = case_when(grepl("23_34A", sample_name_tmp) ~ "23_34A",
                               grepl("C9_34", sample_name_tmp) ~ "C9_34",
                               grepl("8_348", sample_name_tmp) ~ "8_34",
                               grepl("WT", sample_name_tmp) ~ "WT",
                               grepl("F121", sample_name_tmp) ~ "F121"),
         treatment = case_when(grepl("UT", sample_name_tmp) ~ "untreated",
                               grepl("WT", sample_name_tmp) ~ "untreated",
                               grepl("F121", sample_name_tmp) ~ "untreated",
                               grepl("Tdgf1", sample_name_tmp) ~ "Tdgf1",
                               grepl("up1c", sample_name_tmp) ~ "up1c",
                               grepl("up2c", sample_name_tmp) ~ "up2c",
                               grepl("do1c", sample_name_tmp) ~ "do1c",
                               grepl("DelA", sample_name_tmp) ~ "delA",
                               grepl("DelB", sample_name_tmp) ~ "delB"),
         timepoint = "sorting",
         experiment = "E2736",
         flowcytometer = "sorter",
         sample_name = paste0(experiment, "_", cell_line, "_", treatment),
         date = "me20250128"
         ) %>%
  mutate(fcs_filter = case_when(grepl("Single Cells.fcs", file) ~ "SingleCells",
                                grepl("Single Cells.fcs", file) ~ "SingleCells2")) %>%
  filter(fcs_filter == "SingleCells") #use equivalent filtering to the other samples (only filter for doublets once)

```

## Load data
```{r combine_fcs_tibs}

fcs_annotation_tib = bind_rows(fcs_files_E2736_tib) %>%
  mutate(cell_line = factor(cell_line, c("C9_34", "23_34A", "8_34", "WT", "F121"))) %>%
  mutate(landing_pad = case_when(cell_line == "WT" ~ "none",
                                 cell_line == "F121" ~ "none",
                                 .default = split_string(cell_line, "_", 1))) %>%
  mutate(landing_pad = factor(landing_pad, levels = c("none", "C9", "23", "8"),
                               labels = c("no insert", "-161 kb", "-116 kb", "-39 kb")),
         treatment = factor(treatment, levels  = c("untreated", "delA", "delB", "up1c", "up2c", "do1c", "Tdgf1"),
                            labels = c("untreated", "ΔSox2", "ΔCBS_Sox2", "up1c_single", "up2c_single", "do1c_single", "cutting_ctrl"))) %>%
  select(-c(fcs_filter))
```

```{r load_fcs_files_sorter}
fcs_annotation_df = data.frame(fcs_annotation_tib)
rownames(fcs_annotation_df) = fcs_annotation_df$sample_name

flowset = flowCore::read.flowSet(files = fcs_annotation_df$file, alter.names = T, truncate_max_range = F, 
                                 ignore.text.offset = T, 
                                 column.pattern = 'BL.B..530_30.A|YG.D..610_20.A|V.F..450_50.A|FSC.A|SSC.A'
                                 #samples from sorter have different fluorophores, only take the shared ones
)


flowCore::sampleNames(flowset) = fcs_annotation_df$sample_name
pData(flowset) = fcs_annotation_df

flowset_fluo = flowset[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A')]
colnames(flowset_fluo) = c("GFP", "mCherry", "mTurq")

rm(flowset)

```

```{r preprocess_fcs_data}
total_cell_count_tmp = flowCore::keyword(flowset_fluo, "$TOT")[,'$TOT']
total_cell_count = as.numeric(total_cell_count_tmp)
names(total_cell_count) = names(total_cell_count_tmp)

measurement_date = flowCore::keyword(flowset_fluo, "$DATE")[,'$DATE']

median_mTurq = sapply(1:length(flowset_fluo), function(i){
  median(exprs(flowset_fluo[[i]])[,'mTurq'])
})
names(median_mTurq) = sampleNames(flowset_fluo)

median_mCherry = sapply(1:length(flowset_fluo), function(i){
  median(exprs(flowset_fluo[[i]])[,'mCherry'])
})
names(median_mCherry) = sampleNames(flowset_fluo)


median_GFP = sapply(1:length(flowset_fluo), function(i){
  median(exprs(flowset_fluo[[i]])[,'GFP'])
})
names(median_GFP) = sampleNames(flowset_fluo)

pData(flowset_fluo)$median_mTurq = median_mTurq
pData(flowset_fluo)$median_mCherry = median_mCherry
pData(flowset_fluo)$median_GFP = median_GFP

pData(flowset_fluo)$total_cell_count = total_cell_count
pData(flowset_fluo)$measurement_date = measurement_date #date from the fcs file, must be correct

flowset_fluo_data_tib = tibble(pData(flowset_fluo))

```


# Figure sorting data
```{r cuttoffs}

# Step 1: Extract WT samples from flowset_fluo
flowset_WT_ctrls = flowset_fluo[pData(flowset_fluo)$cell_line == "WT"]

# Step 2: Compute the 99th percentile cutoffs for each marker
mCh_neg_99_ls = lapply(1:length(flowset_WT_ctrls), function(i){
  quantile(exprs(flowset_WT_ctrls[[i]])[,'mCherry'], probs = 0.99, na.rm = TRUE)
})
mCh_neg_99 = mean(unlist(mCh_neg_99_ls)) # Cutoff for mCherry negative

GFP_neg_99_ls = lapply(1:length(flowset_WT_ctrls), function(i){
  quantile(exprs(flowset_WT_ctrls[[i]])[,'GFP'], probs = 0.99, na.rm = TRUE)
})
GFP_neg_99 = mean(unlist(GFP_neg_99_ls)) # Cutoff for GFP negative

mT_neg_99_ls = lapply(1:length(flowset_WT_ctrls), function(i){
  quantile(exprs(flowset_WT_ctrls[[i]])[,'mTurq'], probs = 0.99, na.rm = TRUE)
})
mT_neg_99 = mean(unlist(mT_neg_99_ls)) # Cutoff for mTurq negative

```

## Dotplot example

Use 1.5x the border of the WT sample as a cutoff for the GFP+ and mCh+ gates (so we have some space)

```{r A_GFP_mCh_dotplot_one_example}
dotplot_23_delB_GFP_mCh = ggplot(flowset_fluo[2], aes(x = GFP, y= mCherry, fill = after_stat(ncount))) +#if you want each separate panel to end in the highest color
  facet_nested(landing_pad + treatment ~ date) +
  theme_classic() +
  geom_hex(bins = 256) +
  scale_fill_distiller(palette = 'Spectral') +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  ggcyto::scale_y_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  geom_hline(yintercept = c(mCh_neg_99, mCh_neg_99*1.5)) +
  geom_vline(xintercept = c(GFP_neg_99, GFP_neg_99*1.5)) +
  theme(aspect.ratio = 1,
        legend.position = 'none' )

dotplot_23_delB_GFP_mCh

# ggsave('/DATA/projects/Sox2/Figure_Sox2_CRISPR/CM20250423_FACS_GFP_mCherry_example.pdf' , dotplot_23_delB_GFP_mCh, width = 7, units = 'cm' )
```


## gating the sorting data
Use 1.5x the border of the WT sample as a cutoff for the GFP+ and mCh+ gates (so we have some space)

```{r gate_mCh_GFP_level}

double_posGate <- rectangleGate(filterId="GFP+ mCh+",
                                "GFP"=c(1.5*GFP_neg_99, Inf), "mCherry"=c(1.5*mCh_neg_99, Inf))
mCh_negGate <- rectangleGate(filterId="GFP+ mCh-",
                                "GFP"=c(1.5*GFP_neg_99, Inf), "mCherry"=c(-Inf, mCh_neg_99))
GFP_negGate <- rectangleGate(filterId="GFP- mCh+",
                                "GFP"=c(-Inf, GFP_neg_99), "mCherry"=c(1.5*mCh_neg_99, Inf))

doublepos_fcs = flowCore::Subset(flowset_fluo, double_posGate)
pData(doublepos_fcs)$gate = "doublepos"
sampleNames(doublepos_fcs) = paste0(sampleNames(doublepos_fcs), "_", pData(doublepos_fcs)$gate)

mChneg_fcs = flowCore::Subset(flowset_fluo, mCh_negGate)
pData(mChneg_fcs)$gate = "mChneg"
sampleNames(mChneg_fcs) = paste0(sampleNames(mChneg_fcs), "_", pData(mChneg_fcs)$gate)

GFPneg_fcs = flowCore::Subset(flowset_fluo, GFP_negGate)
pData(GFPneg_fcs)$gate = "GFPneg"
sampleNames(GFPneg_fcs) = paste0(sampleNames(GFPneg_fcs), "_", pData(GFPneg_fcs)$gate)


gated_fcs = rbind2(doublepos_fcs, mChneg_fcs)
gated_fcs = rbind2(gated_fcs, GFPneg_fcs)

cell_count_gate = sapply(1:length(gated_fcs), function(i){
  nrow(gated_fcs[[i]])
})
pData(gated_fcs)$cell_count_gate = cell_count_gate
```

## plot density plots

```{r S4A_Fig_mTurq_ridgeline_gated, fig.width= 10, fig.height= 10, cache=FALSE}
# gate the samples of interest, for untreated the GFPneg and mChneg gates are meaningless
gated_samples_oi = (pData(gated_fcs)$treatment %in% c("ΔSox2", "ΔCBS_Sox2")|  pData(gated_fcs)$gate == "doublepos") & pData(gated_fcs)$cell_line != "WT"

#group the gates to be able to color them
pData(gated_fcs)$gate_for_color = case_when(pData(gated_fcs)$treatment == "untreated" ~ 
                                                  paste0(pData(gated_fcs)$gate, "_untreated"),
                                                .default = pData(gated_fcs)$gate)
pData(gated_fcs)$gate_for_color = factor(pData(gated_fcs)$gate_for_color, 
                                             levels = c("doublepos_untreated","doublepos", "GFPneg", "mChneg"))

#get cell counts per gate
cell_count_tib = 
  tibble(pData(gated_fcs[gated_samples_oi])) %>%
  select(experiment, gate, cell_line,gate_for_color, treatment, cell_count_gate, date)

#regular density plot
dens_plot_active_LP_gated  = ggplot(gated_fcs[gated_samples_oi] , 
       aes(x = mTurq, fill = gate_for_color, col = gate_for_color)) +
  # facet_nested(experiment + treatment ~ cell_line, scale = "free_y") +
  facet_nested(treatment ~ cell_line, scale = "free_y", space = 'free_y') +
  geom_density_ridges(aes(y = fct_reorder(gate_for_color, as.integer(gate_for_color), .desc= T)),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 1.5,
                      alpha = 0.50) + #, 
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5), limits = c(-10^3, 5*10^5),
                               pos = 4.42 ) +
  scale_y_discrete(expand = expansion( add = c(0, 1.7))) +
  geom_text_npc(data = cell_count_tib, aes(label = cell_count_gate, 
                                                npcy = 1.2-(as.integer(gate_for_color)/4), col = gate_for_color), 
                npcx = 0.95, hjust = "inward") +
  theme_bw(base_size = 14) +
  theme(axis.title.y = element_blank(),
        legend.position = 'none') +
  colscale_gates 

dens_plot_active_LP_gated
# ggsave( "/DATA/usr/m.eder/projects/FACS_data/E2736/ME20250401_FACS_density_gates_active_LPs_E2736.pdf", dens_plot_active_LP_gated, width = 20, height = 20, units = "cm")

```

## plot only mChneg population with Tdgf1 ctrl
```{r Fig_4B_mTurq_ridgline}
# Step 1: Filter data for the specific conditions
cell_count_tib_try = 
  tibble(pData(gated_fcs[gated_samples_oi])) %>%
  filter(
    (treatment == "cutting_ctrl" & gate_for_color == "doublepos") |
    (treatment %in% c("ΔSox2", "ΔCBS_Sox2") & gate_for_color == "mChneg")
  ) %>%
  select(experiment, gate, cell_line, gate_for_color, treatment, cell_count_gate, date)

# Step 2: Select gated samples for plotting
gated_samples_oi_try = 
  (pData(gated_fcs)$treatment == "cutting_ctrl" & pData(gated_fcs)$gate_for_color == "doublepos") |
  (pData(gated_fcs)$treatment %in% c("ΔSox2", "ΔCBS_Sox2") & pData(gated_fcs)$gate_for_color == "mChneg")

# Step 3: Verify data selection (optional)
pData(gated_fcs)[gated_samples_oi_try,]

new_flow = gated_fcs[gated_samples_oi_try,]
tib_new = pData(new_flow) %>% mutate(combined = paste0(gate_for_color, "_", treatment))

pData(new_flow) <- tib_new


# Define custom colors
custom_colors <- c(
  "doublepos_cutting_ctrl" = "grey60",   
  "mChneg_ΔSox2" = "#067276",      
  "mChneg_ΔCBS_Sox2" = "#1cbdc2")

# Step 4: Generate the plot
ggplot(new_flow, 
       aes(x = mTurq, fill = combined, col = combined)) +
  facet_nested(. ~ cell_line) +
  geom_density_ridges(aes(y = combined),
                      size = 0.3,
                      scale = 1.5,
                      alpha = 0.50) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, 
                               breaks = c(-10^3, 0, 10^3, 10^4, 10^5), 
                               limits = c(-10^3, 5*10^5),
                               pos = 4.42) +
  scale_y_discrete(expand = expansion(add = c(0, 1.7))) +
  # geom_text_npc(data = tib_new,
  #               aes(label = cell_count_gate,
  #                   npcy = 1.2 - (as.integer(gate_for_color) / 4),
  #                   col = gate_for_color),
  #               npcx = 0.95, hjust = "inward") +
  theme_bw(base_size = 14) +
  theme(axis.title.y = element_blank(),
        legend.position = 'none') +
  scale_colour_manual(values = custom_colors) + # Custom colors for line borders
  scale_fill_manual(values = custom_colors)     # Custom colors for fills

```

```{r S4B_39kb_mCh_GFPneg}

# Step 1: Filter data for the specific conditions
cell_count_tib_try = 
  tibble(pData(gated_fcs[gated_samples_oi])) %>%
  filter( 
    (treatment == "cutting_ctrl" & gate_for_color == "doublepos") |
    (treatment %in% c("ΔSox2", "ΔCBS_Sox2") & gate_for_color == "GFPneg")
  ) %>%
  filter(cell_line == "8_34") %>%
  select(experiment, gate, cell_line, gate_for_color, treatment, cell_count_gate, date)

# Step 2: Select gated samples for plotting
gated_samples_oi_try = 
  (pData(gated_fcs)$treatment == "cutting_ctrl" & pData(gated_fcs)$gate_for_color == "doublepos" & pData(gated_fcs)$cell_line == "8_34") |
  (pData(gated_fcs)$treatment %in% c("ΔSox2", "ΔCBS_Sox2") & pData(gated_fcs)$gate_for_color == "GFPneg" & pData(gated_fcs)$cell_line == "8_34")

# Step 3: Verify data selection (optional)
pData(gated_fcs)[gated_samples_oi_try,]

new_flow = gated_fcs[gated_samples_oi_try,]
tib_new = pData(new_flow) %>% mutate(combined = paste0(gate_for_color, "_", treatment))

pData(new_flow) <- tib_new


# Define custom colors
custom_colors <- c(
  "doublepos_cutting_ctrl" = "grey60",   
  "GFPneg_ΔSox2" = "#067276",      
  "GFPneg_ΔCBS_Sox2" = "#1cbdc2")



# Step 4: Generate the plot
## Plot mCherry
ggplot(new_flow, 
       aes(x = mCherry, fill = combined, col = combined)) +
  facet_nested(. ~ cell_line) +
  geom_density_ridges(aes(y = combined),
                      size = 0.3,
                      scale = 1.5,
                      alpha = 0.50) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, 
                               breaks = c(-10^3, 0, 10^3, 10^4, 10^5), 
                               limits = c(-10^3, 5*10^5),
                               pos = 4.42) +
  scale_y_discrete(expand = expansion(add = c(0, 1.7))) +
  # geom_text_npc(data = tib_new,
  #               aes(label = cell_count_gate,
  #                   npcy = 1.2 - (as.integer(gate_for_color) / 4),
  #                   col = gate_for_color),
  #               npcx = 0.95, hjust = "inward") +
  theme_bw(base_size = 14) +
  theme(axis.title.y = element_blank(),
        legend.position = 'none') +
  scale_colour_manual(values = custom_colors) + # Custom colors for line borders
  scale_fill_manual(values = custom_colors)     # Custom colors for fills



## Plot mTurq
ggplot(new_flow, 
       aes(x = mTurq, fill = combined, col = combined)) +
  facet_nested(. ~ cell_line) +
  geom_density_ridges(aes(y = combined),
                      size = 0.3,
                      scale = 1.5,
                      alpha = 0.50) +
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, 
                               breaks = c(-10^3, 0, 10^3, 10^4, 10^5), 
                               limits = c(-10^3, 5*10^5),
                               pos = 4.42) +
  scale_y_discrete(expand = expansion(add = c(0, 1.7))) +
  # geom_text_npc(data = tib_new,
  #               aes(label = cell_count_gate,
  #                   npcy = 1.2 - (as.integer(gate_for_color) / 4),
  #                   col = gate_for_color),
  #               npcx = 0.95, hjust = "inward") +
  theme_bw(base_size = 14) +
  theme(axis.title.y = element_blank(),
        legend.position = 'none') +
  scale_colour_manual(values = custom_colors) + # Custom colors for line borders
  scale_fill_manual(values = custom_colors)     # Custom colors for fills
```




# Sorted clones

## load CRISPR clones

```{r wrangle_fcs_tib_clones}
fcs_files_E2350 = readRDS(path_annotation_clones_E2350) %>%
  #remove the WT samples that were measured on a different day (labeled here as E2250, but were actually measured in E2254)
  filter(!name %in% c(paste0('E2250_WT_untreated_none_E', 10:12))) %>%

  mutate(flow_experiment = experiment) %>%
  filter(!cell_line == "C10_34" & !cell_line == "C1_34") %>%
  mutate(GFP_state = case_when(sorted_phenotype == "GFP-" ~ "GFP-",
                               cell_line == "WT" ~ "GFP-",
                               .default = "GFP+"),
         mCh_state = case_when(sorted_phenotype %in% c("mCh-_mT-", "mCh-_mT+", "mCh-_NA") ~ "mCh-",
                               cell_line == "WT" ~ "mCh-",
                               .default = "mCh+")) %>%
  mutate(expected_insert = case_when(cell_line %in% c("23_34A", "8_34.8", "C9_34") ~ "34",
                                     .default = "none")) %>%
  mutate(landing_pad = case_when(expected_insert == "34" ~  str_replace(cell_line, "_.*$", ""),
                                 .default = "none")) %>%
  mutate(control_samples = case_when(cell_line %in% c("WT", "F121-9") ~ cell_line,
                                     cell_line == "23_34A" & treatment == "untreated" ~ "23_34A")) %>%
  mutate(type = case_when(type == "pool" ~ "sorted_pool",
                          .default = type)) %>%
  mutate(sorted_phenotype = case_when(sorted_phenotype == "mCh-_NA" ~ "mCh-_mT+", #correct the pools which were wrongly labeled as mCh-_NA
                                      .default = sorted_phenotype)) %>%

  #only include remeasurements, but also from the sorter (for E2263)
  filter(timepoint == "remeasurement") %>%
  mutate(exp_name = paste0(experiment, "_", sample_name)) %>%
  mutate(cell_line_clone_name = case_when(type == "control" ~ cell_line,
                                          type == "clone" ~ paste0(cell_line, "_", treatment, "_", clone_name))) %>% 
  mutate(treatment_for_color = case_when(cell_line %in% c("WT", "F121-9") ~ "negative control",
                                         .default = treatment))%>%
  mutate(cell_line_treatment = paste0(cell_line, "_", treatment)) %>%
  
  #only include the treatments & sortings relevant here:
  filter(treatment %in% c('untreated', 'delA', 'delB')) %>%
  filter(sorted_phenotype %in% c("mCh-_mT+","none") | (cell_line == "C9_34" & sorted_phenotype =="mCh-_mT-")) %>%
  mutate(treatment = factor(treatment, levels = c('untreated', 'delA', 'delB'))) %>%
  #for deletions only include correctly genotyped clones
  filter((is.na(PCR_band_deletion) | PCR_band_deletion) & 
           (is.na(PCR_band_intact_allele) | PCR_band_intact_allele)) %>%
  #add short filename
  mutate(filename = str_remove(file, '^.*/'))
  

```



```{r load_fcs_data_clones}
fcs_annotation_E2350_df = data.frame(fcs_files_E2350)
rownames(fcs_annotation_E2350_df) = fcs_files_E2350$exp_name

flowset = flowCore::read.flowSet(files = fcs_annotation_E2350_df$file, alter.names = T, truncate_max_range = F, 
                                 ignore.text.offset = T, 
                                 column.pattern = 'BL.B..530_30.A|YG.D..610_20.A|V.F..450_50.A|FSC.A|SSC.A'
                                 #samples from sorter have different fluorophores, only take the shared ones
)

## to run this when files are in a different folder (all in one folder, called MyFolder), run:
# flowset = flowCore::read.flowSet(files = paste0(MyFolder, "/", fcs_annotation_E2350_df$filename), alter.names = T, truncate_max_range = F, 
#                                  ignore.text.offset = T, 
#                                  column.pattern = 'BL.B..530_30.A|YG.D..610_20.A|V.F..450_50.A|FSC.A|SSC.A'
#                                  #samples from sorter have different fluorophores, only take the shared ones
# )


flowCore::sampleNames(flowset) = fcs_annotation_E2350_df$exp_name
pData(flowset) = fcs_annotation_E2350_df

flowset_fluo_clones = flowset[,c('BL.B..530_30.A', 'YG.D..610_20.A', 'V.F..450_50.A')]
colnames(flowset_fluo_clones) = c("GFP", "mCherry", "mTurq")

rm(flowset)
```

```{r preprocess_fcs_data_clones}
total_cell_count_tmp = flowCore::keyword(flowset_fluo_clones, "$TOT")[,'$TOT']
total_cell_count = as.numeric(total_cell_count_tmp)
names(total_cell_count) = names(total_cell_count_tmp)

measurement_date = flowCore::keyword(flowset_fluo_clones, "$DATE")[,'$DATE']

median_mTurq = sapply(1:length(flowset_fluo_clones), function(i){
  median(exprs(flowset_fluo_clones[[i]])[,'mTurq'])
})
names(median_mTurq) = sampleNames(flowset_fluo_clones)

median_mCherry = sapply(1:length(flowset_fluo_clones), function(i){
  median(exprs(flowset_fluo_clones[[i]])[,'mCherry'])
})
names(median_mCherry) = sampleNames(flowset_fluo_clones)


median_GFP = sapply(1:length(flowset_fluo_clones), function(i){
  median(exprs(flowset_fluo_clones[[i]])[,'GFP'])
})
names(median_GFP) = sampleNames(flowset_fluo_clones)

pData(flowset_fluo_clones)$median_mTurq = median_mTurq
pData(flowset_fluo_clones)$median_mCherry = median_mCherry
pData(flowset_fluo_clones)$median_GFP = median_GFP

pData(flowset_fluo_clones)$total_cell_count = total_cell_count
pData(flowset_fluo_clones)$measurement_date = measurement_date #date from the fcs file, must be correct

flowset_fluo_clones_data_tib = tibble(pData(flowset_fluo_clones))

```

### filter good clones
```{r}
good_clones_tib = flowset_fluo_clones_data_tib %>%
  filter(total_cell_count >= 1000 &
           (is.na(PCR_band_deletion) | PCR_band_deletion) & 
          
           (is.na(PCR_band_intact_allele) | PCR_band_intact_allele) 
  )

flowset_fluo_good = flowset_fluo_clones[good_clones_tib$exp_name]


```

### mark highest peak(s)
```{r}
#set up the transformations I use in the plots
fwd_transf  = flowWorkspace::flowjo_biexp(widthBasis = -100, pos = 4.42, inverse = F)
inv_transf = flowWorkspace::flowjo_biexp(widthBasis = -100, pos = 4.42, inverse = T)

#function to draw a line ad the two top peaks
#important: the values for the geom_density_ridges quantile lines need to be sorted non-decreasingly!
fun_high_peaks2 = function(x, ...){
  peaks_obj = getPeaks(fwd_transf(x))
  N_peaks_to_pick = min(2, length(peaks_obj$Peaks))
  # N_peaks_to_pick = 1
  top_2_peaks = order(peaks_obj$P.h, decreasing = T)[1:N_peaks_to_pick] 
  linear_peak = inv_transf(getPeaks(fwd_transf(x))$Peaks[top_2_peaks])
  sort(linear_peak)
}


```

##  Plot sorted pools
pools (and controls) of E2250
```{r S4C_remeasured_pools_E2250, fig.width = 5, fig.height=5}

#E2250 pools
samples_E2250_pools_paper_idx = pData(flowset_fluo_good)$experiment %in% c("E2250") &
  pData(flowset_fluo_good)$type %in% c("control", "sorted_pool") &
  pData(flowset_fluo_good)$sorted_phenotype %in% c("none", "mCh-_mT+") &
  !pData(flowset_fluo_good)$cell_line %in% c("F121-9")

# #order the samples in a sensible order
levels_pools =  pData(flowset_fluo_good[samples_E2250_pools_paper_idx]) %>% 
  mutate(treatment = factor(treatment, levels = c("delA", "delB", "untreated"))) %>%
  mutate(landing_pad = factor(landing_pad, levels = c("8", "23", "C9", "none"))) %>%
  arrange(landing_pad, treatment) %>% pull(exp_name)

ggplot(flowset_fluo_good[samples_E2250_pools_paper_idx], aes(x = mTurq, fill = treatment_for_color)) +
  facet_nested(  factor(landing_pad, levels = c("none", "C9", "23", "8"))  ~  ., scales = 'free_y', space = 'free_y') +
  geom_density_ridges(aes(y =  factor(exp_name, levels = rev(levels_pools))),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 1.5,
                      alpha = 0.75,
                      quantile_lines = TRUE, quantile_fun = fun_high_peaks2) + #, 
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42) +
  scale_y_discrete(expand = expansion( add = c(0, 2))) +
  theme_bw() +
  theme(legend.position = 'bottom',
        axis.title.y = element_blank()) +
  colscale_clones
```

## Plot sorted clones


```{r S4G_sup_clones_23_delA}
#select the samples to plot
samples_clones_23_delA = pData(flowset_fluo_good) %>%
  filter(type %in%  c("clone", "control") &
           landing_pad %in% c("23", "none") &
           sorted_phenotype %in% c("mCh-_mT+", "none") &
           treatment %in% c("untreated", "delA") &
           date == "ME20230426") %>% pull(exp_name)
samples_clones_23_delA_idx = pData(flowset_fluo_good)$exp_name %in% samples_clones_23_delA 

#order them by mTurq value / the controls in a sensible order
level_order_clones = pData(flowset_fluo_good[samples_clones_23_delA_idx]) %>%
  filter(type == "clone") %>%
  arrange(median_mTurq) %>% pull(cell_line_clone_name)
level_order <- c("WT", "F121-9", "23_34A", level_order_clones )

#plot
ggplot(flowset_fluo_good[samples_clones_23_delA_idx], aes(x = mTurq, fill = treatment_for_color)) +
  # facet_nested(   type ~ . , scales = 'free_y', space = 'free_y', render_empty = F) +
  geom_density_ridges(aes(y = factor(cell_line_clone_name, levels = level_order)),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 2,
                      alpha = 0.75,
                      quantile_lines = TRUE, quantile_fun = fun_high_peaks2, vline_width = 0.5) + #+ #,
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42, limits = c(0, 5*10^4)) +
  scale_y_discrete(expand = expansion( add = c(0, 1.5))) +
  # geom_vline(data = filter(peaks_linear_tib, exp_name %in% pData(flowset_fluo_good[samples_clones_23_delA_idx])$exp_name),aes(xintercept = mTurq_peak)) +
  theme_bw() +
  theme(legend.position = 'bottom',
        axis.title.y = element_blank(), 
        aspect.ratio = 1.5) +
  colscale_clones
        # axis.text.y = element_blank()

```

```{r S4Fsup_clones_23_delB}
#select the samples to plot
samples_clones_23_delB = pData(flowset_fluo_good) %>%
  filter(type %in%  c("clone", "control") &
           landing_pad %in% c("23", "none") &
           sorted_phenotype %in% c("mCh-_mT+", "none") &
           treatment %in% c("untreated", "delB") &
           date == "me20230418") %>% pull(exp_name)
samples_clones_23_delB_idx = pData(flowset_fluo_good)$exp_name %in% samples_clones_23_delB 

#order them by mTurq value / the controls in a sensible order
level_order_clones = pData(flowset_fluo_good[samples_clones_23_delB_idx]) %>%
  filter(type == "clone") %>%
  arrange(median_mTurq) %>% pull(cell_line_clone_name)
level_order <- c("WT", "F121-9", "23_34A", level_order_clones )

#plot
ggplot(flowset_fluo_good[samples_clones_23_delB_idx], aes(x = mTurq, fill = treatment_for_color)) +
  # facet_nested(   type ~ . , scales = 'free_y', space = 'free_y', render_empty = F) +
  geom_density_ridges(aes(y = factor(cell_line_clone_name, levels = level_order)),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 2,
                      alpha = 0.75,
                      quantile_lines = TRUE, quantile_fun = fun_high_peaks2, vline_width = 0.5) + #+ #,
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42, limits = c(0, 5*10^4)) +
  scale_y_discrete(expand = expansion( add = c(0, 1.5))) +
  # geom_vline(data = filter(peaks_linear_tib, exp_name %in% pData(flowset_fluo_good[samples_clones_23_delA_idx])$exp_name),aes(xintercept = mTurq_peak)) +
  theme_bw() +
  theme(legend.position = 'bottom',
        axis.title.y = element_blank(), 
        aspect.ratio = 1.5) +
  colscale_clones

```

```{r S4E_sup_clones_8_delB}
#select the samples to plot
samples_clones_8_delB = pData(flowset_fluo_good) %>%
  filter(type %in%  c("clone", "control") &
           landing_pad %in% c("8", "none") &
           sorted_phenotype %in% c("mCh-_mT+", "none") &
           treatment %in% c("untreated", "delB") &
           date == "20231128") %>% pull(exp_name)
samples_clones_8_delB_idx = pData(flowset_fluo_good)$exp_name %in% samples_clones_8_delB 

#order them by mTurq value / the controls in a sensible order
level_order_clones = pData(flowset_fluo_good[samples_clones_8_delB_idx]) %>%
  filter(type == "clone") %>%
  arrange(median_mTurq) %>% pull(cell_line_clone_name)
level_order <- c("WT", "F121-9", "8_34.8", level_order_clones )

#plot
ggplot(flowset_fluo_good[samples_clones_8_delB_idx], aes(x = mTurq, fill = treatment_for_color)) +
  # facet_nested(   type ~ . , scales = 'free_y', space = 'free_y', render_empty = F) +
  geom_density_ridges(aes(y = factor(cell_line_clone_name, levels = level_order)),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 2,
                      alpha = 0.75,
                      quantile_lines = TRUE, quantile_fun = fun_high_peaks2, vline_width = 0.5) + #+ #,
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42, limits = c(0, 5*10^4)) +
  scale_y_discrete(expand = expansion( add = c(0, 1.5))) +
  # geom_vline(data = filter(peaks_linear_tib, exp_name %in% pData(flowset_fluo_good[samples_clones_23_delA_idx])$exp_name),aes(xintercept = mTurq_peak)) +
  theme_bw() +
  theme(legend.position = 'bottom',
        axis.title.y = element_blank(), 
        aspect.ratio = 1.5) +
  colscale_clones

```

```{r S4D_sup_clones_C9_delB}
#select the samples to plot
samples_clones_C9_delB = pData(flowset_fluo_good) %>%
  filter(type %in%  c("clone", "control") &
           landing_pad %in% c("C9", "none") &
           sorted_phenotype %in% c("mCh-_mT+","mCh-_mT-", "none") & #for C9 we sorted mT high and lower clones from the pool, both are mTurq positive and should be included in the plot
           treatment %in% c("untreated", "delB") &
           date == "20231128") %>% pull(exp_name)
samples_clones_C9_delB_idx = pData(flowset_fluo_good)$exp_name %in% samples_clones_C9_delB 

#order them by mTurq value / the controls in a sensible order
level_order_clones = pData(flowset_fluo_good[samples_clones_C9_delB_idx]) %>%
  filter(type == "clone") %>%
  arrange(median_mTurq) %>% pull(cell_line_clone_name)
level_order <- c("WT", "F121-9", "C9_34", level_order_clones )

#plot
ggplot(flowset_fluo_good[samples_clones_C9_delB_idx], aes(x = mTurq, fill = treatment_for_color)) +
  # facet_nested(   type ~ . , scales = 'free_y', space = 'free_y', render_empty = F) +
  geom_density_ridges(aes(y = factor(cell_line_clone_name, levels = level_order)),
                      size =0.3, # rel_min_height  = 0.005,
                      scale = 2,
                      alpha = 0.75,
                      quantile_lines = TRUE, quantile_fun = fun_high_peaks2, vline_width = 0.5) + #+ #,
  ggcyto::scale_x_flowjo_biexp(widthBasis = -100, breaks = c(-10^3, 0, 10^3, 10^4, 10^5),
                               pos = 4.42, limits = c(0, 5*10^4)) +
  scale_y_discrete(expand = expansion( add = c(0, 1.5))) +
  # geom_vline(data = filter(peaks_linear_tib, exp_name %in% pData(flowset_fluo_good[samples_clones_23_delA_idx])$exp_name),aes(xintercept = mTurq_peak)) +
  theme_bw() +
  theme(legend.position = 'bottom',
        axis.title.y = element_blank(), 
        aspect.ratio = 1.5) +
  colscale_clones

```
## Boxplot highest peaks

calculate the normalized / relative medians
```{r}
# get the autofluorescence for all dates
autofluo_tib = good_clones_tib %>%
  group_by(flow_experiment, date) %>%
  filter(control_samples == "WT") %>%
  summarize(autofluo_mTurq = mean(median_mTurq),
            autofluo_mCherry = mean(median_mCherry),
            autofluo_GFP = mean(median_GFP)
            )

#get normalized corrected fluorescence for all samples
good_clones_tib_norm_tmp = good_clones_tib %>% 
  left_join(autofluo_tib) %>%
  #autofluorescence corrected (_cor)
  mutate(median_mTurq_cor = median_mTurq - autofluo_mTurq,
         median_mCherry_cor = median_mCherry - autofluo_mCherry,
         median_GFP_cor = median_GFP - autofluo_GFP) %>%
  #normalized to GFP (_norm)
  mutate(mTurq_norm = median_mTurq_cor / median_GFP_cor,
         mCherry_norm = median_mCherry_cor / median_GFP_cor)

# get standard values (23_34A) for all dates
standard_tib = good_clones_tib_norm_tmp %>% 
  group_by(flow_experiment, date) %>%
  filter(control_samples == "23_34A") %>%
  summarize(
    #mean autofluo corrected
    mTurq_cor_st = mean(median_mTurq_cor),
    mCherry_cor_st = mean(median_mCherry_cor),
    GFP_cor_st = mean(median_GFP_cor),
    # mean of normalized to GFP
    mTurq_norm_st = mean(mTurq_norm),
    mCherry_norm_st = mean(mCherry_norm),
    #mean raw median values
    mTurq_st = mean(median_mTurq),
    mCherry_st = mean(median_mCherry),
    GFP_st = mean(median_GFP)
    )

#get relative expression values for all samples
good_clones_tib_norm = good_clones_tib_norm_tmp %>%
  left_join(standard_tib) %>%
  # autofluorescence corrected and relative to 23_34A (_cor_rel)
  mutate(mTurq_cor_rel = median_mTurq_cor / mTurq_cor_st,
         mCherry_cor_rel = median_mCherry_cor / mCherry_cor_st,
         GFP_cor_rel = median_GFP_cor / GFP_cor_st,
         # autofluoresnce corrected, normalized to GFP and relative to 23_34A
         mTurq_norm_rel = mTurq_norm / mTurq_norm_st,
         mCherry_norm_rel = mCherry_norm / mCherry_norm_st,
         # relative to 23_34A only (no autofluorescence correction etc)
         mTurq_rel = median_mTurq / mTurq_st,
         mCherry_rel = median_mCherry / mCherry_st,
         GFP_rel = median_GFP / GFP_st  )


``` 

find the highest peaks and normalize them
```{r}
#samples to calculate
samples_oi_tib = pData(flowset_fluo_good) %>%
  filter(type %in% c("clone", "control")) %>%
  filter(flowcytometer == "analytical")
samples_oi = samples_oi_tib$exp_name


#find the two highest peaks in the density plot (transformed back to linear space)
# just take the two highest peaks for each sample, if it shows two peaks (unbiased approach)
peaks_linear_list = lapply(samples_oi, function(smpl){
  mTurq_expr = exprs(flowset_fluo_good[[smpl]])[,'mTurq']
  peaks_obj = getPeaks(fwd_transf(mTurq_expr))
  sample_type = pData(flowset_fluo_good) %>% filter(sample_name == smpl) %>% pull(type)
  N_peaks_to_pick = min(2, length(peaks_obj$Peaks)) #get two peaks if there are 2 or more, get only one if one exists
  top_2_peaks = order(peaks_obj$P.h, decreasing = T)[1:N_peaks_to_pick] 
  linear_peak = inv_transf(getPeaks(fwd_transf(mTurq_expr))$Peaks[top_2_peaks])
  linear_peak
})
names(peaks_linear_list) = samples_oi

#arrange in a tibble
peaks_linear_tib = tibble(exp_name = rep(names(peaks_linear_list), lengths(peaks_linear_list)),
                          mTurq_peak = unlist(peaks_linear_list))  %>%
  left_join(samples_oi_tib)


#correct, normalize and scale the peak data
peaks_linear_tib_scaled = peaks_linear_tib %>%
  left_join(good_clones_tib_norm) %>%
  mutate(mTurq_peak_cor = mTurq_peak - autofluo_mTurq,
         mTurq_peak_norm = mTurq_peak_cor/median_GFP_cor,
         mTurq_peak_rel = mTurq_peak_norm/mTurq_norm_st,
         mTurq_peak_rel_uncor = mTurq_peak / mTurq_st) %>%
  #mark the highest peak per sample
  group_by(exp_name) %>%
  mutate(highest_peak = mTurq_peak_rel >=  max(mTurq_peak_rel) & is.finite(mTurq_peak_rel),
         highest_peak_uncor = mTurq_peak_rel_uncor >= max(mTurq_peak_rel_uncor) )

``` 

Plot
```{r S4I_boxplot_CRISPR_clones, fig.width = 6, fig.height = 5}
#highest peak only
tib_highest_peak_clones = filter(peaks_linear_tib_scaled, is.na(flowcytometer) | flowcytometer == "analytical" ) %>%
  filter(GFP_state == "GFP+") %>%
  filter(flow_experiment %in% c("E2250", "E2254", "E2350", "E2427")) %>%
  filter((type %in% c("clone", "control") & flow_experiment != "E2427") ) %>%
  filter(is.na(cell_line) | cell_line != "WT") %>% #cannot be normalized to GFP
  mutate(landing_pad = factor(landing_pad, levels = c("none", "C9", "23", "8"),
                               labels = c("no insert", "-161 kb", "-116 kb", "-39 kb")),
         treatment = factor(treatment, levels  = c("untreated", "delA", "delB"),
                            labels = c("untreated", "ΔSox2", "ΔCBS_Sox2"))) %>%
  filter(highest_peak) 


#adding statistics does not work well for a faceted plot with missing combinations
#instead I just calculate the statistics separately and then add to the plot
library("rstatix")  
test_116 = tib_highest_peak_clones %>%
  ungroup() %>%
  filter(landing_pad == "-116 kb") %>%
  t_test(formula = mTurq_peak_rel ~ treatment,
         var.equal = F) %>%
  mutate(landing_pad = "-116 kb")
         
test_39 = tib_highest_peak_clones %>%
  ungroup() %>%
  filter(landing_pad == "-39 kb") %>%
  mutate(treatment = factor(treatment, levels = c("untreated", "ΔCBS_Sox2"))) %>%
  t_test(formula = mTurq_peak_rel ~ treatment,
         var.equal = F) %>%
  mutate(landing_pad = "-39 kb")

max_vals_plot = tib_highest_peak_clones %>% 
    ungroup() %>%
  group_by(landing_pad, treatment) %>%
  summarize(max_val = max(mTurq_peak_rel))

stats_tib = bind_rows(test_116, test_39) %>%
  left_join(max_vals_plot, by = join_by(landing_pad == landing_pad, group2 == treatment)) %>%
  mutate(adj_y = c(2, 5, 2, 4),
         y.position = max_val + adj_y)

# relative mTurq reporter expression highest peak
pHighest_Peak_mTurq = 
ggplot(tib_highest_peak_clones, aes(x = treatment, y = mTurq_peak_rel, col = treatment_for_color)) +
  facet_nested(.  ~ factor(landing_pad, levels = c("no insert", "-161 kb", "-116 kb", "-39 kb")) , 
               scales = 'free_x', space = 'free_x') +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  geom_hline(yintercept = 0) +
  #add data
  geom_boxplot(position = "dodge", outlier.shape = NA)+
  geom_point(position = position_jitterdodge(jitter.width = 0.6, jitter.height = 0), size = 2.5, alpha = 0.5)+
  
  #add statistics
  stat_pvalue_manual(stats_tib, label = "p") +
  
  #layout
  scale_x_discrete(guide = guide_axis(angle = 90))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  ylab("relative reporter expression\n(high peak)") +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none') +
  # theme(axis.title.x=element_blank(),
  #       axis.text.x=element_blank(),
  #       axis.ticks.x=element_blank())+
  colscale_clones

pHighest_Peak_mTurq



```

```{r S4I_2_raw_GFP_rel_to_23_34A,  fig.width = 6, fig.height = 5}
# Plot relative raw GFP levels (so autofluo subtracted, relative to 23_34A, as in figure S7)


plot_GFP_of_del_clones = 
  ggplot(tib_highest_peak_clones, aes(x = treatment, y = GFP_rel, col = treatment_for_color)) +
  facet_nested(.  ~ factor(landing_pad, levels = c("no insert", "-161 kb", "-116 kb", "-39 kb")) , 
               scales = 'free_x', space = 'free_x') +
  geom_hline(yintercept = 1, linetype = 'dashed') +
  geom_hline(yintercept = 0) +
  #add data
  geom_boxplot(position = "dodge", outlier.shape = NA)+
  geom_point(position = position_jitterdodge(jitter.width = 0.6, jitter.height = 0), size = 2.5, alpha = 0.5)+
  
  # add statistics
  # stat_pvalue_manual(stats_tib_GFP_rel, label = "p") +
  
  #layout
  scale_x_discrete(guide = guide_axis(angle = 90))+
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.1))) +
  ylab("median GFP_cor") +
  theme_bw(base_size = 14) +
  theme(legend.position = 'none') +
  colscale_clones

plot_GFP_of_del_clones
``` 



## Numbers for text
```{r}

median_rep_per_treatment = tib_highest_peak_clones %>% 
  group_by(cell_line, treatment) %>%
  summarize(median_rel_reporter = median(mTurq_peak_rel))

#change vs untreated, LP23 and LP8
median_rep_per_treatment %>%
  filter(cell_line == "23_34A") %>%
  mutate(change_reporter_vs_untr = median_rel_reporter / filter(median_rep_per_treatment, cell_line == "23_34A" & treatment == "untreated")$median_rel_reporter)

median_rep_per_treatment %>%
  filter(cell_line == "8_34.8") %>%
  mutate(change_reporter_vs_untr = median_rel_reporter / filter(median_rep_per_treatment, cell_line == "8_34.8" & treatment == "untreated")$median_rel_reporter)

#deletion LP23 vs LP8
median_rep_per_treatment %>%
  filter(cell_line == "23_34A") %>%
  mutate(del_23_vs_8 = median_rel_reporter / filter(median_rep_per_treatment, cell_line == "8_34.8" & treatment == "ΔCBS_Sox2")$median_rel_reporter)


```

#Hopping experiment

## Load data 
```{r}
#load the -161kb mCh- data (LP C9)
tib_C9_mChneg = read_tsv(path_tib_C9_mChneg_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  ungroup() %>%
  #mark hopped integrations
  mutate(hopped = start != 34598479) 

```
combine and filter
```{r}
#filter the integrations
tib_filtered <- tib_C9_mChneg  %>%
  #add variables used in the plotting (old names of cell lines etc) 
  mutate(population = case_when(sample_type == "control pool" ~ "ctrl",
                                .default = sorted_gate),
         cell_line = case_when(launch_pad_location == "-161 kb" & insert == "Sox2P" & genotype == "WT" ~ "C9_mCh_34",
                               launch_pad_location == "-161 kb" & insert == "Sox2P" & genotype == "Sox2::mCherry deleted" ~ "C9_mCh-_34"
                               ),
         landing_pad = case_when(launch_pad_location == "-116 kb" ~ "23",
                                 launch_pad_location == "-161 kb" ~ "C9")) %>%
  #annotate confidence of mapping (one or both ITRs)
  mutate(mapped_arms = case_when(read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only",
                                 read_count_1 > 0 & read_count_2 > 0 ~ "both_arms")) %>%
  #filter the data
  filter(population == "ctrl" | read_count >= 2) %>% #for ctrl distribution 1 mapping read is enough
  filter(!(start >= 34721183 & start <= 34721192)) %>% #remove contamination from LP8
  filter(! start %in% c(25018987, 35232946) ) %>% #remove contamination clone C1
  filter(landing_pad  == "C9" | start != 34598479) %>% #remove contamination from LP C9
  filter(landing_pad  == "23" | start != 34643961) %>% #remove contamination from LP 23
  filter(landing_pad  == "C10" | start != 35478413) %>% #remove contamination from LP C10
  filter(strand %in% c("+", "-"))


tib_filtered_P1_strict = tib_filtered %>%
  filter(!(population == "P1" & mapped_arms != "both_arms")) #only keep P1 integrations with 2 mapped arms

all_exp = unique(tib_filtered_P1_strict$experiment)
```

### Add expression data
For C9_mCh-_34 we just use the FACS data from that cell line
```{r}
#load median expression values (from sorting pdfs)
median_sorted_mTurq = read_tsv(path_median_mTurq_relevant_cell_lines)

#add mTurq values to the integration tibble
tib_filtered_mTurq_vals = left_join(tib_filtered_P1_strict, median_sorted_mTurq) %>%
  filter(population != "ctrl") %>%
   dplyr::rename(mTurq_expr_gate = mTurq) 

#check that all integrations have a fluorescence value:
tib_filtered_mTurq_vals %>% filter(is.na(mTurq_expr_gate)) %>% ungroup()
```

### Sorting statistics
For the new expression score we need to know, per population and replicate:
1. number of cells sorted
2. fraction of the cells in that gate (take fraction of the mChpos/mChneg total cells, whichever is relevant for that cell_line)

We add a psuedocount of 1 to all the 'recorded_count' when calculating the 'recorded_fraction', to deal with the 0 recorded cells in P1. 
Got the sorting stats from the sorting pdfs & the labguru notes of the experiments

```{r}
sorting_stats_tib_raw = read_xlsx(path_sorting_stats) 

sorting_stats_tib = sorting_stats_tib_raw %>%
  mutate(recorded_fraction = (recorded_count + 1)/total_recorded) %>% #recorded fraction with psuedocount of 1
  mutate(weigh_factor = recorded_fraction / sorted_count, 
         weigh_factor_1M = weigh_factor * 1E6) # weigh factor 1M is just to make the factor more easily human readable
```

Due to the 50K pools, there are more replicates for P2-P6 than for P1. Also, the split gates (P1H/L and P2H/L) are extra replicates for those populations. Therefore, we adjust the weights of the integrations, so the total weight counts up to the same number of replicates for each population.
Note: the adjustment to combine P1H/L into P1 and P2H/L into P2 is done manually (so adjust the code if more replicates are added)

```{r}
sorting_stats_tib_repN = sorting_stats_tib %>%
  group_by(cell_line, population) %>%
  filter(cell_line == "C9_mCh-_34")%>%
  #count the number of reps per population
  summarize(N_reps = n()) %>%
  mutate(N_reps_adj = case_when(
    #manually combine P1_H/L (2 each) into P1 (2) => 2+2 is 4 times coverage of the whole P1
    cell_line == "C9_mCh-_34" & population %in% c("P1", "P1_H", "P1_L") ~ 4,
    #manually combine P2_H/L (1 each) into P2 (4) => 1+4 is 5 times coverage of the whole P1
    cell_line == "C9_mCh-_34" & population %in% c("P2", "P2_H", "P2_L") ~ 5,
    .default = N_reps)  ) %>%
  group_by(cell_line) %>%
  #find the least replicated population(s)
  mutate(min_reps_CL = min(N_reps_adj) ) %>%
  #adjust the other weights
  mutate(rep_weigh_factor = min_reps_CL/N_reps_adj)


# adjust weigh factor by the replicate weigh factor
sorting_stats_tib_repcor = 
  sorting_stats_tib %>%
  filter(cell_line == "C9_mCh-_34")%>%
  left_join(sorting_stats_tib_repN) %>%
  mutate(weigh_factor_old = weigh_factor,
         weigh_factor = weigh_factor_old * rep_weigh_factor)
  

```

## Beeswarm C9_mCh-_34
```{r}
plot_beeswarm_pink = function(TIB, CL = "C9_mCh-_34", EXP = all_exp, P_RANGE = prange_zoom,
                              PLOT_CTCF = F, PLOT_ANNOT = T,
                              TITLE = NULL, POINT_SIZE = 0.75, RANDOM_METHOD = "quasirandom",
                              NEXT_RANGE = NULL){
  tib_for_plot = filter(TIB, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP & start >= P_RANGE[1] & start <= P_RANGE[2])
  LP_tib_cl = select(tib_for_plot, c(cell_line, landing_pad)) %>%
    distinct() %>% left_join(LP_tib)

  ggplot(filter(tib_for_plot, hopped & chr == "chr3" & cell_line %in% CL & experiment %in% EXP),
         aes(x = start, y = population, col = population)) +
    {if(length(CL)>1) facet_grid(cell_line ~ .)} +  #only facet when necessary

    {if (PLOT_ANNOT)
      list( #add elements in a list, you cannot use + inside an if statement
        # annotate enhancer
        geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
        annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
        # annotate gene
        annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red'),
        geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') )} +
    {if(PLOT_ANNOT & length(CL)== 1)
      geom_vline(xintercept = pull(
        filter(LP_tib, landing_pad %in% unique(tib_for_plot$landing_pad)),
        start),
        col = 'darkgrey')} +
    {if(PLOT_ANNOT & length(CL) > 1)
       geom_vline(data = LP_tib_cl, aes(xintercept = start), col = 'darkgrey')} +

    # # annotate CTCFs
    {if(PLOT_CTCF)geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey', lty="11")}+
    {if(!is.null(NEXT_RANGE)) geom_rug(data = tibble(range = NEXT_RANGE),
                                       aes(x = range),
                                       fill = 'grey',
                                       inherit.aes = F)}+
     #annotate landing pad
    labs(x='Genomic coordinate (Mb)',y='sorted gate') +

    #datapoints
    geom_quasirandom(alpha = 0.5, size = POINT_SIZE, varwidth = T, bandwidth = 0.8,
                     # shape = 16, #changes to point without border, so alpha applies to the whole point
                     method = RANDOM_METHOD,
                     stroke = NA) +
    geom_text(data = plyr::count(tib_for_plot, vars = c("cell_line", "population")), aes(label = paste("n = ", freq, sep = ""),
                                                                                         x = Inf),
              hjust = "inward", vjust = "top",
              col = 'black') +
    #layout:
    theme_classic(base_size = 16) +
    theme(legend.position = "none") +
    scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                       limits=P_RANGE, expand=c(0,0)) +
    scale_y_discrete(limits = rev, drop = F) +
    ggtitle(TITLE) +
    colScale7_pink
}
```

```{r S4K_beeswarm_mChneg_prange, fig.width = 6, fig.height = 3}
tib_for_plot = filter(tib_filtered_P1_strict, hopped & chr == "chr3" &  experiment %in% all_exp ) %>%
  mutate(population = factor(population, levels = c("ctrl", "P1_H", "P1", "P1_L",
                                                    "P2_H", "P2", "P2_L",
                                                    paste0("P", 3:6))))

prange_beeswarm =  c(34.5E6, 34.9E6)

C9_mChneg_34_beesw = plot_beeswarm_pink(TIB = tib_for_plot,
                   CL = c("C9_mCh-_34"),
                   PLOT_ANNOT = T,
                   PLOT_CTCF = T,
                   P_RANGE = prange_beeswarm,
                   POINT_SIZE = 1.5
)
C9_mChneg_34_beesw
```



### Orientation CTCF sites
Plotting triangles in a pdf is a nightmare, so instead 1 just make one plot with the CTCF lines colored by orientation and based on that add the CTCF triangles manually to the paper figures.

```{r}
ggplot() +
   # annotate enhancer
  # annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown)+
  geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown)+
  # annotate gene
  # annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf, alpha = .2, fill = 'red')+
  geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red') +
  
  
  geom_vline(data = as_tibble(CTCF_mm10.chr3_extra), aes(xintercept = start, col = strand), size = 2, lty = 'dashed') +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=c(prange_beeswarm), expand=c(0,0)) +
  theme_bw()

```

## Expression score

### adjust for deletion
Because the C9_mCh-_34 cell line contains a deletion of the Sox2 promoter, gene (and upstream CTCF site), calculating the running mean on the normal genome will give bad results close to the deletion (since part of the window will be in the deleted region).
Therefore, we will adjust all integrations after the deletion to shift up by the deletion size, then calculate the running mean and then shift the integrations back.

There are a few (11) integrations mapped inside the deleted region, most likely these are either bleedthrough reads from another cell line (sequenced in the same experiment) or integrations on the other allele. However, they are so few (relative to the 2600 integrations in the region) that we can ignore them. 
```{r}

#Define the deleted range
deletion_range = GRanges(seqnames = "chr3", 
                         ranges = IRanges(start = 34647592,
                                          end = 34652615),
                         strand = "+")
del_width = width(deletion_range) - 1

tib_filtered_mTurq_vals %>% filter(chr == "chr3" & start >= 34647592 & end <= 34652615)

tib_filtered_mTurq_vals %>% filter(chr == "chr3" & start >=  prange_zoom_paper[1] & end <= prange_zoom_paper[2]) %>% nrow()

```

```{r}
#add the adjusted sorting stats to the integration tibble
tib_filtered_mTurq_vals_repcor = left_join(tib_filtered_mTurq_vals, sorting_stats_tib_repcor)

#for C9_mCh-_34 remove clones in deleted region, shift everything after deletion by 'del_width'
tib_filtered_mTurq_vals_repcor_deladj = tib_filtered_mTurq_vals_repcor %>% 
  filter(!(chr == "chr3" & start >= start(deletion_range)  & start <= end(deletion_range) & cell_line == "C9_mCh-_34")) %>%
  mutate(start_original = start) %>%
  mutate(start = case_when(cell_line == "C9_mCh-_34" & 
                             chr == "chr3" & start_original >= end(deletion_range) ~ (start_original - del_width),
                           .default = start_original))
  


```

Note: the 'start' column is changed so that the expression score calculation works
The 'end' column is not changed, and 'start_original' contains the non-shifted coordinates.

### calculate expression score


```{r}
#C9_mCh+_34 5kb window
set.seed(20240328)
min_ints_window = 3
N_bootstrap = 5000
BinSize = 500
CL = "C9_mCh-_34"
all_exp = unique(tib_filtered_mTurq_vals$experiment)


# # calculate running mean on random sample of same size, 500 times
# expr_score_tib = create_bootstrap_tibble(FUN = function(TIB){expr_score_function_new_weigh(TIB,
#                                                                                               binsize = BinSize,
#                                                                                               N_bins_window = 10,
#                                                                                               calculation_window =  prange_zoom,
#                                                                                               CL = CL, EXP = all_exp)},
#                                             TIB = tib_filtered_mTurq_vals_repcor_deladj,
#                                             N_resamples = N_bootstrap)
# saveRDS(expr_score_tib, "/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250328_new_expression_score_C9_mChneg_34_5kb.rds")

expr_score_tib = readRDS("/DATA/projects/Sox2/Figure_reporter_only_hopping/rds_files/CM20250328_new_expression_score_C9_mChneg_34_5kb.rds")



expr_score_sum_tib = bootstrap_summary_fun(expr_score_tib,
                                           score_column = "expr_score_window",
                                           conf_int = 0.95)

#calculate actual expr score (for filtering # of ints per window)
expr_score_real =
  expr_score_function_new_weigh(tib_filtered_mTurq_vals_repcor_deladj,
                                binsize = BinSize,
                                N_bins_window = 10,
                                calculation_window =  prange_zoom,
                                CL = CL,
                                EXP = all_exp)
  
expr_score_sum_tib_filt_C9_mChneg_34 = left_join(expr_score_sum_tib,  expr_score_real) %>%
  mutate(expr_score_window_median_sel = case_when(
    total_ints_window >= min_ints_window ~ expr_score_window_median,
    .default = NA),
    expr_score_window_lower_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_lower,
      .default = NA),
    expr_score_window_upper_sel = case_when(
      total_ints_window >= min_ints_window ~ expr_score_window_upper,
      .default = NA)) %>%
  mutate(cell_line = CL)
  

```

```{r}

# an empty (NA) row in the middle of the deleted interval
fake_row = expr_score_real %>% filter(interval_number == 1) %>%
  mutate(expr_score_window_median  = NA, 
         expr_score_window_lower = NA,
         expr_score_window_upper = NA,
         interval_number = NA,
         expr_score = NA,
         mid_interval = mid(deletion_range)) 

# shift back and add fake row
expr_score_sum_tib_filt_C9_mChneg_34_adj = expr_score_sum_tib_filt_C9_mChneg_34 %>%
  mutate(mid_interval = case_when(mid_interval >= start(deletion_range) ~ (mid_interval + del_width),
                                  TRUE ~ mid_interval)) %>%
  bind_rows(fake_row)


```

```{r}
#load the bootstrapped data from C9_mCh_34 (generated in fig2)
expr_score_sum_tib_filt_C9_34 = readRDS(path_expr_score_bootstrap_C9_mCh_34)

expr_sum_tib_filt_comb = bind_rows(expr_score_sum_tib_filt_C9_34, expr_score_sum_tib_filt_C9_mChneg_34_adj)
```


### Plot
```{r 4C_expr_plot_mChneg_pos, fig.width=6, fig.height = 4}

 ggplot(expr_sum_tib_filt_comb, aes(x = mid_interval)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_C9), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel,
                  fill = cell_line), alpha = 0.3) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel, col = cell_line)) +

  #plot the deletion range
  geom_rug(data = tibble(del_range = c(start(deletion_range), end(deletion_range))), aes(x = del_range), inherit.aes = F) +
  
  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')
  

```

```{r}
# Zoom-out a bit 
ggplot(expr_sum_tib_filt_comb, aes(x = mid_interval)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_C9), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel,
                  fill = cell_line), alpha = 0.3) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel, col = cell_line)) +

  #plot the deletion range
  geom_rug(data = tibble(del_range = c(start(deletion_range), end(deletion_range))), aes(x = del_range), inherit.aes = F) +
  
  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=c(34580000, 34830000), expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')

```


### increase and spreading

How much is the increase in expression, between Sox2 gene and SCR?
First attempt: just take the difference between the two cell lines per bin. This is not possible because the mCh- bins are shifted (due to the deletion)
Still, middle of bins is only off by 23 bp, so I just shift mid-interval a bit and then take the difference

Take the range from the end of Sox2 (deletion) to the end of the SCR:
```{r}
mCh_del_expr_dif_tib = expr_sum_tib_filt_comb %>%
  filter(mid_interval >= end(Sox2_gene) & mid_interval <= end(enhancer)) %>%
  mutate(mid_interval_adj = case_when(cell_line == 'C9_mCh-_34' ~ mid_interval - 23,
                                      .default = mid_interval)) %>%
  ungroup() %>%
  select(mid_interval_adj, cell_line, expr_score_window_median_sel) %>%
  pivot_wider(names_from = cell_line, values_from = expr_score_window_median_sel) %>%
  mutate(dif_mChneg = `C9_mCh-_34` - `C9_mCh_34`,
         fc_mChneg = `C9_mCh-_34` / `C9_mCh_34`)


summary(mCh_del_expr_dif_tib$dif_mChneg)
summary(mCh_del_expr_dif_tib$fc_mChneg)

quantile(mCh_del_expr_dif_tib$fc_mChneg, 0.05)
quantile(mCh_del_expr_dif_tib$fc_mChneg, 0.95)
```
8.4 fold (IQR 6.2-11.8)


Calculate the half-maximum expression and how far this spreads, for each cell line. 
```{r}
half_max_expr_tib = expr_sum_tib_filt_comb %>%
  filter(mid_interval >= end(Sox2_gene) & mid_interval <= end(enhancer)) %>%
  group_by(cell_line) %>%
  summarize(max_mTurq = max(expr_score_window_median_sel)) %>%
  mutate(half_max = max_mTurq/2)

#first bin after the SCR
half_max_after_enhancer = expr_sum_tib_filt_comb %>% 
  select(interval_number, mid_interval, cell_line, expr_score_window_median_sel) %>%
  left_join(half_max_expr_tib) %>%
  filter(expr_score_window_median_sel <= half_max & mid_interval >= end(enhancer)) %>% 
  arrange(mid_interval) %>%
  group_by(cell_line) %>%
  summarize(first_int_after_SCR = min(mid_interval))


half_max_after_enhancer %>% 
  pivot_wider(names_from = cell_line, values_from = first_int_after_SCR) %>%
  mutate(dif_in_position = `C9_mCh-_34` - `C9_mCh_34`)

#last bin before gene
half_max_before_sox2 = expr_sum_tib_filt_comb %>% 
  select(interval_number, mid_interval, cell_line, expr_score_window_median_sel) %>%
  left_join(half_max_expr_tib) %>%
  filter(expr_score_window_median_sel <= half_max & mid_interval <= start(Sox2_gene)) %>% 
  arrange(desc(mid_interval)) %>%
  group_by(cell_line) %>%
  summarize(last_int_before_sox2 = max(mid_interval))


half_max_before_sox2 %>% 
  pivot_wider(names_from = cell_line, values_from = last_int_before_sox2) %>%
  mutate(dif_in_position = `C9_mCh-_34` - `C9_mCh_34`)

16500 + 6000
```
"With the positions of the half-maximum expression extending by 22.5 kb (16.5kb upstream, 6kb downstream)."

check the locations visually
```{r}

ggplot(expr_sum_tib_filt_comb, aes(x = mid_interval, col = cell_line, fill = cell_line)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  # Line downstream SCR spread of activity
  geom_hline(data = half_max_expr_tib, aes(yintercept = half_max, col = cell_line)) +
  geom_vline(data = half_max_before_sox2, aes(xintercept = last_int_before_sox2, col = cell_line)) +
  geom_vline(data = half_max_after_enhancer, aes(xintercept = first_int_after_SCR, col = cell_line)) +

  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel), alpha = 0.2, colour = NA) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel)) +

  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=prange_zoom_paper, expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')


```

Range of any expression. As in figure 2, define detectable expression as expression score >= mean mTurq P5/P6.
We cannot really calculate this after the SCR, because the expression score of C9_mCh_34 never drops far enough (not enough data for the lower expression levels)

```{r}
# Get stats
## Look what the levels of P5 and P6 is for the three different cell lines in that plot C9, C10, 23

tib_P_stats = median_sorted_mTurq %>%
  filter(cell_line %in% c("23_34A", "C10_34", "C9_mCh_34") & population %in% c("P5", "P6")) %>%
  group_by(population) %>%
  summarise(mean_P = mean(mTurq)) %>%
  summarise(any_expr_threshold = mean(mean_P))

expr_threshold = tib_P_stats$any_expr_threshold
expr_threshold_20prc = 1.2*expr_threshold

#last bin before gene
any_expr_before_sox2 = expr_sum_tib_filt_comb %>% 
  select(interval_number, mid_interval, cell_line, expr_score_window_median_sel) %>%
  filter(expr_score_window_median_sel <= expr_threshold & mid_interval <= start(Sox2_gene)) %>% 
  arrange(desc(mid_interval)) %>%
  group_by(cell_line) %>%
  summarize(last_int_before_sox2 = max(mid_interval))


any_expr_before_sox2 %>% 
  pivot_wider(names_from = cell_line, values_from = last_int_before_sox2) %>%
  mutate(dif_in_position = `C9_mCh-_34` - `C9_mCh_34`)

start(Sox2_gene) - pull(filter(any_expr_before_sox2, cell_line == "C9_mCh-_34"), last_int_before_sox2)

```
And detectable expression spreads 70kb upstream compared to the C9_mCh+_34
Or: there is detectable expression up to 88kb upstream of the Sox2 gene

```{r}
ggplot(expr_sum_tib_filt_comb, aes(x = mid_interval, col = cell_line, fill = cell_line)) +
  # annotate enhancer
  list(
    annotate("rect", xmin = start(enhancer), xmax = end(enhancer), ymin = -Inf, ymax = Inf, alpha = .2, fill = brown),
    geom_vline(xintercept = c(start(enhancer), end(enhancer)), col = brown),
    # annotate gene
    annotate("rect", xmin = start(Sox2_gene), xmax = end(Sox2_gene), ymin = -Inf, ymax = Inf,alpha = .2, fill = 'red'),
    geom_vline(xintercept = c(start(Sox2_gene), end(Sox2_gene)), col = 'red'),
    geom_vline(xintercept = start(landingPad_23), col = 'darkgrey'),
    geom_vline(xintercept = start(CTCF_mm10.chr3_extra), col = 'darkgrey',linetype = 'dotted')
  ) +

  # Line downstream SCR spread of activity
  geom_hline(yintercept = expr_threshold) +
  geom_vline(xintercept = pull(filter(any_expr_before_sox2, cell_line == "C9_mCh-_34"), last_int_before_sox2)) +
  ## plot the confidence interval based on the resampled data
  geom_ribbon(aes(ymin = expr_score_window_lower_sel, ymax = expr_score_window_upper_sel), alpha = 0.2, colour = NA) +
  # geom_line( aes(y = expr_score_window_mean), col = 'orange') +
  ## plot the median based on the resampled data
  geom_line(aes(y = expr_score_window_median_sel)) +

  #axes
  scale_x_continuous(breaks = scales::pretty_breaks(n = 6), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL),
                     limits=c(34.55E6, 34.88E6), expand=c(0,0)) +
  scale_y_continuous(expand=expansion(mult = c(0, 0.05)),
                     breaks = scales::pretty_breaks(n=6),
                     limits = c(0, NA)) +
  labs(x='Genomic coordinate (Mb)',y='expression score') +
  theme_classic(base_size = 16) +
  theme(legend.position = 'none')

```


## FACS hopping


### load FACSDiva file E2687
```{r}
diva_ws_E2687 <- open_diva_xml(diva_xml_file_E2687_round2_sorting)
  

gs_E2687 <- diva_to_gatingset(diva_ws_E2687,
                        name = 1, #the group to import
                        path = path_FACS_sorting_E2687_round2,
                        swap_cols = F, #apparently we don't need to swap the FSC/SSC-H/W here (this is a known diva 'bug', if you import the xml into flowjo the columns are swapped but apparently not here)
                        worksheet = "global",
                        verbose = F,
                        execute = T)

pdata_tib_E2687 = pData(gs_E2687) %>%
  as_tibble() %>%
  mutate(name_short = str_remove(str_remove(name, "me20241031_E2687_round2_"), "_([0-9]){3}.fcs"),
         cell_line = case_when(grepl("WT_ctrl", name_short) ~ "WT",
                               grepl("CM2285", name_short) ~ "F121_9",
                               grepl("CM2267", name_short) ~ "23_34A_delB_mCh-",
                               grepl("CM2268", name_short) ~ " 23_34A_delB_GFP-",
                               grepl("CM2350", name_short) ~ "F121_9",
                               grepl("23_34", name_short) ~ "23_34A",
                               grepl("C9_34_mCh-", name_short) ~ "C9_mCh-_34",
                               grepl("C9_34_P6", name_short) ~ "P6_mCh-_34",
                               grepl("C9_41_P6", name_short) ~ "P6_mCh-_41",
                               grepl("C10_P7", name_short) ~ "C10_34"),
         treatment = case_when(grepl("PB", name_short) ~ "PB",
                               grepl("SB", name_short) ~ "SB",
                               grepl("Martijn", name_short) ~ "other_experiment",
                               .default = "untreated"),
         recording = case_when(grepl("_SORT", name_short) ~ "sorting",
                               .default = "pre-recording"),
         sample_type = case_when(cell_line %in% c("WT", "F121_9") | treatment == "untreated" ~ "control",
                                 .default = "sample")
  ) 


pdata_tib_E2687_df = as.data.frame(pdata_tib_E2687)
rownames(pdata_tib_E2687_df) = pdata_tib_E2687_df$name

pData(gs_E2687) = pdata_tib_E2687_df

```

### plot gates sorting

```{r, S4J_plot_8_gates_C9_mChneg_34, fig.height = 6, fig.width = 6}
gs_oi_E2687 = subset(gs_E2687, sample_type == "sample"  & cell_line == "C9_mCh-_34"  & 
                        (recording == "sorting" | treatment == "PB")) #sorting only, combining 2 data sets doesn't give the right 
ggcyto::ggcyto(gs_oi_E2687, aes(x = "GFP", y = "mTurqoise2"), subset = "mCh_neg") + 
  #facet by cell line / treatment
  facet_grid(cell_line ~ treatment) +

  #plot cells
  geom_hex(bins = 128, aes( fill = after_stat(ncount))) +
  scale_fill_distiller(palette = 'Spectral') + #manually adding a fill scale also forces the scale to be linear again.

  #plot_gates
  ggcyto::geom_gate(c("P1_high_mCh-", "P1_low_mCh-", "P2_high_mCh-", "P2_low_mCh-",  
                      paste0("P", 3:6, "_mCh-")), 
                    col = 'black') +
  ggcyto::geom_stats(adjust = c(-0.3, 0.5), digits = 2, size = 3) +
  # ggcyto::geom_stats(data = perc_for_plot, aes(label = percent), 
  #              inherit.aes = FALSE)+

  #layout
  theme_bw(base_size = 14) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.position = "none",
        aspect.ratio = 1) +
  ggtitle(element_blank())+
  ggcyto::axis_x_inverse_trans() +
  ggcyto::axis_y_inverse_trans() +
  ggcyto::ggcyto_par_set(limits = list(x = c(1.2, 4.5), y = c(0, 4.5))) +
  scale_y_continuous(expand = c(0,0)) +
  scale_x_continuous(expand = c(0,0))+
  ggcyto::labs_cyto(labels = "marker")

```

Note: the percentages for P1/2-H/L are wrong, because this is the percentage of the parent population, which is P1/2 for these.
Instead, we want to show the percentage of the whole mCh_neg population.
Here I calculate the correct percentages, which we can then manually change in the figure
```{r}
# percentages on the gates
count_tib_for_perc = gs_pop_get_stats(gs_oi_E2687, nodes = c("P1_high_mCh-", "P1_low_mCh-", "P2_high_mCh-", "P2_low_mCh-",  
                      paste0("P", 1:6, "_mCh-"), "mCh_neg"))
count_tib_for_perc_tib = as_tibble(count_tib_for_perc) %>%
  group_by(sample) %>%
  mutate(count_mChneg = count[pop == "mCh_neg"]) %>%
  mutate(percent = count/count_mChneg *100)

count_tib_for_perc_tib %>%
  dplyr::filter(pop %in% c("P1_high_mCh-", "P1_low_mCh-", "P2_high_mCh-", "P2_low_mCh-"))

```

# ChIP around Sox2 gene
When we delete the Sox2 gene, we observe silencing of the -116kb reporter over time. 
We expect this might be epigenetic silencing, due to the PRC2 peaks around the gene encroaching on the (less strong) reporter.
Here I make some plots to show that these peaks are present around the gene (and where the reporter is).

```{r }
prange_Sox2_gene_gr = GRanges(seqnames = 'chr3', IRanges(start = 34630000, end = 34680000))

## get the Sox2 gene annotation from RefSeq
# Define UCSC mm10 RefSeq annotation URL
url_mm10_refseq <- "http://hgdownload.soe.ucsc.edu/goldenPath/mm10/database/ncbiRefSeqCurated.txt.gz"

# Define a temporary file to store the downloaded data
temp_file <- tempfile(fileext = ".gz")

# Download the file
download.file(url_mm10_refseq, temp_file)

# Read the file (gunzipping automatically)
refseq_data <- read.table(gzfile(temp_file), header = F, sep = "\t", stringsAsFactors = FALSE)

# Assign proper column names based on UCSC schema
colnames(refseq_data) <- c("bin", "transcript_id", "chromosome", "strand", "txStart", "txEnd", 
                           "cdsStart", "cdsEnd", "exonCount", "exonStarts", "exonEnds",
                           "score", "gene_symbol",  "cdsStartStat", "cdsEndStat", "exonFrames"   )

#Get Sox2 gene
Sox2_tib = refseq_data %>% as_tibble() %>%
  filter(gene_symbol == "Sox2") %>%
  select(1:9, gene_symbol)

plot_gene_annot = ggplot(data = Sox2_tib) +
  geom_rect(aes(xmin = txStart, xmax = txEnd), ymin = 0.2, ymax = 0.4) +
  geom_rect(aes(xmin = cdsStart, xmax = cdsEnd), ymin = 0, ymax = 0.6) +
  geom_vline(xintercept = start(landingPad_23 ), col = 'darkgrey') +
  geom_vline(xintercept = start(CTCF_mm10.chr3_extra ), col = 'darkgrey',  lty="11") +
  
  scale_x_continuous(limits = c(start(prange_Sox2_gene_gr), end(prange_Sox2_gene_gr)), expand = c(0,0),
                     breaks = scales::pretty_breaks(n = 4), labels = scales::unit_format(scale = 1E-6, accuracy = 0.01, unit=NULL)) + scale_y_continuous(limits = c(0,0.6))+
  theme_classic() +
  theme(legend.position = 'none')

```

```{r S4H_ChIP_around_Sox2_gene_PRC1, fig.width = 6, fig.height = 5}
paths_ChIP_ls = split(unname(paths_ChIP_files),names(paths_ChIP_files))

cowplot::plot_grid(plotlist = c(list(plot_gene_annot,
                                     BigwigTrack(prange_Sox2_gene_gr, paths_ChIP_ls[2], smooth = 100) +
                                       scale_x_continuous(expand= c(0,0)),
                                     BigwigTrack(prange_Sox2_gene_gr, paths_ChIP_ls[1], smooth = 100) +
                                       scale_x_continuous(expand= c(0,0)),
                                     BigwigTrack(prange_Sox2_gene_gr, paths_ChIP_ls[3], smooth = 100) +
                                       scale_x_continuous(expand= c(0,0)),
                                     BigwigTrack(prange_Sox2_gene_gr, paths_ChIP_ls[4], smooth = 100) +
                                       scale_x_continuous(expand= c(0,0)),
                                     BigwigTrack(prange_Sox2_gene_gr, paths_ChIP_ls[5], smooth = 100) +
                                       scale_x_continuous(expand= c(0,0))
)),
                   ncol = 1, align="v", axis="lr")


```



