---
title: "CM20250424_hopping_statistics"
author: "Mathias & Christine"
date: "2025-03-31"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

#Goal
Make a table of the number of integrations per cell line and population. This is the source for table S7.


```{r load_libraries}

# Load dependencies
library(GenomicRanges)
# library(rtracklayer)
library(tidyverse)
# library(gtools)
library(tidyverse)
# library(dtplyr)
library(dplyr)
# library(tidyr)
# library(ggbeeswarm)
# library(ggplot2)
library(Biostrings)
library(stringr)
library(readr)
library(knitr)
# library(RColorBrewer)
# library(cowplot)
# library(RcppRoll)
# library(viridis)
# library(ggh4x)
# library(ggrastr)
# library(caTools)
# library(readxl)
# import::from(flowCore, .except=c("filter")) #I don't load the whole flowCore and ggcyto libraries, because they overwrite dplyr's filter function :0
# import::from(ggcyto, 'fortify_fs')
# library(GGally)
# library(GENOVA)
# library(ggpubr)
# library(ggridges)
# library(ggpmisc)

split_string <- function(vect,sep,N1,N2=N1){
  library(stringr)
  sapply(vect, function(X){
    paste(str_split(X,sep)[[1]][N1:N2],collapse = sep)
  },USE.NAMES = F)
}


```

# Paths & parameters
```{r paths_parameters}
#datatag
datetag = paste0("CM",format(Sys.time(), '%Y%m%d'))

# Prepare output 
output_dir <- paste0("/DATA/projects/Sox2/Figure_reporter_only_hopping/analysis_",datetag)
dir.create(output_dir, showWarnings = FALSE)

```

```{r knits setup}

library(knitr)
opts_chunk$set(cache = T,
               message = F, 
               warning = F,
               dev=c('png', 'pdf'), 
               dpi = 600,
               fig.path = file.path(output_dir, "figures/"),
               cache.lazy = F) 
pdf.options(useDingbats = FALSE)
```


## File paths
```{r}
path_CTCF_sites = "/DATA/usr/v.franceschini/Workspaces/2024_01_MATHIAS_CTCF_PEAKS/VF240812_calling_CTCF_motifs_again/02_OUTPUTS/CTCF_sites/6764_2_ME_E2_CGATGT_S2_R1_001_peaks_motifs.bed"

# path_tib_23_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-116kb_Sox2P_pools.txt"
path_tib_23_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/still_to_upload/mapped_integrations_-116kb_Sox2P_pools.txt"

path_tib_C9_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-161kb_Sox2P_pools.txt"

path_tib_C10_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/still_to_upload/mapped_integrations_+715kb_Sox2P_pools.txt"

path_tib_C9_41_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_-161kb_Sox2P_CDS_pools.txt"
path_tib_C9_mChneg_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/still_to_upload/mapped_integrations_-161kb_Sox2mCherrydel_Sox2P_pools.txt"


path_median_mTurq_relevant_cell_lines =  "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250318_median_mTurq_sorting_for_paper.tsv"
path_sorting_stats = "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250327_stats_for_expression_score_for_paper.xlsx"

path_CRE6_ints = "/DATA/projects/Sox2/GEO_collection/Tagmentation/mapped_integrations_short_neutral_insert.txt"

```

## Annotations
```{r}
prange_zoom_paper = c(34.59E6, 34.83E6)
range_2Mb = c(33.643960E6, 35.643960E6)

```

# Load all data

mCh+ Sox2P hopping (3LPs)
```{r}
#load the -116kb data (LP23)
tib_23_all_data = read_tsv(path_tib_23_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) 

#the control pool in E2285 (rep3) was sequenced as 10 separate libraries, combine them into one sample
tib_23_E2285_ctrls = tib_23_all_data %>%
  filter(replicate == "rep3" & insert == "Sox2P" & sample_type == "control pool") %>%
  mutate(sample_name = "Sox2P-reporter -116 kb, unsorted control pool rep3")%>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") 

tib_23 = tib_23_all_data %>%
  #select the rest of the data
  filter(!(replicate == "rep3" & insert == "Sox2P" & sample_type == "control pool")) %>%
  select(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name,
         read_count, read_count_1, read_count_2)%>%
  #add merged controls back in
  bind_rows(tib_23_E2285_ctrls) %>%
  #mark hopped integrations
  mutate(hopped = start != 34643961)

#load the -161kb data (LP C9)
tib_C9 = read_tsv(path_tib_C9_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  #combine the 2 libraries of "Sox2P-reporter -161 kb, population P1 pool1 rep2"
  mutate(to_merge = case_when(replicate == "rep2" & insert == "Sox2P" & sorted_gate == "P1" ~ "merge",
                              .default = sample_name)) %>%
  mutate(sample_name = case_when(to_merge == "merge" ~ "Sox2P-reporter -161 kb, population P1 pool1 rep2",
                             .default = sample_name)) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  #mark hopped integrations
  mutate(hopped = start != 34598479)
  
#load the +715kb data (LP C10)
tib_C10 = read_tsv(path_tib_C10_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  #mark hopped integrations
  mutate(hopped = start != 35478413)
  
```

Sox2P mCh-

```{r}

#load the -161kb mCh- data (LP C9)
tib_C9_mChneg = read_tsv(path_tib_C9_mChneg_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  ungroup() %>%
  #mark hopped integrations
  mutate(hopped = start != 34598479) 

```

Sox2P_CDS
```{r}
#load the -161kb mCh+ Sox2P_CDS data (C9_mCh_41)
tib_C9_41 = read_tsv(path_tib_C9_41_ints) %>%
  filter(chr %in% paste0("chr", c(1:19, "X"))) %>%
  mutate(chr = factor(chr, levels = paste0("chr", c(1:19, "X")))) %>%
  group_by(chr, start, end, seq, strand, replicate, experiment, launch_pad_location, insert, genotype, sample_type, sorted_gate, pool_nr, sample_name) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2), .groups = "keep") %>%
  ungroup() %>%
  #mark hopped integrations
  mutate(hopped = start != 34598479)
```

combine and filter
```{r}
#filter the integrations
tib_filtered <- bind_rows(tib_23, tib_C9, tib_C10, tib_C9_41, tib_C9_mChneg)  %>%
  #add variables used in the plotting (old names of cell lines etc) 
  mutate(population = case_when(sample_type == "control pool" ~ "ctrl",
                                .default = sorted_gate),
         cell_line = case_when(launch_pad_location == "-116 kb" & insert == "Sox2P" ~ "23_34A",
                               launch_pad_location == "-161 kb" & insert == "Sox2P" & genotype == "WT" ~ "C9_mCh_34",
                               launch_pad_location == "-161 kb" & insert == "Sox2P" & genotype == "Sox2::mCherry deleted" ~ "C9_mCh-_34",
                               launch_pad_location == "-161 kb" & insert == "Sox2P_CDS" & genotype == "WT" ~ "C9_mCh_41",
                               launch_pad_location == "+715 kb" & insert == "Sox2P" ~ "C10_34"),
         landing_pad = case_when(launch_pad_location == "-116 kb" ~ "23",
                                 launch_pad_location == "-161 kb" ~ "C9",
                                 launch_pad_location == "+715 kb" ~ "C10")) %>%
  #annotate confidence of mapping (one or both ITRs)
  mutate(mapped_arms = case_when(read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only",
                                 read_count_1 > 0 & read_count_2 > 0 ~ "both_arms")) %>%
  #filter the data
  filter(population == "ctrl" | read_count >= 2) %>% #for ctrl distribution 1 mapping read is enough
  filter(!(start >= 34721183 & start <= 34721192)) %>% #remove contamination from LP8
  filter(! start %in% c(25018987, 35232946) ) %>% #remove contamination clone C1
  filter(landing_pad  == "C9" | start != 34598479) %>% #remove contamination from LP C9
  filter(landing_pad  == "23" | start != 34643961) %>% #remove contamination from LP 23
  filter(landing_pad  == "C10" | start != 35478413) %>% #remove contamination from LP C10
  filter(strand %in% c("+", "-"))


tib_filtered_P1_strict = tib_filtered %>%
  filter(!(population == "P1" & mapped_arms != "both_arms")) #only keep P1 integrations with 2 mapped arms

all_exp = unique(tib_filtered_P1_strict$experiment)
```


# make tables
```{r}

tib_filtered_P1_strict %>% 
  filter(population != "ctrl" & chr == "chr3") %>%
  group_by(cell_line) %>%
  summarize(N = n())

```
Shape of the table:
one block per cell line, one row per population
one column per location (genomewide, chr3, within 2Mb region)
```{r}

tib_stats_per_pop = tib_filtered_P1_strict %>% 
  mutate(population = factor(population, levels = c("ctrl", "P1_H", "P1", "P1_L", "P2_H", "P2", "P2_L", paste0("P", 3:6)))) %>%
  mutate(chr3 = chr == "chr3",
         in_wide_locus = start >= range_2Mb[1] & start <= range_2Mb[2],
         in_narrow_locus = start >= prange_zoom_paper[1] & start <= prange_zoom_paper[2]) %>%
  group_by(cell_line, population) %>%
  summarize(N_genomewide = n(),
            # N_chr3 = sum(chr3),
            N_in_wide_locus = sum(in_wide_locus),
            N_in_narrow_locus = sum(in_narrow_locus))

tib_stats_per_cl = tib_filtered_P1_strict %>% 
  mutate(population = factor(population, levels = c("ctrl", "P1_H", "P1", "P1_L", "P2_H", "P2", "P2_L", paste0("P", 3:6)))) %>%
  mutate(chr3 = chr == "chr3",
         in_wide_locus = start >= range_2Mb[1] & start <= range_2Mb[2],
         in_narrow_locus = start >= prange_zoom_paper[1] & start <= prange_zoom_paper[2]) %>%
  group_by(cell_line) %>%
  summarize(N_genomewide = n(),
            # N_chr3 = sum(chr3),
            N_in_wide_locus = sum(in_wide_locus),
            N_in_narrow_locus = sum(in_narrow_locus))

formatted_stats_tib = tib_stats_per_cl %>% 
  mutate(population = "all") %>%
  bind_rows(tib_stats_per_pop) %>%
  select(cell_line, population, N_genomewide, N_in_wide_locus, N_in_narrow_locus) %>%
  mutate(population = 
           factor(population, levels = c("all", "ctrl", "P1_H", "P1", "P1_L", "P2_H", "P2", "P2_L", paste0("P", 3:6))),
         cell_line = factor(cell_line, levels = c("C9_mCh_34", "23_34A", "C10_34", "C9_mCh-_34", "C9_mCh_41"))) %>%
  arrange(cell_line, population) %>%
  mutate(cell_line_name = case_when(cell_line == "23_34A" ~ "Sox2P-reporter -116 kb",
                                    cell_line == "C10_34" ~ "Sox2P-reporter +715 kb",
                                    cell_line == "C9_mCh_34" ~ "Sox2P-reporter -161 kb",
                                     cell_line == "C9_mCh-_34" ~ "Sox2P-reporter -161 kb Sox2::mCherry deleted",
                                    cell_line == "C9_mCh_41" ~ "Sox2P_CDS-reporter -161 kb"))
  
# write_tsv(formatted_stats_tib, "/DATA/projects/Sox2/Figure_reporter_only_hopping/CM20250424_integration_stats.tsv")
```


```{r}

filter(tib_filtered_P1_strict, hopped & experiment %in% all_exp & cell_line %in% c("23_34A", "C9_mCh_34", "C10_34")) %>%
  filter(population != "ctrl") %>%
  group_by(cell_line) %>%
  summarize(N = n())

```


# calculate and add proof-of-principle experiment


```{r}
tib_E2226 <- read_tsv(path_CRE6_ints)

# Change mapping quality to numeric
tib_E2226 <- tib_E2226 %>%
  mutate(mapq_1 = as.numeric(mapq_1),
         mapq_2 = as.numeric(mapq_2),
         mapq_1 = replace_na(mapq_1, 0),
         mapq_2 = replace_na(mapq_2, 0))

# Factors for chr
chromosomes <- paste0("chr", c(1:19, "X"))

tib_E2226 <- tib_E2226 %>%
  mutate(chr = factor(chr, levels = chromosomes)) %>%
  drop_na(chr)

tib_E2226 = tib_E2226 %>% 
  mutate(cell_line = "CRE6",
         population = "ctrl") %>%
  mutate(mapped_arms = case_when(read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only",
                                 read_count_1 > 0 & read_count_2 > 0 ~ "both_arms")) %>%
  mutate(hopped = start != 34643961) 

#Filter table
tib_E2226_filt = tib_E2226 %>% 
  filter(population == "ctrl" | read_count >= 2) %>% #require at least two reads except for ctrl
  filter(!(start >= 34721183 & start <= 34721192)) %>% #remove contamination from LP8
  filter(strand %in% c("+", "-")) #because we sometimes split by strand, I'd prefer to remove the ambiguous integrations from all analyses

#E2226 had 10 control reactions sequenced with separate indices, here we combine them to get comparable data to the other experiments (with 1 index pair for the control pool)
tib_E2226_comb = tib_E2226_filt %>%
  group_by(cell_line, population, chr, start, end, seq, strand, experiment) %>%
  summarize(read_count = sum(read_count),
            read_count_1 = sum(read_count_1),
            read_count_2 = sum(read_count_2)) %>%
  mutate(sample = paste0(cell_line, "_", population))  %>%
  mutate(mapped_arms = case_when(read_count_1 == 0 ~ "rv_only",
                                 read_count_2 == 0 ~ "fw_only",
                                 read_count_1 > 0 & read_count_2 > 0 ~ "both_arms")) %>%
  mutate(hopped = start != 34643961) 


E2226_stats_tib = tib_E2226_comb %>%
  mutate(chr3 = chr == "chr3",
         in_wide_locus = start >= range_2Mb[1] & start <= range_2Mb[2],
         in_narrow_locus = start >= prange_zoom_paper[1] & start <= prange_zoom_paper[2]) %>%
  group_by(cell_line) %>%
  summarize(N_genomewide = n(),
            # N_chr3 = sum(chr3),
            N_in_wide_locus = sum(in_wide_locus),
            N_in_narrow_locus = sum(in_narrow_locus))

E2226_stats_tib
```